### 2021-10-13 README.Rmd
initial commit

---
title: "README"
author: "Roland Hassan Nasser"
date: "13/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## digiRhythm

digiRhythm is an R library developed at Agroscope and provides a set of tools to 
analyze and visualize the rhythmic behavior of animals.

**TO DO:**

- [ ] Define data inside the library
- [ ] Create a function that read Icetag data stored in a CSV file.
- [ ] Add the previous function as an example to the README file.
- [ ] Create a function that visualize the actogram.
- [ ] Add the previous function as an example to the README file.
- [ ] Create test functions.
- [ ] Create a function that computes the degree of functional coupling.
- [ ] Create a function that computes the diurnality index.
- [ ] Shiny app?

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
#devtools::install_github("nasserdr/digiRhythm")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#library(digiRhythm)
## basic example code
```

### 2021-10-29 README.Rmd
Update the readme file with some more examples

@@
 ---
-title: "README"
-author: "Roland Hassan Nasser"
-date: "13/10/2021"
-output: html_document
+title: "readme"
+output: github_document
 ---
@@
 
-**TO DO:**
+**TO DO CORE FUNCTIONALITIES:**
 
-- [ ] Define data inside the library
-- [ ] Create a function that read Icetag data stored in a CSV file.
-- [ ] Add the previous function as an example to the README file.
-- [ ] Create a function that visualize the actogram.
-- [ ] Add the previous function as an example to the README file.
+- [X] Create a function that read Icetag data stored in a CSV file.
+- [X] Add the import_raw_icetage_data as an example to the README file.
+- [X] Create a function that computes the degree of functional coupling.
+- [X] Add the the dfc as an example to the README file.
+- [ ] Create a function that computes the diurnality index.
+- [ ] Add the the dfc as an example to the README file.
+
+**TO DO - VISUALIZATION FUNCTIONS:**
+
+- [X] Create a function that visualize the actogram.
+- [ ] Create a function that visualize the actogram + average activity.
+- [ ] Create a function that visualize the actogram + average activity + DFC.
+
+**TO DO - UTILITIES AND DOCUMENTATIONS:**
+
+- [X] Define data inside the library
+- [X] Add and example about the improt_raw_icetag_data to the README.
+- [ ] Create a function to test if a dataset is digiRhythm friendly.
+- [ ] Create a proper documentation for the data set.
+- [ ] configure data sets with lazy loading
 - [ ] Create test functions.
-- [ ] Create a function that computes the degree of functional coupling.
-- [ ] Create a function that computes the diurnality index.
-- [ ] Shiny app?
 
@@
 
-``` r
-# install.packages("devtools")
-#devtools::install_github("nasserdr/digiRhythm")
+```{r}
+#Uncomment the below two lines if you use the library for the first time of if
+#you want to update the library
+#install.packages("devtools")
+#devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
 ```
@@
 
-```{r example}
-#library(digiRhythm)
-## basic example code
+```{r example, message=FALSE, warning=FALSE, paged.print=FALSE}
+library(digiRhythm)
+
+#The file name with the path
+url <- 'https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv'
+download.file(url, destfile = '516b_2.csv')
+filename <- file.path(getwd(), '516b_2.csv')
+
+#The columns that we are interested in
+colstoread <- c("Date", "Time", "Motion Index", 'Steps')
+
+#Reading the activity data from the csv file
+data <- import_raw_icetag_data(filename = filename, skipLines = , act.cols.names = colstoread, sampling = 15)
+
+print(head(data))
+
 ```
+
+This is an example on how to visualize the actogram
+
+```{r visualize, message=FALSE, warning=FALSE, paged.print=FALSE}
+data("df516b_2", package = "digiRhythm")
+df <- remove_activity_outliers(df516b_2)
+df_act_info(df)
+activity = names(df)[2]
+start <- "2020-30-04"
+end <- "2020-06-05"
+save <- TRUE
+outputdir <- 'testresults'
+outplotname <- 'myplot'
+width <- 10
+device <- 'tiff'
+height <-  5
+actogram(data, activity, start, end, save = FALSE,
+     outputdir = 'testresults', outplotname = 'actoplot', width = 10,
+     height =  5, device = 'tiff')
+```
+
+This is an example on how to compute the degree of functional coupling.
+
+```{r dfc_example, results = FALSE, warning=FALSE, paged.print=FALSE}
+data("df516b_2", package = "digiRhythm")
+df <- remove_activity_outliers(df516b_2)
+df_act_info(df)
+activity = names(df)[2]
+my_dfc <- dfc(df, activity , sampling = 15, show_lsp_plot = FALSE)
+
+#You may want to explore the two list inside my_dfc.
+#DFC and SPECTRUM are saved inside my_dfc, each as a list
+```
+
+This is an example on how to compute the diurnality index:
+
+```{r dindex_example, results = FALSE, warning=FALSE, paged.print=FALSE}
+data("df516b_2", package = "digiRhythm")
+df <- remove_activity_outliers(df516b_2)
+df_act_info(df)
+activity = names(df)[2]
+d_index <- diurnality(data, activity, plot = TRUE)
+
+```
+
+This is an example on how you can resample your data:
+
+```{r resample, results = FALSE, warning=FALSE, paged.print=FALSE}
+data("df516b_2", package = "digiRhythm")
+df <- df516b_2
+df <- remove_activity_outliers(df)
+new_sampling <- 30
+new_dgm <- resample_dgm(df, new_sampling)
+
+```

### 2021-10-29 R/import_raw_icetag_data.R

@@ -18,7 +18,7 @@
 #'                                act.cols.names = colstoread,
 #'                                sampling = 15,
 #'                                verbose = TRUE)
-
+#'
 #' @param filename The file name (full or relative path with extension)
 #' @param skipLines The number of non-useful lines to skip (lines to header)
 #' @param act.cols.names A vector containing the names of columns to read

### 2021-10-29 original_codes/original_import_raw_icetag_data.R

@@ -1,5 +1,6 @@
 library(readr) #read_csv
 library(tidyr) #unite
+library(xts)
 
 
 url <- 'https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv'

### 2021-10-29 R/is_dgm_friendly.R
Updated is_dgm_friendly so that it notifies when the data is null or when it contains less than 7 days.

@@ -21,28 +21,47 @@ is_dgm_friendly <- function(data, verbose = FALSE){
 
   is_dgm <- TRUE
 
-  if("POSIXct" %in%  class(data[,1])  | "POSIXt" %in%  class(data[,1])){
-    message <- paste0(green('v Correct time format: '), 'First column has a Posixct Format')
-    print_v(message, verbose)
-  }else {
-    message <- paste0(red('x Inorrect time format: '), 'First column does not a Posixct Format')
+  if(is.null(data)){
+    message <- red('Data cannot be NULL')
     print_v(message, verbose)
     is_dgm <- FALSE
-  }
+  } else{
+    #Checking if the first column meets the requirements
+    if("POSIXct" %in%  class(data[,1])  | "POSIXt" %in%  class(data[,1])){
+      message <- paste0(green('v Correct time format: '), 'First column has a POSIXct Format')
+      print_v(message, verbose)
+    }else {
+      message <- paste0(red('x Inorrect time format: '), 'First column does not a POSIXct Format')
+      print_v(message, verbose)
+      is_dgm <- FALSE
+    }
 
 
-  if(ncol(data) == 1){
-    message <- paste0(red('x Illogical number of columns: '), 'The dataset has only one column. Minimum number of columns is 2')
-    print_v(message, verbose)
-    is_dgm <- FALSE
-  }else{
-    for(i in 2:ncol(data)){
-      if("numeric" %in% class(data[,i]) | "integer" %in%  class(data[,i])){
-        message <- green(paste('v Correct numeric format - Column', i, '==>', colnames(data)[i]))
-        print_v(message, verbose)
-      } else {
-        message <- red(paste('x Inorrect numeric format - Column', i, '==>', colnames(data)[i]))
-        print_v(message, verbose)
+    #Checking if data contains at least 7 days of data
+    if(length(unique(as.Date(data[,1]))) >= 7){
+      message <- paste0(green('v Number of days: '), 'Bigger or equal to 7')
+      print_v(message, verbose)
+    }else {
+      message <- paste0(red('x Number of days: '), 'Less than 7 (can\'t run the DFC algorithm later on')
+      print_v(message, verbose)
+      is_dgm <- FALSE
+    }
+
+    #Checking if we have more than one column
+    if(ncol(data) == 1){
+      message <- paste0(red('x Illogical number of columns: '), 'The dataset has only one column. Minimum number of columns is 2')
+      print_v(message, verbose)
[…]

### 2021-10-29 original_codes/original_is_dgm_friendly.R

@@ -1,36 +1,54 @@
 library(crayon)
+library(digiRhythm)
 data("df516b_2", package = "digiRhythm")
 data <- df516b_2
 
 verbose = TRUE
 is_dgm <- TRUE
 
-#Add isNull option
-#Add minimum 2 days options
-#Add warning that if less than 7 days we cant compute the DFC
 
-if("POSIXct" %in%  class(data[,1])  | "POSIXt" %in%  class(data[,1])){
-  message <- paste0(green('v Correct time format: '), 'First column has a Posixct Format')
-  print_v(message, verbose)
-}else {
-  message <- paste0(red('x Inorrect time format: '), 'First column does not a Posixct Format')
+if(is.null(data)){
+  message <- red('Data cannot be NULL')
   print_v(message, verbose)
   is_dgm <- FALSE
-}
+} else{
+
+  #Checking if the first column meets the requirements
+  if("POSIXct" %in%  class(data[,1])  | "POSIXt" %in%  class(data[,1])){
+    message <- paste0(green('v Correct time format: '), 'First column has a POSIXct Format')
+    print_v(message, verbose)
+  }else {
+    message <- paste0(red('x Inorrect time format: '), 'First column does not a POSIXct Format')
+    print_v(message, verbose)
+    is_dgm <- FALSE
+  }
 
 
-if(ncol(data) == 1){
-  message <- paste0(red('x Illogical number of columns: '), 'The dataset has only one column. Minimum number of columns is 2')
-  print_v(message, verbose)
-  is_dgm <- FALSE
-}else{
-  for(i in 2:ncol(data)){
-    if("numeric" %in% class(data[,i]) | "integer" %in%  class(data[,i])){
-      message <- green(paste('v Correct numeric format - Column', i, '==>', colnames(data)[i]))
-      print_v(message, verbose)
-    } else {
-      message <- red(paste('x Inorrect numeric format - Column', i, '==>', colnames(data)[i]))
-      print_v(message, verbose)
+  #Checking if data contains at least 7 days of data
+  if(length(unique(as.Date(data[,1]))) >= 7){
+    message <- paste0(green('v Number of days: '), 'Bigger or equal to 7')
+    print_v(message, verbose)
+  }else {
+    message <- paste0(red('x Number of days: '), 'Less than 7 (can\'t run the DFC algorithm later on')
+    print_v(message, verbose)
+    is_dgm <- FALSE
+  }
+
[…]

### 2021-10-29 README.Rmd
Added the starting point of the na_dgm function

@@
 - [X] Add the the dfc as an example to the README file.
-- [ ] Create a function that computes the diurnality index.
-- [ ] Add the the dfc as an example to the README file.
+- [x] Create a function that computes the diurnality index.
+- [x] Add the the dfc as an example to the README file.
+- [ ] Consider adding other sample data sets (check which ones appear to have
+non-noisy data from the previous analysis and visualizations).
+- [ ] Enhance the actogram (from the original codes).
+- [ ] Add a function that fills NA values in the activity data set.
+- [ ] Add a function that creates a histogram for the significant and harmonic
+frequencies within the DFC (Patricia's case).
 
@@
 - [ ] Create a function that visualize the actogram + average activity + DFC.
+- [ ] Create a function that visualizes all daily activities with control over
+the number of columns.
 
@@
 - [X] Define data inside the library
-- [X] Add and example about the improt_raw_icetag_data to the README.
-- [ ] Create a function to test if a dataset is digiRhythm friendly.
-- [ ] Create a proper documentation for the data set.
-- [ ] configure data sets with lazy loading
-- [ ] Create test functions.
+- [X] Add and example about the import_raw_icetag_data to the README.
+- [x] Create a function to test if a data set is digiRhythm friendly.
+- [x] Create a proper documentation for the data set.
+- [x] configure data sets with lazy loading
+- [ ] Create test functions for all the functionalities.
+- [ ] Add a note about the fact that visualizations were created in a way they
+are aligned with the standards of common journals and give details about these 
+settings + Add a note where the code can be customized to change these settings.
+- [ ] Make a series of videos about the digiRhythm library.
 
@@
 ```{r visualize, message=FALSE, warning=FALSE, paged.print=FALSE}
-data("df516b_2", package = "digiRhythm")
-df <- remove_activity_outliers(df516b_2)
+data("df691b_1", package = "digiRhythm")
+df <- remove_activity_outliers(df691b_1)
 df_act_info(df)
 activity = names(df)[2]
-start <- "2020-30-04"
-end <- "2020-06-05"
+start <- "2020-08-25"
+end <- "2020-10-11"
 save <- TRUE
@@
 height <-  5
-actogram(data, activity, start, end, save = FALSE,
+actogram(df, activity, start, end, save = FALSE,
      outputdir = 'testresults', outplotname = 'actoplot', width = 10,

### 2021-10-29 original_codes/original_actogram.R

@@ -1,4 +1,5 @@
 library(ggplot2)
+library(xts)
 data("df516b_2")
 df <- df516b_2
 df <- remove_activity_outliers(df)
@@ -38,10 +39,10 @@ df_filtered$date <- as.Date(df_filtered[[activity]], tz = "CET", origin = df_fil
 df_filtered$time <- format(df_filtered[,1], format = "%H:%M", tz = "CET")
 
 
-p <- ggplot(df_filtered,
+ggplot(df_filtered,
             aes(x = time,
                 y = date,
-                color = activity)) +
+                color = Motion.Index)) +
   geom_tile() +
   ylab("Date") +
   xlab("Time") +
@@ -52,8 +53,16 @@ p <- ggplot(df_filtered,
     panel.background = element_rect(fill = "white"),
     axis.line = element_line(size = 0.5),
   ) +
-  scale_x_discrete(breaks = c("03:00", "09:00", "15:00", "21:00")) +
-  theme(legend.position = "none")
+  scale_colour_gradient(
+    low = "#000000",
+    high = "#FFFFFF",
+    space = "Lab",
+    na.value = "grey50",
+    guide = "colourbar",
+    aesthetics = "colour"
+  ) +
+  scale_x_discrete(breaks = c("03:00", "09:00", "15:00", "21:00"))
+#  theme(legend.position = "none")
 
 if (save == TRUE) {
 

### 2021-10-29 original_codes/original_fill_na_dgm.R

@@ -1,0 +1,15 @@
+#imports
+library(xts) #xts #period.apply #endpoints
+library(lubridate) #date
+library(ggplot2)
+
+activity <- 'Motion.Index'
+
+#Configs
+data("df516b_2", package = "digiRhythm")
+data <- df516b_2
+
+
+data <- remove_activity_outliers(data)
+df_act_info(data)
+activity = names(data)[2]

### 2022-07-05 R/import_raw_activity_data.R
Update vignettes and one dataset. Added NA check in import_activity function

@@ -101,6 +101,9 @@ import_raw_activity_data <- function(filename,
                      skip = skipLines,
                      delim = sep,
                      show_col_types = FALSE)[, act.cols.names]
+
+  data <- na.omit(data)
+
   data <- data %>%
     mutate(across(where(is.character), str_trim))
 
@@ -225,6 +228,7 @@ import_raw_activity_data <- function(filename,
   }
 
   df = df[!duplicated(df$datetime),]
+
 
  gc()
  return(df)

### 2022-07-05 R/utils.R

@@ -36,18 +36,18 @@ df_act_info <- function(df){
 
 remove_activity_outliers <- function(df){
 
-  data <- df
+  data <- as.data.frame(df)
   for (i in 2:ncol(df)) {
-    Q <- quantile(data[,i], probs = c(.25, .75), na.rm = FALSE)
+    Q <- quantile(data[,i], probs = c(.25, .75), na.rm = TRUE)
     iqr <- IQR(data[,i])
     up <-  Q[2] + 1.5*iqr # Upper Range
     low <- Q[1] - 1.5*iqr # Lower Range
     non_outliers <- which(data[,i] <= up | data[,i] >= low)
     outliers <- which(data[,i] > up | data[,i] < low)
-    mean_without_outliers <- mean(data[non_outliers,i])
+    mean_without_outliers <- mean(data[non_outliers,i], na.rm = TRUE)
     data[outliers, i] <- mean_without_outliers
   }
-  return(data)
+  data
 }
 
 

### 2022-07-05 original_codes/original_import_raw_activity_data.R

@@ -6,24 +6,24 @@ library(dplyr)
 library(stringr)
 
 #Read a sample file from github
-# url <- 'https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv'
-# download.file(url, destfile = '516b_2.csv')
-# filename <- file.path(getwd(), '516b_2.csv')
-# act.cols.names <- c("Date", "Time", "Motion Index", 'Steps')
-# date_format <- "%d.%m.%Y"
-# time_format <- "%H:%M:%S"
-# sep = ','
-# skiplines <- 7
+url <- 'https://raw.githubusercontent.com/nasserdr/digiRhythm_sample_datasets/main/689b_3.csv'
+download.file(url, destfile = 'df689b_3.csv')
+filename <- file.path(getwd(), 'df689b_3.csv')
+act.cols.names <- c("Date", "Time", "Motion Index", 'Steps')
+date_format <- "%d.%m.%Y"
+time_format <- "%H:%M:%S"
+sep = ','
+skipLines <- 7
 
 #OR Read a file from local file system (Aska's data)
-dir <- '~/mnt/Data-Work-RE/26_Agricultural_Engineering-RE/262.2_VT_Nutztierhaltung/Rhythmizität_Milchkühe/PM_4_semaines/rawdatamin/raw_data_binded/all_variables_separated_missing_days'
-file <- list.files(dir)[26]
-filename <- file.path(dir, file)
-act.cols.names <- c("Date", "Time", "move_x", 'move_y')
-date_format <- "%Y-%m-%d"
-time_format <- "%H:%M:%S"
-sep = ';'
-skipLines <- 0
+# dir <- '~/mnt/Data-Work-RE/26_Agricultural_Engineering-RE/262.2_VT_Nutztierhaltung/Rhythmizität_Milchkühe/PM_4_semaines/rawdatamin/raw_data_binded/all_variables_separated_missing_days'
+# file <- list.files(dir)[26]
+# filename <- file.path(dir, file)
+# act.cols.names <- c("Date", "Time", "move_x", 'move_y')
+# date_format <- "%Y-%m-%d"
+# time_format <- "%H:%M:%S"
+# sep = ';'
+# skipLines <- 0
 
 
 #OR Read a file from local file system (Marie's data)
@@ -54,6 +54,9 @@ data <- read_delim(filename,
                    skip = skipLines,
                    delim = sep,
[…]

### 2022-07-05 R/actogram.R
Updated for cran changes

@@ -29,8 +29,7 @@
 #' start <- "2020-05-01" #year-month-day
 #' end <- "2020-08-13" #year-month-day
 #' activity_alias <- 'Motion Index'
-#' save <- 'sample_results/actogram' #if NULL, don't save the image
-#' actogram(df, activity, activity_alias, start, end, save)
+#' actogram(df, activity, activity_alias, start, end, save = NULL)
 
 actogram <- function(
   df,

### 2022-07-07 R/globals.R
Revamped the Lomb Scargle Periodogram function and made it more customized to the DigiRhythm package. Removed the lomb dependency. Updated a couple of examples and vignettes.

@@ -5,5 +5,12 @@ utils::globalVariables(
     'where',
     'na.omit',
     'Date',
-    'Diurnalit')
+    'Diurnality',
+    'fibsearch',
+    'frequency_hz',
+    'period_seconds',
+    'status_harmonic',
+    'power',
+    'period_hours',
+    'new_h')
   )

### 2022-07-07 R/lomb_scargle_periodogram.R

@@ -1,0 +1,194 @@
+#' Computes the Lomb Scargle Periodogram and returns the information needed for
+#' computing the DFC and HP. A plot visualizing the Harmonic Frequencies presence
+#' in the spectrum is possible. The function is inspired from the Lomb library in
+#' a great part, with modifications to fit the requirements of harmonic powers and
+#' computation of the DFC
+#'
+#' @param data a digiRhythm friendly dataframe of only two columns
+#' @param alpha the statistical significance for the false alarm
+#' @param plot if TRUE, the LSP will be plotted
+#'
+#' @return a list that contains a dataframe (detailed below), the significance
+#' level and alpha (for the record). The dataframe contains the power the frequency,
+#' the frequency in HZ, the p values according to Baluev 2008, the period that corresponds
+#' to the frequency in seconds and in hours and finally, a boolean to tell whether
+#' the frequency is harmonic or not.
+#'
+#' @import ggplot2
+#' @importFrom pracma fibsearch
+#'
+#' @export
+#'
+#' @examples
+#' data("df516b_2", package = "digiRhythm")
+#' data <- df516b_2[1:672, c(1,2)]
+#' sig <- 0.01
+#' lomb_scargle_periodogram(data, alpha = sig, plot = TRUE)
+
+
+lomb_scargle_periodogram <- function (data, alpha = 0.01, plot = TRUE) {
+
+  if (!is_dgm_friendly(data, verbose = TRUE)) {
+    warning('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
+  }
+
+  x <- data
+  start <- as.Date(x[1,1])
+  # print(start)
+  end <- as.Date(x[nrow(x),1])
+  # print(end)
+
+  from_to = paste('From',
+                  start,
+                  'To',
+                  end)
+
+  ofac <- 1
+  names <- colnames(x)
+  times <- x[, 1]
+  x <- x[, 2]
+
+  times <- times[!is.na(x)]
+  x <- x[!is.na(x)]
+
+  nobs <- length(x)
+
+  times <- as.numeric(times)
+  start <- min(times)
+  end <- max(times)
+  av.int <- mean(diff(times))
[…]

### 2022-07-07 R/utils.R

@@ -72,7 +72,7 @@ print_v <- function(
 
 #' Returns the periodicity of a digiRhythm dataframe
 #'
-#' @param data a gigiRhythm friendly dataframe
+#' @param data a digiRhythm friendly dataframe
 #'
 #' @importFrom xts periodicity as.xts
 #'
@@ -92,3 +92,46 @@ dgm_periodicity <- function(data){
   xts::periodicity(xts_data)
 
 }
+
+
+#' Returns p-value of a frequency peack according to pbaluev (2008) given Z,
+#' fmax and tm. Copied from the LOMB library.
+#'
+#' @param Z the power of the frequency
+#' @param fmax the maximum frequency in the spectrum
+#' @param tm the time grid of the original time series
+#'
+#' @export
+
+#Copied from the LOMB library
+pbaluev <- function(Z, fmax, tm) {
+  #code copied from astropy timeseries (https://docs.astropy.org/en/stable/timeseries/index.html)
+  N = length(tm)
+  Dt = mean(tm^2) - mean(tm)^2
+  NH = N - 1
+  NK <- N - 3
+  fsingle <- (1 - Z) ^ (0.5 * NK)
+  Teff <- sqrt(4 * pi * Dt) # Effective baseline
+  W <- fmax * Teff
+  ggamma_NH <- sqrt(2 / N) * exp(lgamma(N / 2) - lgamma((N - 1) / 2))
+  tau <- ggamma_NH * W * (1 - Z) ^ (0.5 * (NK - 1))*sqrt(0.5 * NH * Z)
+  p = -(exp(-tau) - 1) + fsingle * exp(-tau)
+  return(p)
+}
+
+#' Returns the level given the p-value computed with pbaluev (2008). Copied from
+#'  the LOMB library.
+#'
+#' @param Z the power of the frequency
+#' @param fmax the maximum frequency in the spectrum
+#' @param tm the time grid of the original time series
+#' @param alpha the significance level
+#'
+#' @export
+
+
+levopt <-  function(Z,alpha,fmax,tm){
+  prob = pbaluev(Z,fmax,tm)
+  (log(prob) - log(alpha))^2
+}
+

### 2022-07-07 examples/code_snipets_statsoft_paper.R

@@ -1,0 +1,32 @@
+library(ggplot2)
+library(magrittr)
+library(lubridate)
+library(dplyr)
+library(lomb)
+#Example
+data("df516b_2")
+df <- df516b_2
+
+head(df)
+str(df)
+
+# Fourier transform for signals added together ----------------------------
+
+par(mfrow=c(3,2), mar = c(2, 2, 2, 2))
+plot(t_grid, signal0.5, type = 'l', xlab = "", ylab = "", main = "1/2 Hz")
+t_grid = seq(0, 9.99, sampling)
+ft <- Mod(fft(signal0.5))
+f_grid <- seq(0, 1/(2*sampling), length.out = N/2+1)
+plot(f_grid, ft[1:(N/2+1)], type = 'l')
+
+plot(t_grid, signal4, type = 'l', xlab = "", ylab = "", main = "4 Hz")
+ft <- Mod(fft(signal4))
+f_grid <- seq(0, 1/(2*sampling), length.out = N/2+1)
+plot(f_grid, ft[1:(N/2+1)], type = 'l')
+
+
+combi1 = signal0.5 + signal4
+plot(t_grid, combi1, type = 'l', xlab = "", ylab = "", main = "0.5 Hz + 4 Hz")
+ft <- Mod(fft(combi1))
+f_grid <- seq(0, 1/(2*sampling), length.out = N/2+1)
+plot(f_grid, ft[1:(N/2+1)], type = 'l')

### 2022-07-07 old_codes/dfc_old.R

@@ -1,4 +1,3 @@
-
 #' Computes the Degree of Function coupling (DFC), Harmonic Part (HP) and Weekly
 #' Lomb-Scargle Spectrum (LSP Spec) for one variable in an activity dataset.
 #' The dataset should be digiRhythm friendly.
@@ -24,8 +23,6 @@
 #' @param data The activity data set.
 #' @param activity The name of the activity.
 #' @param sampling The sampling period of the data set in minutes.
-#' @param sliding_window_day recommended to be 7 days. The sliding window on which
-#' the Lomb Scargle Periodogram is computed.
 #' @param sig The significance level that should be used to determine the
 #' significant frequency component.
 #' @param plot if TRUE, the DFC/HP plot will be shown.
@@ -47,26 +44,28 @@
 #' @importFrom lubridate date
 #' @importFrom dplyr filter
 #' @importFrom stats ts
+#' @importFrom lomb lsp
 #'
 #' @export
 #' @examples
 #' data("df516b_2", package = "digiRhythm")
-#' df <- df516b_2[1:672, c(1,2)]
+#' df <- df516b_2
 #' df <- remove_activity_outliers(df)
 #' df_act_info(df)
-#' activity = names(df)[2]
-#' my_dfc <- dfc(df, activity, sliding_window_day = 3, sampling = 15)
+#' activity <- names(df)[2]
+#' my_dfc <- dfc(df, activity , sampling = 15)
+
+
 
 #######################################################
 dfc <- function(
-    data,
-    activity,
-    sampling = 15, #in minutes
-    sliding_window_day = 7,
-    sig = 0.05,
-    plot = TRUE,
-    plot_harmonic_part = TRUE,
-    verbose = TRUE
+  data,
+  activity,
+  sampling = 15, #in minutes
+  sig = 0.05,
+  plot = TRUE,
+  plot_harmonic_part = TRUE,
+  verbose = TRUE
 )
 {
 
@@ -85,8 +84,8 @@ dfc <- function(
               by = 1)
 
 
-  if (length(days) < sliding_window_day) {
-    stop(paste('You need at least', sliding_window_day, 'days of data to run the Degree of Functional Coupling algorithm'))
+  if (length(days) < 7) {
+    stop('You need at least 7 days of data to run the Degree of Functional Coupling algorithm')
   }
 
   ##Change (removed)
@@ -109,21 +108,21 @@ dfc <- function(
[…]

### 2022-07-15 R/dfc.R
internal updates

@@ -24,7 +24,6 @@
 #' @param data The activity data set.
 #' @param activity The name of the activity.
 #' @param sampling The sampling period of the data set in minutes.
-#' @param sliding_window_day recommended to be 7 days. The sliding window on which
 #' the Lomb Scargle Periodogram is computed.
 #' @param sig The significance level that should be used to determine the
 #' significant frequency component.
@@ -55,14 +54,13 @@
 #' df <- remove_activity_outliers(df)
 #' df_act_info(df)
 #' activity = names(df)[2]
-#' my_dfc <- dfc(df, activity, sliding_window_day = 3, sampling = 15)
+#' my_dfc <- dfc(df, activity, sampling = 15)
 
 #######################################################
 dfc <- function(
     data,
     activity,
     sampling = 15, #in minutes
-    sliding_window_day = 7,
     sig = 0.05,
     plot = TRUE,
     plot_harmonic_part = TRUE,
@@ -85,9 +83,6 @@ dfc <- function(
               by = 1)
 
 
-  if (length(days) < sliding_window_day) {
-    stop(paste('You need at least', sliding_window_day, 'days of data to run the Degree of Functional Coupling algorithm'))
-  }
 
   ##Change (removed)
   # if (length(which(diff(days) != 1)) > 0) {
@@ -109,21 +104,21 @@ dfc <- function(
                      power = numeric(),
                      pvalue = numeric()) #The data frame for SPEC
 
-  n_days_scanned <- length(days) - sliding_window_day - 1
+  n_days_scanned <- length(days) - 6
 
   i = 1
 
   for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
 
     if (verbose) {
-      cat("Processing dates ", as.character(days[i]), " until ", as.character(days[(i + sliding_window_day - 1)]), "\n")
+      cat("Processing dates ", as.character(days[i]), " until ", as.character(days[(i + 6)]), "\n")
 
     }
 
     samples_per_day = 24*60/sampling #The number of data points per day
 
     #Filtering the next seven days by date (not by index - in case of missing data, filtering by index would make errors)
-    data_week <- df %>% filter(date >= days[i]) %>%  filter(date <= days[i + sliding_window_day - 1])
+    data_week <- df %>% filter(date >= days[i]) %>%  filter(date <= days[i + 6])
 
[…]

### 2022-07-15 R/lomb_scargle_periodogram.R

@@ -5,6 +5,7 @@
 #' computation of the DFC
 #'
 #' @param data a digiRhythm friendly dataframe of only two columns
+#' @param sampling the sampling period in minutes. default = 15 min.
 #' @param alpha the statistical significance for the false alarm
 #' @param plot if TRUE, the LSP will be plotted
 #'
@@ -26,10 +27,15 @@
 #' lomb_scargle_periodogram(data, alpha = sig, plot = TRUE)
 
 
-lomb_scargle_periodogram <- function (data, alpha = 0.01, plot = TRUE) {
+lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = TRUE) {
 
   if (!is_dgm_friendly(data, verbose = TRUE)) {
-    warning('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
+    stop('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
+  }
+
+  if (length(unique(as.Date(data[,1]))) != 7 ) {
+    warning('This LSP function is customized to use data with 7 days span only. But, the data contains less or more than 7 days.
+         A dynamic number of days will be introduced in a later version.')
   }
 
   x <- data
@@ -138,7 +144,7 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, plot = TRUE) {
   lsp_data <- lsp_data %>% mutate(period_seconds = (1/frequency_hz))
   lsp_data <- lsp_data %>% mutate(period_hours = period_seconds/3600)
 
-  harmonic_periods <- 24/seq(1,12)
+  harmonic_periods <- 24/seq(1,12) #24h, 12h, 8h, ....
   l <- lapply(harmonic_periods, function(x){
     which.min(abs(x-lsp_data$period_hours))
   })
@@ -150,9 +156,11 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, plot = TRUE) {
 
   output <- list(lsp_data = lsp_data, sig.level = level, alpha = alpha)
 
-  lsp_data <- lsp_data[1:96,]
+  len <- 24*60/sampling
+  lsp_data <- lsp_data[1:len,]
   if(plot){
     hdata <- lsp_data %>% filter(status_harmonic ==TRUE) %>%
+      # filter(power > 0.01) %>%
       select(frequency_hz, power, period_hours) %>%
[…]

### 2022-07-15 examples/code_snipets_statsoft_paper.R

@@ -2,31 +2,88 @@ library(ggplot2)
 library(magrittr)
 library(lubridate)
 library(dplyr)
+library(digiRhythm)
 library(lomb)
+library(patchwork)
+
+# LSP on purely synthetic data ----------------------------
+#Creating time grid
+sampling = 15*60 #15 minutes
+time = seq(
+  c(ISOdate(2022,3,20, 0,0,0)),
+  c(ISOdate(2022,3,26, 23, 59, 0)),
+  by = 15*60) #a timedate sequence over 1 week
+
+#Creating a signal with a period = 24 and computing its lsp
+period <- 24*3600 #period in seconds (24h * 3600 s)
+freq <- 1/period #freq in hertz
+sign24h <- sin(2*pi*freq*as.numeric(time))
+
+df <- data.frame(
+  datetime = time,
+  activity = sign24h
+)
+
+plot(df$datetime, df$activity, type = 'l', xlab = 'Signal', ylab = 'Date', main = 'Sine wave with a period of 24h')
+lomb_scargle_periodogram(df, alpha = 0.01, sampling = 15, plot = TRUE)
+
+#Creating a signal with a period = 12 and computing its lsp
+
+period <- 12*3600 #period in seconds (12h * 3600 s)
+freq <- 1/period #freq in hertz
+sign12h <- sin(2*pi*freq*as.numeric(time))
+
+df <- data.frame(
+  datetime = time,
+  activity = sign12h
+)
+
+lomb_scargle_periodogram(df, alpha = 0.01, sampling = 15, plot = TRUE)
+plot(df$datetime, df$activity, type = 'l', xlab = 'Signal', ylab = 'Date', main = 'Sine wave with a period of 12h')
+
+#Creating a signal with a period = 6 and computing its lsp
+
+period <- 6*3600 #period in seconds (6h * 3600 s)
+freq <- 1/period #freq in hertz
+sign06h <- sin(2*pi*freq*as.numeric(time))
+
+df <- data.frame(
+  datetime = time,
+  activity = sign06h
+)
+lomb_scargle_periodogram(df, alpha = 0.01, sampling = 15, plot = TRUE)
+plot(df$datetime, df$activity, type = 'l', xlab = 'Signal', ylab = 'Date', main = 'Sine wave with a period of 06h')
+
[…]

### 2022-07-16 R/daily_acticity_wrap_plot.R
Few internal changes after CRAN comments (added tempdir, removed @export tag for non-public functions, added @return tag for public functions)

@@ -31,7 +31,6 @@
 #' activity_alias <- 'Motion Index'
 #' start <- "2020-05-01" #year-month-day
 #' end <- "2020-05-07" #year-month-day
-#' save <- 'sample_results/daily_activity_wrap' #if NULL, don't save the image
 #' ncols <- 3
 #' sampling_rate <- 30
 #' my_dwp <- daily_activity_wrap_plot(df, activity, activity_alias, start, end, sampling_rate,

### 2022-07-16 R/utils.R

@@ -5,6 +5,8 @@
 #'
 #' @export
 #'
+#' @return No return value. Prints the head and tail as well as the starting and
+#'  end date of a digiRhythm friendly dataframe.
 
 df_act_info <- function(df){
 
@@ -32,6 +34,9 @@ df_act_info <- function(df){
 #' @importFrom stats IQR quantile
 #'
 #' @export
+#'
+#' @return return a dataframe where columns start the second one have undergone
+#' an outlier removal.
 
 
 remove_activity_outliers <- function(df){
@@ -57,12 +62,10 @@ remove_activity_outliers <- function(df){
 #' @param verbose if TRUE, print the string
 #'
 #' @importFrom crayon red green
-#'
-#' @export
 
 print_v <- function(
-  string,
-  verbose){
+    string,
+    verbose){
 
   if(verbose) {
     cat(string, '\n')
@@ -81,6 +84,8 @@ print_v <- function(
 #' df <- df516b_2
 #' dgm_periodicity(df)
 #' @export
+#'
+#' @return returns a periodicity object of type xts.
 
 dgm_periodicity <- function(data){
 
@@ -100,8 +105,7 @@ dgm_periodicity <- function(data){
 #' @param Z the power of the frequency
 #' @param fmax the maximum frequency in the spectrum
 #' @param tm the time grid of the original time series
-#'
-#' @export
+
 
 #Copied from the LOMB library
 pbaluev <- function(Z, fmax, tm) {
@@ -126,8 +130,6 @@ pbaluev <- function(Z, fmax, tm) {
 #' @param fmax the maximum frequency in the spectrum
 #' @param tm the time grid of the original time series
 #' @param alpha the significance level
-#'
-#' @export
 
 
 levopt <-  function(Z,alpha,fmax,tm){

### 2022-11-07 R/dfc.R
Changed the graphics fonts and color in all functions (size is now 15 and color is black)

@@ -78,25 +78,15 @@ dfc <- function(
 
 
   df$date <- date(df$datetime)
-  ##Change (took into account all days including missing days)
-  # days <- unique(df$date)
+
   days <- seq(from = df$date[1],
               to = last(df$date),
               by = 1)
 
-
-
-  ##Change (removed)
-  # if (length(which(diff(days) != 1)) > 0) {
-  #   warning('There is an interruption in the days sequence, i.e., there are non consecutive
-  #         days in the data')
-  #   print('Interruption is at the following days:')
-  #   cat(which(diff(days) != 1), '\n')
-  # }
-
   dfc <- NULL
   spec <- NULL
-  dfc <- data.frame(date = character(),
+  dfc <- data.frame(from = character(),
+                    to = character(),
                     dfc = numeric(),
                     hp = numeric()) #The data frame for DFC
 
@@ -113,7 +103,7 @@ dfc <- function(
   for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
 
     if (verbose) {
-      cat("Processing dates ", as.character(days[i]), " until ", as.character(days[(i + 6)]), "\n")
+      cat("Processing dates ", format(days[i]), " until ", format(days[(i + 6)]), "\n")
 
     }
 
@@ -122,16 +112,11 @@ dfc <- function(
     #Filtering the next seven days by date (not by index - in case of missing data, filtering by index would make errors)
     data_week <- df %>% filter(date >= days[i]) %>%  filter(date <= days[i + 6])
 
-    #data_week <- df %>% filter(date >= days[i]) %>%  filter(date <= (days[i]+6))
-
 
     #Selecting the first column (datetime) and the activity column
     df_var <- data_week %>% select(1, `activity`)
 
-    lsp <- lomb_scargle_periodogram(df_var, alpha = sig, sampling = sampling, plot = plot_lsp)
-
-    # lomb::lsp(df_var, alpha =  sig, plot = TRUE)
-
+    lsp <- lomb_scargle_periodogram(df_var, alpha = sig, plot = TRUE)
 
     #Computing the p-values for each frequency
     # From timbre: seems they did not take the case where p>0.01 into account
@@ -155,19 +140,9 @@ dfc <- function(
 
 
     sumall <- sum(lsp_data$power) #sum of all powers
-
-    #
[…]

### 2022-11-07 R/diurnality.R

@@ -149,12 +149,11 @@ diurnality <- function(data,
     ylab("Diurnality Index") +
     xlab("Date") +
     theme(
-      axis.text.x = element_text(color = "#000000"),
-      axis.text.y = element_text(color = "#000000"),
-      text = element_text(size = 12),
+      axis.text = element_text(color = "#000000"),
+      text = element_text(size = 15),
       panel.background = element_rect(fill = "white"),
       axis.line = element_line(size = 0.5),
-    )
+    ) + geom_hline(yintercept= 0, linetype='dotted', col = 'black')
 
   if (!is.null(save)) {
 

### 2022-11-07 original_codes/original_daily_acticity_wrap_plot.R

@@ -45,6 +45,7 @@ avg_plot <- ggplot(data_to_plot, aes(
   ggtitle(paste(activity_alias, "(", data_to_plot$date[1], "-", last(data_to_plot$date),")")) +
   xlab("Time") +
   theme(
+    axis.text = element_text(color = 'black'),
     panel.background = element_rect(fill = "white"),
     legend.key = element_rect(fill = "white"),
     legend.key.width= unit(0.5, "cm"),

### 2022-11-07 original_codes/original_daily_average_activity.R

@@ -62,9 +62,8 @@ avg_act_plot <- ggplot(s,
   theme_classic() +
   theme(plot.title = element_text(hjust = 0.5)) +
   theme(
-    axis.text.x = element_text(color="#000000"),
-    axis.text.y = element_text(color="#000000"),
-    text=element_text(size = 12),
+    axis.text = element_text(color = 'black'),
+    text=element_text(size = 15),
     panel.background = element_rect(fill = "white"),
     axis.line = element_line(size = 0.5))
 

### 2022-11-07 original_codes/original_dfc.R

@@ -100,14 +100,6 @@ if (length(days) < 7) {
   stop('You need at least 7 days of data to run the Degree of Functional Coupling algorithm')
 }
 
-##Change (removed)
-# if (length(which(diff(days) != 1)) > 0) {
-#   warning('There is an interruption in the days sequence, i.e., there are non consecutive
-#         days in the data')
-#   print('Interruption is at the following days:')
-#   cat(which(diff(days) != 1), '\n')
-# }
-
 dfc <- NULL
 spec <- NULL
 dfc <- data.frame(from = character(),
@@ -138,10 +130,6 @@ for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
   data_week <- df %>% filter(date >= days[i]) %>%  filter(date <= days[i + 6])
 
 
-  # print(paste('Length Unique days', length(unique(as.Date(data_week[,1])))))
-  #data_week <- df %>% filter(date >= days[i]) %>%  filter(date <= (days[i]+6))
-
-
   #Selecting the first column (datetime) and the activity column
   df_var <- data_week %>% select(1, `activity`)
 
@@ -169,19 +157,9 @@ for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
 
 
   sumall <- sum(lsp_data$power) #sum of all powers
-
-  #
-  # new_lsp_data <- lsp_data %>% filter(
-  #   power >= lsp$sig.level | p_values <= sig | harm_power == TRUE
-  # )
-
   ssh <- sum(lsp_data$power[lsp_data$power >= lsp$sig.level
                             & lsp_data$status_harmonic == TRUE])
-
   sumsig <- sum(lsp_data$power[which(lsp_data$power >= lsp$sig.level)])  #sum of all significant
-
-  # sumsig <- sum(lsp_data$power[which(lsp_data$p_values <= sig)])
-
 
   # frequencies (each one has a power)
   # sumall: sum of powers for all frequencies (96) ==> 100: ALL
@@ -209,20 +187,22 @@ for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
   }
 }
 
-# dfc$date <- date(dfc$date, format("%Y-%m-%d"))
-dfc$date <- as.Date(dfc$date)
+dfc$from <- as.Date(dfc$from, format("%Y-%m-%d"))
+dfc$to <- as.Date(dfc$to, format("%Y-%m-%d"))
 dfc$dfc <- as.numeric(dfc$dfc)
[…]

### 2022-11-07 original_codes/original_lomb_scargle_periodogram.R

@@ -1,37 +1,10 @@
-#' Computes the Lomb Scargle Periodogram and returns the information needed for
-#' computing the DFC and HP. A plot visualizing the Harmonic Frequencies presence
-#' in the spectrum is possible. The function is inspired from the Lomb library in
-#' a great part, with modifications to fit the requirements of harmonic powers and
-#' computation of the DFC. This function is inspired by the lsp function from the
-#' lomb package and adapted to add different colors for harmonic and non harmonic
-#' frequencies in the signal. For more information about lomb::lsp, please refer
-#' to: https://cran.r-project.org/web/packages/lomb/
-#'
-#' @param data a digiRhythm friendly dataframe of only two columns
-#' @param sampling the sampling period in minutes. default = 15 min.
-#' @param alpha the statistical significance for the false alarm
-#' @param plot if TRUE, the LSP will be plotted
-#' @param extra_info_plot if True, extra information will be shown on the plot
-#'
-#' @return a list that contains a dataframe (detailed below), the significance
-#' level and alpha (for the record). The dataframe contains the power the frequency,
-#' the frequency in HZ, the p values according to Baluev 2008, the period that corresponds
-#' to the frequency in seconds and in hours and finally, a boolean to tell whether
-#' the frequency is harmonic or not.
-#'
-#' @import ggplot2
-#' @importFrom pracma fibsearch
-#'
-#' @export
-#'
-#' @examples
-#' data("df516b_2", package = "digiRhythm")
-#' data <- df516b_2[1:672, c(1,2)]
-#' sig <- 0.01
-#' lomb_scargle_periodogram(data, alpha = sig, plot = TRUE)
-
-
-lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = TRUE, extra_info_plot = TRUE) {
+library(pracma)
+data("df516b_2", package = "digiRhythm")
+data <- df516b_2[1:672, c(1,2)]
+alpha = 0.01
+sampling = 15
+plot = TRUE
+extra_info_plot = TRUE
 
   if (!is_dgm_friendly(data, verbose = TRUE)) {
     stop('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
@@ -197,7 +170,9 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = 
       p <- p + ggtitle(paste('LSP for ', datanames[2],from_to))
     }
     print(p)
-    }
+  }
 
-  return(output)
-}
+
+
+
+

### 2022-11-07 R/dfc.R
Updated graphs inside DFC

@@ -136,12 +136,12 @@ dfc <- function(
 
 
     lsp_data <- lsp$lsp_data[1:len,]
-    harm_power <- lsp_data$power[lsp_data$status_harmonic == TRUE] #The harmonic powers
+    harm_power <- lsp_data$power[lsp_data$status_harmonic == 'Harmonic'] #The harmonic powers
 
 
     sumall <- sum(lsp_data$power) #sum of all powers
     ssh <- sum(lsp_data$power[lsp_data$power >= lsp$sig.level
-                              & lsp_data$status_harmonic == TRUE])
+                              & lsp_data$status_harmonic == 'Harmonic'])
     sumsig <- sum(lsp_data$power[which(lsp_data$power >= lsp$sig.level)])  #sum of all significant
 
     # frequencies (each one has a power)

### 2022-11-07 R/lomb_scargle_periodogram.R

@@ -176,8 +176,9 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = 
       #          x = max(lsp_data$frequency_hz),
       #          y = level*1.05,
       #          label = paste("P<", alpha), size = 6, vjust = 0) +
-      labs(fill = 'Status', y = 'Power', x = 'Frequency (Hz)') +
-      theme(
+      labs(fill = 'Status', y = 'Power', x = 'Frequency (Hz)')
+    if(extra_info_plot){
+      p <- p + theme(
         panel.background = element_rect(fill = "white"),
         axis.text = element_text(color = "#000000"),
         text = element_text(size = 15),
@@ -187,15 +188,28 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = 
         legend.justification ="right",
         legend.position = c(1,0.89),
         plot.margin = margin(t = 50)) +
-      geom_text(data = hdata, mapping = aes(
-        x = frequency_hz,
-        y = power,
-        label = new_h,
-        angle = 90,
-        hjust = -0.4))
-    if(extra_info_plot){
-      p <- p + ggtitle(paste('LSP for ', datanames[2],from_to))
+        geom_text(data = hdata, mapping = aes(
+          x = frequency_hz,
+          y = power,
+          label = new_h,
+          angle = 90,
+          hjust = -0.4)) + ggtitle(paste('LSP for ', datanames[2],from_to))
+    } else {
+      p <- p + theme(
+        panel.background = element_rect(fill = "white"),
+        axis.text = element_text(color = "#000000"),
+        text = element_text(size = 15),
+        axis.line = element_line(size = 0.5),
+        legend.position = 'none',
+        plot.margin = margin(t = 50)) +
+        geom_text(data = hdata, mapping = aes(
+          x = frequency_hz,
+          y = power,
+          label = new_h,
+          angle = 90,
+          hjust = -0.4))
     }
+
     print(p)
     }
 

### 2022-11-07 original_codes/original_dfc.R

@@ -153,12 +153,12 @@ for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
 
 
   lsp_data <- lsp$lsp_data[1:len,]
-  harm_power <- lsp_data$power[lsp_data$status_harmonic == TRUE] #The harmonic powers
+  harm_power <- lsp_data$power[lsp_data$status_harmonic == 'Harmonic'] #The harmonic powers
 
 
   sumall <- sum(lsp_data$power) #sum of all powers
   ssh <- sum(lsp_data$power[lsp_data$power >= lsp$sig.level
-                            & lsp_data$status_harmonic == TRUE])
+                            & lsp_data$status_harmonic == 'Harmonic'])
   sumsig <- sum(lsp_data$power[which(lsp_data$power >= lsp$sig.level)])  #sum of all significant
 
   # frequencies (each one has a power)
@@ -222,7 +222,6 @@ if(plot_harmonic_part){
     theme(
       axis.text.x = element_text(size=rel(1.5), color = 'black'),
       axis.text.y = element_text(size=rel(1.5), color = 'black'),
-      panel.background = element_rect(fill = "white"),
       axis.line = element_line(size = 0.5),
       legend.key = element_rect(fill = "white"),
       legend.key.width = unit(0.5, "cm"),

### 2022-11-07 original_codes/original_lomb_scargle_periodogram.R

@@ -149,8 +149,9 @@ extra_info_plot = TRUE
       #          x = max(lsp_data$frequency_hz),
       #          y = level*1.05,
       #          label = paste("P<", alpha), size = 6, vjust = 0) +
-      labs(fill = 'Status', y = 'Power', x = 'Frequency (Hz)') +
-      theme(
+      labs(fill = 'Status', y = 'Power', x = 'Frequency (Hz)')
+    if(extra_info_plot){
+      p <- p + theme(
         panel.background = element_rect(fill = "white"),
         axis.text = element_text(color = "#000000"),
         text = element_text(size = 15),
@@ -160,15 +161,28 @@ extra_info_plot = TRUE
         legend.justification ="right",
         legend.position = c(1,0.89),
         plot.margin = margin(t = 50)) +
-      geom_text(data = hdata, mapping = aes(
-        x = frequency_hz,
-        y = power,
-        label = new_h,
-        angle = 90,
-        hjust = -0.4))
-    if(extra_info_plot){
-      p <- p + ggtitle(paste('LSP for ', datanames[2],from_to))
+        geom_text(data = hdata, mapping = aes(
+          x = frequency_hz,
+          y = power,
+          label = new_h,
+          angle = 90,
+          hjust = -0.4)) + ggtitle(paste('LSP for ', datanames[2],from_to))
+    } else {
+      p <- p + theme(
+        panel.background = element_rect(fill = "white"),
+        axis.text = element_text(color = "#000000"),
+        text = element_text(size = 15),
+        axis.line = element_line(size = 0.5),
+        legend.position = 'none',
+        plot.margin = margin(t = 50)) +
+        geom_text(data = hdata, mapping = aes(
+          x = frequency_hz,
+          y = power,
+          label = new_h,
+          angle = 90,
+          hjust = -0.4))
     }
+
     print(p)
   }
 

### 2022-11-07 original_codes/original_dfc.R
Changed the date of DFC

@@ -202,7 +202,7 @@ for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
     lsp_data$power,
     lsp_data$p_values))
 
-  dfc[i,] <-  c(format(days[i]), format(days[i+6]), DFC, HP)
+  dfc[i+6,] <-  c(format(days[i]), format(days[i+6]), DFC, HP)
 
   if (verbose) {
     print(dfc[i,])

### 2022-11-07 original_codes/original_dfc.R
changed the day of DFC

@@ -183,7 +183,7 @@ for (i in 1:n_days_scanned) {# Loop over the days (7 by 7)
   dfc[i+6,] <-  c(format(days[i]), format(days[i+6]), DFC, HP)
 
   if (verbose) {
-    print(dfc[i,])
+    print(dfc[i+6,])
   }
 }
 

### 2022-11-10 R/daily_average_activity.R
Updated codes after the emergence of new R version

@@ -71,7 +71,8 @@ daily_average_activity <- function(
   s <- sum_of_activity_over_all_days_per_sample
 
   s$datetime <- paste(data_to_plot$date[1], s$time)
-  s$datetime <- as.POSIXct(s$datetime, format("%Y-%m-%d %H:%M"))
+  s$datetime <- as.POSIXct(s$datetime, format("%Y-%m-%d %H:%M"), tz = tz(df$datetime))
+
   s <- s %>% select(datetime, average)
 
   avg_act_plot <- ggplot(s,

### 2022-11-10 R/resample_dgm.R

@@ -22,10 +22,11 @@
 
 resample_dgm <- function(data, new_sampling){
 
-  xts_data <- data
-  rownames(xts_data) <- data[,1]
-  xts_data[1] <- NULL
-  xts_data <- as.xts(xts_data)
+  xts_data <- xts(
+    x = data[,c(2:ncol(df))],
+    order.by = data[,1]
+  )
+
 
   original_sampling <- xts::periodicity(xts_data)$frequency
 

### 2022-11-10 original_codes/original_daily_average_activity.R

@@ -47,7 +47,7 @@ for(t in unique(data_to_plot$time)){
 s <- sum_of_activity_over_all_days_per_sample
 
 s$datetime <- paste(data_to_plot$date[1], s$time)
-s$datetime <- as.POSIXct(s$datetime, format("%Y-%m-%d %H:%M"))
+s$datetime <- as.POSIXct(s$datetime, format("%Y-%m-%d %H:%M"), tz = tz(df$datetime))
 s <- s %>% select(datetime, average)
 
 avg_act_plot <- ggplot(s,

### 2022-11-10 original_codes/original_dfc.R

@@ -7,13 +7,13 @@ library(ggplot2)
 #Arguments configuration
 
 #Dataset without interruption
-# data("df516b_2", package = "digiRhythm")
-# df <- df516b_2
-# activity = names(df)[2]
-# sampling = 15
-# sig <- 0.05
-# plot <- TRUE
-# verbose = TRUE
+data("df516b_2", package = "digiRhythm")
+df <- df516b_2
+activity = names(df)[2]
+sampling = 15
+sig <- 0.05
+plot <- TRUE
+verbose = TRUE
 
 
 ##Dataset with interruption
@@ -51,7 +51,7 @@ library(ggplot2)
 # sampling = 15
 # plot <- TRUE
 # verbose = TRUE
-# plot_harmonic_part = TRUE
+plot_harmonic_part = TRUE
 
 #Example from Marie's dataset
 
@@ -59,29 +59,29 @@ library(ggplot2)
 #We assume that the first column is a datetime column and the other columns are activity columns
 #df should be a dataframe
 
-target_tz <- 'GMT'
-data <- import_raw_activity_data("team/marie/12112.csv",
-                               skipLines = 7,
-                               act.cols.names = c("Date", "Time", "Motion Index", "Steps"),
-                               date_format = "%d.%m.%Y",
-                               time_format = "%H:%M:%S",
-                               sep = ",",
-                               original_tz = "CET",
-                               target_tz = target_tz,
-                               sampling = 15,
-                               trim_first_day = TRUE,
-                               trim_middle_days = TRUE,
-                               trim_last_day = TRUE,
-                               verbose = FALSE)
+# target_tz <- 'GMT'
+# data <- import_raw_activity_data("team/marie/12112.csv",
+#                                skipLines = 7,
+#                                act.cols.names = c("Date", "Time", "Motion Index", "Steps"),
+#                                date_format = "%d.%m.%Y",
+#                                time_format = "%H:%M:%S",
+#                                sep = ",",
+#                                original_tz = "CET",
+#                                target_tz = target_tz,
+#                                sampling = 15,
+#                                trim_first_day = TRUE,
+#                                trim_middle_days = TRUE,
+#                                trim_last_day = TRUE,
+#                                verbose = FALSE)
+#
+# sig = 0.05
+# activity = names(data)[2]
+# sampling = 15
+# plot <- TRUE
[…]

### 2022-11-10 original_codes/original_resample.R

@@ -10,9 +10,10 @@ new_sampling <- 30
 #Ensuring that the dataframe has a dataframe class instead of tibble
 df <- as.data.frame(df)
 xts_data <- df
-rownames(xts_data) <- df[,1]
-xts_data[1] <- NULL
-xts_data <- as.xts(xts_data)
+xts_data <- xts(
+  x = df[,c(2:ncol(df))],
+  order.by = df[,1]
+)
 
 original_sampling <- xts::periodicity(xts_data)$frequency
 

### 2022-11-10 R/daily_average_activity.R
Few more updates amid the new version

@@ -17,7 +17,7 @@
 #' @importFrom magrittr %>%
 #' @importFrom stats time
 #' @import ggplot2
-#' @importFrom lubridate date
+#' @importFrom lubridate date tz
 #' @import dplyr
 #'
 #' @export

### 2022-11-10 R/globals.R

@@ -13,5 +13,7 @@ utils::globalVariables(
     'power',
     'period_hours',
     'new_h',
-    'to')
+    'to',
+    'tz',
+    'df')
   )

### 2022-11-10 R/resample_dgm.R

@@ -23,7 +23,7 @@
 resample_dgm <- function(data, new_sampling){
 
   xts_data <- xts(
-    x = data[,c(2:ncol(df))],
+    x = data[,c(2:ncol(data))],
     order.by = data[,1]
   )
 
@@ -60,4 +60,5 @@ resample_dgm <- function(data, new_sampling){
     coredata(sampled_xts)
   )
 
+  return(new_data)
 }

### 2023-01-05 build_cran_package.sh
Changed the description file

@@ -1,0 +1,32 @@
+#After finish working on the file in the current directory,
+#one needs to move some files to the ../dgm folder in order
+#to execute an R CMD check independently for a CRAN submission
+
+#Moving the files
+rm -rf ../dgm/*
+
+#Data files
+cp -R data ../dgm
+cp -R inst ../dgm
+
+#Manuals
+cp -R man ../dgm
+
+cp -R Meta ../dgm
+cp -R doc ../dgm/inst
+
+#Code
+cp -R R ../dgm
+
+#Vignettes
+cp -R vignettes ../dgm
+
+#Desc, namespace and readme files
+cp -R DESCRIPTION ../dgm
+cp -R NAMESPACE ../dgm
+cp -R README.md ../dgm
+
+cd ..
+
+/opt/local/R/4.1.3/bin/R CMD check --as-cran dgm
+/opt/local/R/4.1.3/bin/R CMD build dgm

### 2023-03-31 R/data.R
comments

@@ -113,7 +113,7 @@
 #' A dataset of start and endtime of the morningmilking and eveningmilking on a dairy farm.
 #'
 #' @format A data frame of 4 columns
-#' \describe {
+#' \describe{
 #'  \item{day_start}{a POSIX formatted datetime}
 #'  \item{day_end}{a POSIX formatted datetime}
 #'  \item{night_start}{a POSIX formatted datetime}

### 2023-03-31 jstatsoft/article.Rnw
Added the jstatsoft paper

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algpseudocode}

%% another package (only for this demo article)
\usepackage{framed}
\usepackage{graphicx}
\graphicspath{ {./figures} }

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}
% \newcommand{\ds}{'df516b_2'}
%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("digiRhythm")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation (and optionally ORCID link)
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
   \And Joanna Stachowicz~\orcidlink{0000-0003-4921-8353}\\Agroscope - Digital Production
   \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institut für \"Okologischen Landbau
   \And Christina Umst\"atter~\orcidlink{0000-0002-7638-3867}\\Johann Heinrich von Th\"unen\\ Institut für Agrartechnologie}
\Plainauthor{Hassan-Roland Nasser, Joanna Stachowicz, Marie Schneider, Christina Umst\"atter}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmic behavior assessment in livestock}
\Plaintitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
\Shorttitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}

%% - \Abstract{} almost as usual
\Abstract{
Animals' welfare is becoming increasingly attractive, especially in livestock farming. One method to quantify an animal's welfare is measuring its rhythmicity from activities acquired with sensors (e.g., accelerometers) mounted on the animal. The Degree of functional coupling (DFC) is one of these rhythmicity indicators. Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks. Despite the importance of this indicator and its use in the literature, no open source library is available for computing it. In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides off-the-shelves functionalities for measuring animal's rhythmicity, and other valuable add-ons such as creating actrograms, computing and visualizing average activities, and the calculation of the diurnality index. We gave special attention to integrating the package in the research workflow, from raw data to creating publication-ready content.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
\Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by a comma).

\Address{
  Hassan-Roland Nasser, Joanna Stachowicz\\
  Agroscope, Digital Production\\
  Tanikon 1\\
  8352, Ettenhaussen \\
  Switzerland\\
  E-mail: \email{roland.nasser@agroscope.admin.ch, joa.stachowi@gmail.com}\\
  URL: \url{https://ch.linkedin.com/in/nasserha/, https://orcid.org/0000-0003-4921-8353}\\
  \newline

  Marie Schneider\\
  Johann Heinrich von Th\"unen Institut für \"Okologischen Landbau\\
  Trenthorst 32\\
  23847 Westerau\\
  Deutschland \\
  \emph{and}\\
  Universit\"at Hohenheim – Institut f\"ur Verfahrenstechnik der Tierhaltungssysteme \\
  Garbenstra{\ss}e 9 \\
  70599 Stuttgart\\
  Deutschland\\
  E-mail: \email{marie.schneider@thuenen.de}\\
  \newline

  Christina Umst\"atter\\
  Johann Heinrich von Th\"unen Institut f\"ur Agrartechnologie \\
  Bundesallee 47 \\
  38116 Braunschweig\\
  Deutschland\\
  E-mail: \email{christina.umstaetter@thuenen.de}\\
  URL: \url{https://www.thuenen.de/de/fachinstitute/agrartechnologie/personal/leitung/prof-dr-christina-umstaetter}\\

}

\begin{document}

\section[Introduction: Degree of functional coupling and animal welfare ]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}

The Degree of functional coupling (DFC) is a metric used to describe the extent to which a measured activity is significantly cyclic to 24 hours and thus harmonically synchronized with the periodicity of the environment. The activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal. The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep and used by \cite{scheibe1999comparative} to zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors. With the increasing availability of sensor systems to collect behavioral data continuously, more studies using the DFC were undertaken \citep{sarout2018assessment, fuchs2022detecting}. Throughout multiple studies (some refs), it has been proven that the DFC can be used as a meaningful indicator to assess an animal's welfare, thus the importance of having an open source package specifically for this type of computation. Therefore, an \proglang{R} package was developed and published with name \pkg{digiRhythm} \citep{dgm}.

Two design considerations were followed while developing the package to avoid errors and make the library's outputs easily integrated with third-party \proglang{R} packages and other analysis workflows. The first consideration concerns the input data, whereas a specific tabular data structure is required to avoid function misuse and error propagation. The second consideration deals with the user-friendliness of the data interoperability, where the functions returns are designed to be easily exploited and integrated with other analysis workflows.

This paper is organized as follows. In section \ref{sec:dfc}, the mathematical details of computing the DFC are explained. A deep dive into the statistical significance of frequency peaks needed for the DFC computation is also highlighted. In section \ref{sec:usage}, a detailed overview on all the \pkg{digiRhythm} package functionalities is provided. Finally, conclusions are drawn in section \ref{sec:summary}.

\section{Understanding the degree of functional coupling}{\label{sec:dfc}}

Computing the DFC involves several steps. First, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activityis computed. The second step investigates the spectrum by looking at specific frequencies (harmonics and non-harmonics, further details in Figure \ref{fig:reallsp}). Afterward, the statistical significance of frequency peaks is computed. Finally, a ratio of the sum of frequency powers is calculated (a detailed breakdown of the calculation is given in Figure \ref{dfc_sum}). It was decided to follow a top down approach and start with the  explaining the the ratio of frequency powers. Afterwards, we deep dive into how these frequency powers and their statistical significance can be computed in details.

According to  \cite{scheibe1999comparative}, the DFC can be computed as follows:
%
\begin{equation} \label{eq:dfc_for}
DFC = \frac{ssh}{sumsig}
\end{equation}
%
Another indicator, called the harmonic power, can also be computed similarly:
%
\begin{equation} \label{eq:hp_for}
HP = \frac{ssh}{sumall}
\end{equation}
%
Where $ssh$ is the sum of powers of frequencies that are harmonic and statistically significant, $sumsig$ is the sum of powers of frequencies that are significant (harmonic and non-harmonic), and $sumall$ is the sum of the power of all frequencies in the spectrum.

Both DFC and HP are used to study the rhythmic behaviors of animals. Both indicators can be computed using the \pkg{digiRhythm} package.

\subsection{The Lomb-Scargle Periodogram (LSP)}{\label{sub:lsp}}
The Lomb-Scargle Periodogram is the first step in computing the DFC and HP. The Lomb–Scargle periodogram (LSP) is widely used within the astronomy community and is well-known for detecting and characterizing periodicity in unevenly sampled time-series \citep{vanderplas2018understanding}.

The activityis in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) using the LSP. In the time domain, we can see the magnitude of the activity at a particular time stamp. In the frequency domain we can see the magnitude (power) of oscillations of specific frequencies present in the activity. For example, a  sine wave of 24 hours (Figure \ref{fig:signalVslsp}, left, first row) period could be represented with a one-peak periodogram (Figure \ref{fig:signalVslsp}, right, first row); this peak represents the frequency of a sine wave of 24 hours period. The same idea could be illustrated for a sine wave of 6 hours period (Figure \ref{fig:signalVslsp}, second row), and so on. Moreover, an activitycomposed of the addition of 24 hours and 6 hours sine waves (Figure \ref{fig:signalVslsp}, left, third row) will result in having two peaks for the periodogram (Figure \ref{fig:signalVslsp}, right, third row). Consequently, a periodogram is a tool that allows us to infer the power of each possible sine wave present in the activity.

\begin{figure}
  \includegraphics[width=1\columnwidth]{signalVslsp}
  \caption{Examplary harmonics (left) and their corresponding LSP (right). The first two rows show pefect sine waves acivity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP. The bottom row shows the LSP (right) of two sine waves (of 24 hours period and 6 hours period) added (left) together.}
  \label{fig:signalVslsp}
\end{figure}

Activity signals acquired from sensors on animals are far from perfect sine waves. Therefore, investigating periodic events inside these signals needs a conversion through LSP. Figure \ref{fig:realsignal} shows an example of a real activity example - the motion index (which is a proprietary metric of the
overall animal activity measured in three dimensions (\cite{icetagiowa})) of a cow. The corresponding LSP is shown in Figure \ref{fig:reallsp}.

\begin{figure}
\centering
<<realsignal, fig=TRUE, width=9>>==
data("df516b_2", package = 'digiRhythm')
df <- df516b_2
#We extract seven days from the dataset only for illustration purposes
sub_df <- df[1:672 , c(1,2)]
plot(sub_df$datetime, sub_df$Motion.Index, type = 'l',
     xlab = 'DateTime', ylab = 'Motion Index')
@
\caption{\label{fig:realsignal} The \emph{Motion Index} activity column in the dataset \fct{df516b\_2}.}
\end{figure}

Mathematically, the LSP is computed as follows:
%
\begin{equation}{\label{eq:lsp}}
P_{LS}(f) = \frac{1}{2}
\Biggl\{
\frac{(\sum_n g_n \cos{(2 \pi nf[t_n - \tau])})^2}{\sum_n g_n \cos^2{(2 \pi nf[t_n - \tau])}}
+
\frac{(\sum_n g_n \sin{(2 \pi nf[t_n - \tau])})^2}{\sum_n g_n \sin^2{(2 \pi nf[t_n - \tau])}}
\Biggl\}
\end{equation}

Where $g_n$ is the $n_{th}$ sample in the activity, $t_n$ the $n_{th}$ time index, $f$ is the frequency (in hertz) and $\tau$, which depends on $f$ and ensures time-shift invariance, is presented by:

\begin{equation}{\label{eq:tau_lsp}}
\tau = \frac{1}{4\pi f} \tan^{-1} \frac{\sum_n \sin{(4\pi ft_n)}}{\sum_n \cos{(4\pi ft_n)}}
\end{equation}

Before computing the LSP, we have to determine the frequency grid on which we will apply the formula \ref{eq:lsp}. The frequency grid goes from 0 to $f_{max}$ which is the maximum frequency whom power can be computed from the activity $g_n$. $f_{max}$ is determined according to the Shanon-Nyquist theorem (\cite{shannon1928certain, shannon1949mathematical}).

Figure \ref{fig:reallsp} represent the Lomb-Scargle periodogram computed from the activitypresent in Figure \ref{fig:realsignal}.

\begin{figure}
\centering
<<reallsp, fig=TRUE, width=9>>==
lsp <- lomb_scargle_periodogram(df[1:672 , c(1,2)],
                         alpha = 0.01, sampling = 15, plot = TRUE)
head(lsp$lsp_data)
@
\caption{The Lomb-Scargle Periodogram of the activity data show in Figure \ref{fig:realsignal}. Red bars show the amplitude of harmonic frequencies (with the corresponding harmonic period), while green bars represent non-harmonic frequencies. The dashed line represents the threshold of significance (frequencies with power less than this threshold are considered non-significant). In this particular example, there are only two frequencies with significant power; one of them - which corresponds to a 12-hours period - is harmonic.}
\label{fig:reallsp}
\end{figure}

\subsection{Harmonic frequencies and the intuition behind}{\label{sub:harmonic}}
Animal species exhibit a circadian rhythm reflected in behavioral changes following a 24 hours cycle. Therefore, rhythmic behavioral events are reflected in the power of harmonic frequencies of the primary cycle (24 hours). These harmonics are the frequencies that correspond to 24 hours period divided by any integer smaller than 24 (24/1, 24/2, …). At this point, it is important to recall the subset of frequency powers previously described. These subsets are illustrated in Figure \ref{dfc_sum} for clarification.

\begin{figure}[t]
  \includegraphics[width=1\columnwidth]{sum_dfc}
  \caption{An illustration shows the frequency peaks found in a Lomb Scargle Periodogram in the context of animals' rhythmicity. The whole frequency space consists of frequencies with significant and non-significant peaks, whose powers sum up to $sumall$. The powers of the frequencies with significant peaks sum up to $sumsig$. A subset of the significant peaks represents harmonic frequencies (24h, 24h/2, 24h/3, \dots) whose powers sum up to $ssh$.}
  \label{dfc_sum}
\end{figure}

The relationship between rhythmicity and frequencies would now be clear for the reader. If the significant frequencies are only the harmonic ones (i.e., ssh = sumsig), then the DFC = 1, and therefore the animal exhibits a highly rhythmic behavior. On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior.

\subsection{Measuring the statistical significance of a frequency peak}{\label{sub:significance}}
After computing the periodogram, measuring the statistical significance of  frequency peaks is the second step in the DFC process. The latter would show a peak per frequency. Determining whether a certain peak reflects a activitycomponent or a noise component is a crucial step. In fact, for a small number of samples and a low signal-to-noise ratio, the noise peaks become comparable to the activity peaks. Hence, the importance of having a tool that measures the statistical significance of a peak.
In this regard, the False Alarm Probability (FAP) is the measure that is typically used. The FAP reflects the probability that data with no activitywould lead to a peak of similar magnitude.

According to \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom. This implies that the probability of observing periodogram values less than a given $Z=P(f_0)$ in an activityformed by only Gaussian noise is:
%
\begin{equation}
  P_{single} = 1 - \exp{Z}
\end{equation}
%
To compute the FAP, there are multiple approaches. According to \citep{suveges2015comparative}, the Baluev method \citep{baluev2008assessing} is one of the top performing ones. This method was chosen to be implemented in \pkg{digiRhythm}.

According to Baluev \citep{baluev2008assessing}, the FAP could be written as:

\begin{equation}
  FAP(z) \approx 1 - [P_{single}(z)]\exp(-\tau(z))
\end{equation}

Where:
\begin{equation}
\tau(z) = W(1-z)^{(N-4)/2}\sqrt{z}
\end{equation}
and $W = f_{max}\sqrt{4\pi var(t)}$

\section{digiRhythm functionalities and usage}\label{sec:usage}
In addition to the ability to compute the DFC and HP, \pkg{digiRhythm} contains much more functionalities. These functionalities were complied with animal behavior scientists to provide the commonly needed tools for a rhythmicity analysis topic. Currently, the library offer the following tools:

\begin{itemize}
  \item \fct{lomb\_scargle\_periodogram}: Computing and plotting the Lomb Scargle Periodogram. This function also offers comprehensive visualization of the LSP.
  \item \fct{dfc}: Computing and plotting the Degree of functional coupling and the harmonic part.
  \item \fct{daily\_average\_activity}: Computing and plotting the daily average activity.
  \item \fct{diurnality}: Computing and plotting the diurnality index.
  \item \fct{actogram}: Actogram Visualization.

\end{itemize}

The \fct{dfc} function is a wrapper of the  that users could also use should they wish to investigate more about
\subsection{The input data format}
As mentioned in section \ref{sec:intro}, the library accepts input Data frame s with a specific structure. A typical data frame sampled at 15 min intervals is shown below. This data frame is provided with the \pkg{digiRhythm} package and could be fetched using:

<<input_data>>==
data("df516b_2", package = 'digiRhythm')
df <- df516b_2
head(df)
str(df)
@

The above snippet shows a sample activity dataframe where columns represent date-time, the motion index and the number of steps measured during a 15 minutes time interval.

To make sure that users are using the right format, a utility routine that can check whether the data is digiRhythm friendly or not was included. This could be done using the function \fct{is\_dgm\_friendly} as follows:

<<is_dgm>>==
is_dgm_friendly(data = df, verbose = TRUE)
@

This function takes a data frame as an argument and checks whether the data is digiRhythm friendly or not by returning a boolean. The data is considered \pkg{digiRhythm} friendly if it fulfills all these conditions:

\begin{itemize}
  \item A data frame is not a NULL object.
  \item The first column follows the POSIX standard.
  \item The data frame contains at least two columns.
  \item All columns - except the first one - are numeric.
\end{itemize}

The verbose argument helps understand why (or why not) the data is considered (or not) to be digiRhythm friendly. Verbose = TRUE will output information about each condition checked.
A warning message will appear if the data contains less than seven days because at least seven days are needed to compute the DFC and HP, yet other functionalities could be performed.

\subsection[The dfc function]{The \fct{dfc} function}{\label{sub:dgm_fct}}
The DFC and HP are both implemented (\ref{alg:dfc}) in the \fct{dfc} function. The function takes a \pkg{digiRhythm} friendly data frame, a start, an end date, and the variable for which the user wants to run the analysis. It then applies a Lomb Scargle Periodogram over a sliding window of seven days at a time. For instance, the DFC and HP are computed for the days one to seven, then for the days two to eight, and so on. The last six days of the data set would not be considered because it is not recommended to use the LSP (in the context of computing the DFC) for less than seven days. Figure \ref{fig:dfc} shows a typical DFC/HP output.

\begin{algorithm}
\caption{The Degree of Functional Coupling algorithm}\label{alg:dfc}
\begin{algorithmic}
\Require {data}
\Ensure {data is compatible with the library}
\State $days \gets Total\ Number\ of\ Unique\ Days$
\State $output \gets NULL$
\For{\texttt{day in 7:days}}
        \State \texttt{Select data from the day ($day-7$) to day ($day$)}
        \State \texttt{Compute the Lomb Scargle Periodogram of the selected data}
        \State \texttt{Compute the sum of power for all frequencies $sumall$}
        \State \texttt{Select Harmonic frequencies and compute $ssh$}
        \State \texttt{Select Harmonic frequencies and compute $sumsig$}
        \State {dfc $\gets$ ssh/sumsig}
        \State {hp $\gets$ ssh/sumall}
        \State {Output[day] $\gets$ \{day, dfc, hp\}} \Comment{Append the data frame  with a new line}
      \EndFor
\end{algorithmic}
\end{algorithm}

\subsection[Using digiRhythm to compute DFC and HP]{Using \pkg{digiRhythm} to compute DFC and HP}{\label{sub:dgm_dfc_hp}}

The DFC and HP can be computed using the \fct{dfc} function. A recommended workflow is shown below. After reading the data and verfying its compatibility with the library, the data (originally sampled with 1 min) is resampled to 15 minutes. In this context, resamplings mean simple summation. For instance, if the dataset contains the number of steps with 1 minute sampling period, resampling to 15 minutes means adding the steps during each 15 minutes interval. Further, an activity is chosen (in this example, we choose the activity from the second column (motion index)), then the DFC is computed using the \fct{dfc} function as shown below:

<<dfc>>==
<<dfc_snippet, results=hide>>=
df_resampled <- resample_dgm(df, 15)
activity = names(df_resampled)[2]
my_dfc <- dfc(df_resampled, activity = activity,
              plot = TRUE, verbose = FALSE)
@

The returned data frame  contains the DFC and HP measured along a sliding window of seven days as follows:

<<print_dfc>>==
head(my_dfc$data)
print(class(my_dfc))
@

The returned variable \code{my_dfc} contains both the data and the plot as it is a \code{ggplot2} object.

\begin{figure}[t!]
\centering
<<print_dfc, fig=TRUE, cache = TRUE, width=9>>==
library("ggplot2")
plot(my_dfc)
@
\caption{\label{fig:dfc} The DFC and HP of the dataset \fct{df516b\_2}.}
\end{figure}

It is possible to access the computed data and how the user can change the plot based on the returned plot.
Calling the data field from the variable \code{my_dfc} will allow accessing the computed DFC and HP as shown below:

It is, therefore, easy to access the resulting DFC and HP and therefore use them in further analysis like statistical modeling or more customized workflows. Because the returned object has the \pkg{ggplot2} type \citep{wickham2010layered}, it is also possible to change the plot directly by adding a layer on the aesthetics of the \code{my_dfc} object as shown below:

\begin{figure}[h!]
\centering
<<dfc_plot_modified, fig=TRUE, width=9>>==
new_plot <- my_dfc +
   theme(
     text = element_text(family = "Times", size = 8),
     axis.text.y = element_text(size = 15, colour = "red"))
print(new_plot)
@
%
\caption{\label{fig:dfc_modified} The output plot from Figure \ref{fig:dfc} modified by adding a layer of aesthetics causing some modifications to the \fct{ggplot} output returned by the \fct{dfc}. In particular, the font family was changed to \emph{Times}, the font size to 8 points, and the y-axis text was changed to a bigger size (15) and colored red.}
\end{figure}

This is an illustrative yet straightforward example to sketch the feature of resuing the graphical outputs of the library, but the application comes in handy when users need to produce publication-quality pictures where they should abide by a specific format. It is important to mention that this feature is possible because the functions were designed in a way they return \fct{ggplot} objects rather than base \proglang{R} objects.

\subsection[Recommendations to use the dfc function]{Recommendations to use the dfc function}{\label{sub:recom_dfc}}

Although the \fct{dfc} function can accept data with any sampling frequency, it is suggested by several authors \citep{sinz1976systemanalyse, sarout2018assessment, fuchs2022detecting, scheibe1999comparative} that optimal insights could be obtained using seven days sliding window with 15 minutes sampling period. For the sake of simplicity, an option for changing the sliding window length (for example, 6 or 8, or more days) was not implemented. However, this could be done in future versions if proven useful.

\subsection[Actogram visualization]{Actogram visualization}{\label{sub:acto}}

The actogram is often the first-to-go tool when working with animal activity. It is a tiled 2D representation where the x-axis represents the time of the day, the y-axis represents the number of days (or date), and the tiles are color-coded according to the activity intensity at each sample.

Figure \ref{fig:actogram} shows a code snippet for visualizing the actogram of the dataset \fct{df516b\_2} and its corresponding actogram.

\begin{figure}[h!]
\centering
<<actogram, fig=TRUE, width=9>>==
start <- "2020-05-01" #year-month-day
end <- "2020-06-14" #year-month-day -->
activity_alias <- 'Motion Index'
actogram(df, activity, activity_alias , start, end, save = NULL)
@
%
\caption{\label{fig:actogram} The actogram of the of the dataset \fct{df516b\_2}.}
\end{figure}

\subsection[Average activity]{Average activity}{\label{sub:avg}}

The average activity is an indicator of how active an animal is on average during times of the day. The average activity is computed using the following formula:

\begin{equation}
  activity_{average}(T_i) = \frac{1}{d} \sum_d {A(T_{id})}
\end{equation}

Where $T_i$ is a specific time of the day, $d$ is the number of days, and $A(T_{id})$ is the activity value at the time $T_i$ during the day $d$.

The function \fct{daily\_average\_activity} compute and plot this indicator as shown in the Figure \ref{fig:avg_act}.

\begin{figure}[h!]
\centering
<<avg_activity, fig=TRUE, width=9>>==
daily_average_activity(df, activity, activity_alias,
                                 start, end, save = NULL)
@
\caption{\label{fig:avg_act} The average activity of the dataset \fct{df516b\_2}.}
\end{figure}

\subsection[Diurnality Index (DI)]{Diurnality Index (DI)}{\label{sub:di}}

The Diurnality index is also an important measure when it comes to assessing animals' rhythmicity. This measure indicates whether the animal exercises its activity more during the daytime or the nighttime. Using the Diurnality function, the user has the liberty to choose the day and light time. The user can also omit some periods during the day, such as milking time, where the milking process controls the animal's activity rather than its behavior. The Diurnality index is computed according to \cite{hoogenboom1984seasonal}:

\begin{equation}
  DI = \frac{\frac{C_d}{Td} - \frac{C_n}{Tn}}{\frac{C_d}{Td} + \frac{C_n}{Tn}}
\end{equation}

Where $C_d$ and $C_n$ are the summations of activity during the daytime and nighttime, respectively. $T_d$ and $T_n$ are the numbers of activity samples during the daytime and nighttime. The number of samples is usually measured from the sampling time. For instance, if the activity is measured with 15 min time sample, and the daytime is considered to be between 06h00 AM and 04h00 PM, then $T_d$ is considered equal to 4 samples per hour × 10 hours = 40 samples.
The Diurnality function takes the daytime and nighttime as arguments, with default daytime as 06h30 AM till 04h30 PM and default nighttime as 18h00 pm till 05 AM the next day. To change these times, the configuration should follow an HH:MM:SS format as shown below:

<<setting_times>>==
day_time <- c("06:30:00", "16:30:00")
night_time <- c("18:00:00", "05:00:00")
@

Figure \ref{fig:di} shows how to compute and visualize the diurnality index for an example dataset.

\begin{figure}[h!]
\centering
<<diurnality, fig=TRUE>>==
my_di <- diurnality(df, activity, day_time, night_time)
@
%
\caption{\label{fig:di} The Diurnality index of the dataset \fct{df516b\_2}.}
\end{figure}

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Discussion} \label{sec:summary}
The \pkg{digiRhythm} package can make a significant contribution to the animal welfare assessment as it makes the analysis of animal rhythmicity more accessible. The tools is easy to uses and comes in handy especially it is delivered as an \proglang{R} package. The theoretical details and intuition behind the DFC were explained in section \ref{sec:dfc}. The usage of the package was explained in the section \ref{sec:usage}. Further usage examples could be found in the library's vignettes.

At the date of writing this paper, the current version (v1.1) of the library does not offer the possibility to compute rhythmicity with other techniques such as cosinor analysis (\cite{refinetti2016}). Therefore, there is potential to expand the library's functionality in the future to include other techniques.

Furthermore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various sources or techniques. In our study, we obtained rhythmicity measures from sensors on animals. It would be valuable to examine the correlation between the rhythmicity of these measures and those obtained from other sources within the same animals, such as blood samples. For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons. Additionally, \cite{plaza2022gps} used GPS data to estimate rhythmicity. Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.

Finally, We invite other researchers to contribute to this library by submitting data, code, and techniques that can be shared with the scientific community. By collaborating and sharing these techniques, we can collectively advance our understanding of rhythmicity and its role in various biological systems.

\bibliography{refs}
\newpage

\end{document}

### 2023-03-31 jstatsoft/runs.R

@@ -1,0 +1,9 @@
+#For runs
+library(devtools)
+library("tools")
+
+#install_github('https://github.com/nasserdr/digiRhythm')
+
+setwd("~/projects/digiRhythm/jstatsoft")
+Sweave('article.Rnw')
+texi2pdf("article.tex")

### 2023-04-28 jstatsoft/article.Rnw
Added version Marianne

@@
 
-%% - \Abstract{} almost as usual
+%% - \Abstract{} almost as usual # comr: changed abstract throughout.
 \Abstract{
-Animals' welfare is becoming increasingly attractive, especially in livestock farming.
-One method to quantify an animal's welfare is measuring its rhythmicity from activities acquired with sensors (e.g., accelerometers) mounted on the animal.
-The Degree of functional coupling (DFC) is one of these rhythmicity indicators.
+The welfare of farm animals is a growing concern to the public, yet indicators that allow to reliably quantify the welfare of these animals are scarce.
+One method that has been proposed to serve this purpose is measuring the organisms rhythmicity from activities acquired with sensors (e.g., accelerometers).
+The Degree of Functional Coupling (DFC) is one of these rhythmicity indicators.
 Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks.
-Despite the importance of this indicator and its use in the literature, no open source library is available for computing it.
-In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides off-the-shelves functionalities for measuring animal's rhythmicity, and other valuable add-ons such as creating actrograms, computing and visualizing average activities, and the calculation of the diurnality index.
-We gave special attention to integrating the package in the research workflow, from raw data to creating publication-ready content.
+In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from activity data.
+Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of a diurnality index are also integrated in this package.
+The package is customized for integration in the research workflow, from raw data to creating publication-ready content and therefore allows the simple application of these sophisticated methods.
 }
@@
 %% - Should be comma-separated and in sentence case.
-\Keywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
+\Keywords{Lomb-Scargle periodogram, Degree of Functional Coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
 \Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
@@
 
-  Christina Umst\"atter\\
-  Johann Heinrich von Th\"unen Institut f\"ur Agrartechnologie \\
-  Bundesallee 47 \\
-  38116 Braunschweig\\
-  Deutschland\\
-  E-mail: \email{christina.umstaetter@thuenen.de}\\
-  URL: \url{https://www.thuenen.de/de/fachinstitute/agrartechnologie/personal/leitung/prof-dr-christina-umstaetter}\\
+ Marianne Cockburn\\
+  Agroscope, Equines \\
+  Swiss National Stud \\
+  Les Longs-Prés\\
+  1580, Avenches \\
+  Switzerland\\
+  E-mail: \email{marianne.cockburn@agroscope.admin.ch}\\
+  URL: \url{}\\
+  \newline
 
@@
 
-\section[Introduction: Degree of functional coupling and animal welfare ]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}
+\section[Introduction: The Degree of Functional Coupling as a welfare indicator]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}
 
@@
 
-The activityis in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) using the LSP.
+The activity is in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) using the LSP.
 In the time domain, we can see the magnitude of the activity at a particular time stamp.

### 2023-04-28 jstatsoft/article_comr.Rnw
Added version Marianne

--- 2023-04-28 jstatsoft/article.Rnw
@@
 \author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
-   \And Joanna Stachowicz~\orcidlink{0000-0003-4921-8353}\\Agroscope - Digital Production
    \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institut für \"Okologischen Landbau
-   \And Christina Umst\"atter~\orcidlink{0000-0002-7638-3867}\\Johann Heinrich von Th\"unen\\ Institut für Agrartechnologie}
-\Plainauthor{Hassan-Roland Nasser, Joanna Stachowicz, Marie Schneider, Christina Umst\"atter}
+   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope}
+\Plainauthor{Hassan-Roland Nasser, Marie Schneider, Marianne Cockburn}
 
@@
 
-\section[Introduction: The Degree of Functional Coupling as a welfare indicator]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}
+\section[Introduction: The Degree of Functional Coupling as a welfare indicator]{Introduction: The Degree of Functional Coupling as a welfare indicator} \label{sec:intro}
 
-The Degree of functional coupling (DFC) is a metric used to describe the extent to which a measured activity is significantly cyclic to 24 hours and thus harmonically synchronized with the periodicity of the environment.
-The activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal.
-The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep and used by \cite{scheibe1999comparative} to zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
-With the increasing availability of sensor systems to collect behavioral data continuously, more studies using the DFC were undertaken \citep{sarout2018assessment, fuchs2022detecting}.
-Throughout multiple studies (some refs), it has been proven that the DFC can be used as a meaningful indicator to assess an animal's welfare, thus the importance of having an open source package specifically for this type of computation.
-Therefore, an \proglang{R} package was developed and published with name \pkg{digiRhythm} \citep{dgm}.
+Digitalisation has reached livestock farming and sensors are now regularly mounted on animals with the intention of improving management through early detection of physiological abnormalities.
+To date, the reliability of sensor based abnormality detection still proves potential for optimization.
+The Degree of functional coupling (DFC) presents a unique approach to activity evaluation.
+The DFC is a metric used to describe the extent to which a measured activity is significantly cyclic to 24 hours and thus harmonically synchronized with the periodicity of the environment.
+Such activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal, but could also be the visiting times of feeding troughs or watering stations.
+The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep and used by \cite{scheibe1999comparative} to evaluate the welfare of zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
+With the increasing availability of sensor systems to collect behavioral data continuously, more studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}.
+These studies were able to demonstrate that the rhythm of animals is linked to their state of physical well being.
+However, the application of the DFC was always limited to those with extensive knowledge of statistical programming.
+An open source package specifically for this type of computation opens up the opportunities for livestock research to easily apply these algorithms.
+Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}.
 
-Two design considerations were followed while developing the package to avoid errors and make the library's outputs easily integrated with third-party \proglang{R} packages and other analysis workflows.
+% Two design considerations were followed while developing the package to avoid errors and make the library's outputs easily integrated with third-party \proglang{R} packages and other analysis workflows.
 The first consideration concerns the input data, whereas a specific tabular data structure is required to avoid function misuse and error propagation.
@@
 
-This paper is organized as follows.
+This paper describes the functions of the package and is organized as follows.
 In section \ref{sec:dfc}, the mathematical details of computing the DFC are explained.
-A deep dive into the statistical significance of frequency peaks needed for the DFC computation is also highlighted.
-In section \ref{sec:usage}, a detailed overview on all the \pkg{digiRhythm} package functionalities is provided.
-Finally, conclusions are drawn in section \ref{sec:summary}.
+Further, the statistical significance of frequency peaks needed for the DFC computation is explained.
+In section \ref{sec:usage}, a detailed overview of the \pkg{digiRhythm} package functions is provided.
+Finally, we indicate how this package can be useful for researchers in section \ref{sec:summary}.
 
@@
 Computing the DFC involves several steps.
-First, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activityis computed.
+First, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed.
 The second step investigates the spectrum by looking at specific frequencies (harmonics and non-harmonics, further details in Figure \ref{fig:reallsp}).
-Afterward, the statistical significance of frequency peaks is computed.
+In a a next step, the statistical significance of frequency peaks is computed.
 Finally, a ratio of the sum of frequency powers is calculated (a detailed breakdown of the calculation is given in Figure \ref{dfc_sum}).
-It was decided to follow a top down approach and start with the  explaining the the ratio of frequency powers.
-Afterwards, we deep dive into how these frequency powers and their statistical significance can be computed in details.
+We started by explaining the the ratio of frequency powers.
+Ultimately, we describe in detail how these frequency powers and their statistical significance can be computed.
 
@@
 
-The activity is in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) using the LSP.
+Activity is in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) by applying the LSP.
 In the time domain, we can see the magnitude of the activity at a particular time stamp.
@@
 The same idea could be illustrated for a sine wave of 6 hours period (Figure \ref{fig:signalVslsp}, second row), and so on.
-Moreover, an activitycomposed of the addition of 24 hours and 6 hours sine waves (Figure \ref{fig:signalVslsp}, left, third row) will result in having two peaks for the periodogram (Figure \ref{fig:signalVslsp}, right, third row).
+Moreover, an activity composed of the addition of 24 hours and 6 hour sine waves (Figure \ref{fig:signalVslsp}, left, third row) will result in having two peaks for the periodogram (Figure \ref{fig:signalVslsp}, right, third row).
 Consequently, a periodogram is a tool that allows us to infer the power of each possible sine wave present in the activity.
@@
   \includegraphics[width=1\columnwidth]{signalVslsp}
-  \caption{Examplary harmonics (left) and their corresponding LSP (right).
-The first two rows show pefect sine waves acivity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP.
+  \caption{Examplary harmonics (left) and their corresponding Lomb-Scargle Periodogram (LSP) (right).
+The first two rows show regular sine waves activity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP.
 The bottom row shows the LSP (right) of two sine waves (of 24 hours period and 6 hours period) added (left) together.}
@@
 
-Activity signals acquired from sensors on animals are far from perfect sine waves.
-Therefore, investigating periodic events inside these signals needs a conversion through LSP.
-Figure \ref{fig:realsignal} shows an example of a real activity example - the motion index (which is a proprietary metric of the
+Activity signals acquired from sensors on animals are irregular.
+Therefore, investigating periodic events inside these signals requires a conversion through LSP.
+Figure \ref{fig:realsignal} shows an example of activity data - the motion index (which is a proprietary metric of the
 overall animal activity measured in three dimensions (\cite{icetagiowa})) of a cow. The corresponding LSP is shown in Figure \ref{fig:reallsp}.
@@
 Before computing the LSP, we have to determine the frequency grid on which we will apply the formula \ref{eq:lsp}.
-The frequency grid goes from 0 to $f_{max}$ which is the maximum frequency whom power can be computed from the activity $g_n$.
+The frequency grid goes from 0 to $f_{max}$ which is the maximum frequency power that be computed from the activity $g_n$.
 $f_{max}$ is determined according to the Shanon-Nyquist theorem (\cite{shannon1928certain, shannon1949mathematical}).
 
-Figure \ref{fig:reallsp} represent the Lomb-Scargle periodogram computed from the activitypresent in Figure \ref{fig:realsignal}.
+Figure \ref{fig:reallsp} represent the Lomb-Scargle periodogram computed from the activity present in Figure \ref{fig:realsignal}.
 
@@
 @
-\caption{The Lomb-Scargle Periodogram of the activity data show in Figure \ref{fig:realsignal}.
+\caption{The Lomb-Scargle Periodogram of the activity data shown in Figure \ref{fig:realsignal}.
 Red bars show the amplitude of harmonic frequencies (with the corresponding harmonic period), while green bars represent non-harmonic frequencies.
@@
 Therefore, rhythmic behavioral events are reflected in the power of harmonic frequencies of the primary cycle (24 hours).
-These harmonics are the frequencies that correspond to 24 hours period divided by any integer smaller than 24 (24/1, 24/2, …).
+These harmonics are the frequencies that correspond to 24 hour periods divided by any integer smaller than 24 (24/1, 24/2, …).
 At this point, it is important to recall the subset of frequency powers previously described.
-These subsets are illustrated in Figure \ref{dfc_sum} for clarification.
+These subsets are illustrated in Figure \ref{dfc_sum}.
 
@@
   \includegraphics[width=1\columnwidth]{sum_dfc}
-  \caption{An illustration shows the frequency peaks found in a Lomb Scargle Periodogram in the context of animals' rhythmicity.
+  \caption{Frequency peaks found in a Lomb Scargle Periodogram in the context of animals' rhythmicity.
 The whole frequency space consists of frequencies with significant and non-significant peaks, whose powers sum up to $sumall$.
@@
 If the significant frequencies are only the harmonic ones (i.e., ssh = sumsig), then the DFC = 1, and therefore the animal exhibits a highly rhythmic behavior.
-On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior.
+On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior., which in turn would indicate a poor physiological state.
 
 \subsection{Measuring the statistical significance of a frequency peak}{\label{sub:significance}}
-After computing the periodogram, measuring the statistical significance of  frequency peaks is the second step in the DFC process.
+After computing the periodogram, measuring the statistical significance of  frequency peaks is the second step in the DFC computation.
 The latter would show a peak per frequency.
-Determining whether a certain peak reflects a activitycomponent or a noise component is a crucial step.
+Determining whether a certain peak reflects a valid activity component or a noise component is a crucial step.
 In fact, for a small number of samples and a low signal-to-noise ratio, the noise peaks become comparable to the activity peaks.
-Hence, the importance of having a tool that measures the statistical significance of a peak.
-In this regard, the False Alarm Probability (FAP) is the measure that is typically used. The FAP reflects the probability that data with no activitywould lead to a peak of similar magnitude.
+Hence, it is important to verify the data by applying a tool that measures the statistical significance of a peak.
+In this regard, the False Alarm Probability (FAP) is the measure that is typically used. The FAP reflects the probability that data with no activity would lead to a peak of similar magnitude.
 
 According to \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom.
-This implies that the probability of observing periodogram values less than a given $Z=P(f_0)$ in an activityformed by only Gaussian noise is:
+This implies that the probability of observing periodogram values less than a given $Z=P(f_0)$ in an activity formed by only Gaussian noise is:
 %
@@
 %
-To compute the FAP, there are multiple approaches.
+To compute the FAP, multiple approaches are available.
 According to \citep{suveges2015comparative}, the Baluev method \citep{baluev2008assessing} is one of the top performing ones.
@@
 \section{digiRhythm functionalities and usage}\label{sec:usage}
-In addition to the ability to compute the DFC and HP, \pkg{digiRhythm} contains much more functionalities.
-These functionalities were complied with animal behavior scientists to provide the commonly needed tools for a rhythmicity analysis topic.
-Currently, the library offer the following tools:
+In addition to the ability to compute the DFC and HP, \pkg{digiRhythm} contains much more functions.
+These functions were complied with animal behavior scientists to provide the commonly needed tools for rhythmicity analysis.
+Currently, the library offers the following tools:
 
@@
   \item \fct{lomb\_scargle\_periodogram}: Computing and plotting the Lomb Scargle Periodogram. This function also offers comprehensive visualization of the LSP.
-  \item \fct{dfc}: Computing and plotting the Degree of functional coupling and the harmonic part.
+  \item \fct{dfc}: Computing and plotting the Degree of Functional Coupling and the harmonic part.
   \item \fct{daily\_average\_activity}: Computing and plotting the daily average activity.
@@
 
-The \fct{dfc} function is a wrapper of the  that users could also use should they wish to investigate more about
+The \fct{dfc} function is a wrapper of the !MISSING WORD! that users could also use should they wish to investigate more about!MISSING WORD!
 \subsection{The input data format}
-As mentioned in section \ref{sec:intro}, the library accepts input Data frame s with a specific structure.
+As mentioned in section \ref{sec:intro}, the library accepts input Data frames with a specific structure.
 A typical data frame sampled at 15 min intervals is shown below.
@@
 The verbose argument helps understand why (or why not) the data is considered (or not) to be digiRhythm friendly. Verbose = TRUE will output information about each condition checked.
-A warning message will appear if the data contains less than seven days because at least seven days are needed to compute the DFC and HP, yet other functionalities could be performed.
+A warning message will appear if the data contains less than seven days because at least seven days are required to compute the DFC and HP, yet other functionalities could be performed.
 
@@
 The DFC and HP are both implemented (\ref{alg:dfc}) in the \fct{dfc} function.
-The function takes a \pkg{digiRhythm} friendly data frame, a start, an end date, and the variable for which the user wants to run the analysis.
+The function evaluates a \pkg{digiRhythm} friendly data frame, by analyzing the start, and end date, and the variable for which the user wants to run the analysis.
 It then applies a Lomb Scargle Periodogram over a sliding window of seven days at a time.
@@
 After reading the data and verfying its compatibility with the library, the data (originally sampled with 1 min) is resampled to 15 minutes.
-In this context, resamplings mean simple summation.
+In this context, resampling means simple summation.
 For instance, if the dataset contains the number of steps with 1 minute sampling period, resampling to 15 minutes means adding the steps during each 15 minutes interval.
@@
 \centering
-<<print_dfc, fig=TRUE, cache = TRUE, width=9>>==
+<<plot_dfc, fig=TRUE, cache = TRUE, width=9>>==
 library("ggplot2")
@@
 
-It is possible to access the computed data and how the user can change the plot based on the returned plot.
+It is possible to access the computed data and the user can adjust the plot features.
 Calling the data field from the variable \code{my_dfc} will allow accessing the computed DFC and HP as shown below:
 
-It is, therefore, easy to access the resulting DFC and HP and therefore use them in further analysis like statistical modeling or more customized workflows.
-Because the returned object has the \pkg{ggplot2} type \citep{wickham2010layered}, it is also possible to change the plot directly by adding a layer on the aesthetics of the \code{my_dfc} object as shown below:
+It is, therefore, easy to access the resulting DFC and HP and use them in further analysis like statistical modeling or more customized workflows.
+Because the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, it is also possible to change the plot directly by adding a layer on the aesthetics of the \code{my_dfc} object as shown below:
 
@@
 @
-%
-\caption{\label{fig:dfc_modified} The output plot from Figure \ref{fig:dfc} modified by adding a layer of aesthetics causing some modifications to the \fct{ggplot} output returned by the \fct{dfc}.
+
+\caption{\label{fig:dfc_modified} Output plot from Figure \ref{fig:dfc} modified by adding a layer of aesthetics causing some modifications to the \fct{ggplot} output returned by the \fct{dfc}.
 In particular, the font family was changed to \emph{Times}, the font size to 8 points, and the y-axis text was changed to a bigger size (15) and colored red.}
@@
 
-This is an illustrative yet straightforward example to sketch the feature of resuing the graphical outputs of the library, but the application comes in handy when users need to produce publication-quality pictures where they should abide by a specific format.
+This is an illustrative yet straightforward example to sketch the feature of reusing the graphical outputs of the library, but the application is useful when users need to produce publication-quality pictures in a specific format.
 It is important to mention that this feature is possible because the functions were designed in a way they return \fct{ggplot} objects rather than base \proglang{R} objects.
@@
 
-The function \fct{daily\_average\_activity} compute and plot this indicator as shown in the Figure \ref{fig:avg_act}.
+The function \fct{daily\_average\_activity} computes and plots this indicator as shown in Figure \ref{fig:avg_act}.
 
@@
 
-The Diurnality index is also an important measure when it comes to assessing animals' rhythmicity.
-This measure indicates whether the animal exercises its activity more during the daytime or the nighttime.
-Using the Diurnality function, the user has the liberty to choose the day and light time.
+The Diurnality Index is also an important measure when it comes to assessing animals' rhythmicity.
+This measure indicates whether the animal exercises its activity more during diurnal (daytime) or nocturnal (nighttime) periods.
+Using the Diurnality function, the user can specify day and night time.
 The user can also omit some periods during the day, such as milking time, where the milking process controls the animal's activity rather than its behavior.
@@
 \section{Discussion} \label{sec:summary}
-The \pkg{digiRhythm} package can make a significant contribution to the animal welfare assessment as it makes the analysis of animal rhythmicity more accessible.
-The tools is easy to uses and comes in handy especially it is delivered as an \proglang{R} package.
-The theoretical details and intuition behind the DFC were explained in section \ref{sec:dfc}.
-The usage of the package was explained in the section \ref{sec:usage}.
-Further usage examples could be found in the library's vignettes.
+The \pkg{digiRhythm} package can make a significant contribution to the animal welfare assessment as it makes the analysis of animal rhythmicity more accessible to researchers.
+The tool is easy to use, especially it is delivered as an \proglang{R} package.
+The theoretical details and intuition behind the DFC is explained in section \ref{sec:dfc}.
+The usage of the package is explained in the section \ref{sec:usage}.
+Further usage examples can be found in the library's vignettes.
 

### 2023-04-28 jstatsoft/runs.R

@@ -5,5 +5,8 @@ library("tools")
 #install_github('https://github.com/nasserdr/digiRhythm')
 
 setwd("~/projects/digiRhythm/jstatsoft")
-Sweave('article.Rnw')
-texi2pdf("article.tex")
+Sweave('article_comr.Rnw')
+texi2pdf("article_comr.tex")
+
+# To rerun the pictures:
+# Check: ~/projects/digiRhythm/examples/code_snipets_statsoft_paper.R

### 2023-05-07 jstatsoft/article_comr.Rnw
Marie_little changs

@@
 Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks.
+Additionally, categorizing the significant peaks in frequencies that are harmonically synchronised with the periodicity of the environment.
 In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from activity data.
@@
 If the significant frequencies are only the harmonic ones (i.e., ssh = sumsig), then the DFC = 1, and therefore the animal exhibits a highly rhythmic behavior.
-On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior., which in turn would indicate a poor physiological state.
+On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior, which in turn could be initiated by a poor physiological state or other physiological events.
 

### 2023-05-12 jstatsoft/article_comr.Rnw
Marianne input

--- 2023-04-28 jstatsoft/article_comr.Rnw
@@
    \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institut für \"Okologischen Landbau
-   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope}
+   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group}
 \Plainauthor{Hassan-Roland Nasser, Marie Schneider, Marianne Cockburn}
@@
 \Abstract{
-The welfare of farm animals is a growing concern to the public, yet indicators that allow to reliably quantify the welfare of these animals are scarce.
+Indicators that allow to reliably quantify the welfare of animals are scarce.
 One method that has been proposed to serve this purpose is measuring the organisms rhythmicity from activities acquired with sensors (e.g., accelerometers).
@@
 Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of a diurnality index are also integrated in this package.
-The package is customized for integration in the research workflow, from raw data to creating publication-ready content and therefore allows the simple application of these sophisticated methods.
+The package is customized for integration in the research workflow, from raw data to creating publication-ready content and allows the simple application of these sophisticated methods.
 }
@@
 \Address{
-  Hassan-Roland Nasser, Joanna Stachowicz\\
+  Hassan-Roland Nasser\\
   Agroscope, Digital Production\\
@@
  Marianne Cockburn\\
-  Agroscope, Equines \\
+  Agroscope, Equine Research Group \\
   Swiss National Stud \\
@@
   E-mail: \email{marianne.cockburn@agroscope.admin.ch}\\
-  URL: \url{}\\
+  URL: \url{https://orcid.org/0000-0003-3193-0641}\\
   \newline
@@
 
-\section[Introduction: The Degree of Functional Coupling as a welfare indicator]{Introduction: The Degree of Functional Coupling as a welfare indicator} \label{sec:intro}
+\section[Introduction: The Degree of Functional Coupling in the context of livestock production ]{Introduction: The Degree of Functional Coupling in the context of livestock production} \label{sec:intro}
 
-Digitalisation has reached livestock farming and sensors are now regularly mounted on animals with the intention of improving management through early detection of physiological abnormalities.
+### new start
+The real-time monitoring of animal behaviour is becoming increasingly important in agriculture.
+It can provide new opportunities to enhance the efficiency and sustainability of livestock production and is used to enhance animal welfare (Wemelsfelder et al., 2001).
+Animal behaviour can be altered by stress, pain or disease, but monitoring these changes in welfare is still a challenge, since some behavioural definitions are unclear and not specifically linked to a unique adverse event (Theurer et al., 2013; Levitis et al., 2009).
+Relevant information about animal welfare can be obtained by studying individual differences in the capacity for behavioural adaptation (Cooper et al., 2005).
+Indeed, during the animals life, significant changes can occur due to internal and external stimuli; these alterations arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours (Berger et al., 2003).
+Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of animal behaviour data are regularly recorded, and can be used to calculate parameters describing biological rhythms (Theurer et al., 2013; Scheibe and Gromann, 2006; Berger et al.
+2003).
+Scheibe et al.
+(1999) developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding.
+The DFC provides an estimate of synchronization between an individual and its environment.
+It is calculated to each 24 h period within rolling seven day periods and predicts values between 0% and 100% (Scheibe et al., 1999, Berger et al., 2003).
+An organism with a high DFC is highly synchronized with its environment, while lower DFC's indicate a lower synchronization.
+Breakdowns in the DFC response can be caused by stressors (e.g.
+transport, disease, etc.) that affect the time structure of behaviour (Berger et al., 2003; Scheibe et al., 1999).
+In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i.
+e.
+24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations (Scheibe et al., 1999; Berger et al., 2003).
+Previous studies demonstrated that calves change their behaviour as illness develops (reduced feeding, less social behaviour and increased rest) as an adaptive response to cope with the disease (Owen-Ashley et al., 2006, Dantzer and Kelley, 2007; Borderas et al.
+2008).
+A better understanding of changes in behaviour due to illness or stress of handling is valuable to improve animal welfare and productivity as well as handler safety (Grandin et al., 1989; Kilgore et al., 2005; Borderas et al., 2008; Robért et al., 2011).
+Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).
+### new end
+
+suggest to delete start:
+With the progression of digital technologies livestock production has changed significantly.
+Sensors are now regularly mounted on animals, with the intention of improving management practices through monitoring performance, such as daily milk yields.
+However, existing sensor systems are also used for the early detection of physiological changes.
 To date, the reliability of sensor based abnormality detection still proves potential for optimization.
+While processes on farms, including livestock production, are increasingly automated, it is still not possible to reliably assess the welfare state of a farm animals.
 The Degree of functional coupling (DFC) presents a unique approach to activity evaluation.
@@
 Such activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal, but could also be the visiting times of feeding troughs or watering stations.
-The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep and used by \cite{scheibe1999comparative} to evaluate the welfare of zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
-With the increasing availability of sensor systems to collect behavioral data continuously, more studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}.
-These studies were able to demonstrate that the rhythm of animals is linked to their state of physical well being.
+It is assumed that organisms in a good welfare state have a rhythmicity close to 1, whereas the rhythmicity of the organism drops when its physiological state is compromised.
+The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep.
+Hereby, the authors confirmed that the DFC of sheep dropped before they died.
+\cite{scheibe1999comparative} further used the  DFC to evaluate the welfare of zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
+suggest to delete end:
+
+With the increasing availability of sensor technologies to collect behavioral data continuously, further studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}.
+These studies were able to confirm that the rhythm of animals is linked to their state of physical wellbeing.
 However, the application of the DFC was always limited to those with extensive knowledge of statistical programming.
@@
 \section{Discussion} \label{sec:summary}
-The \pkg{digiRhythm} package can make a significant contribution to the animal welfare assessment as it makes the analysis of animal rhythmicity more accessible to researchers.
-The tool is easy to use, especially it is delivered as an \proglang{R} package.
+The \pkg{digiRhythm} package can make a significant contribution to the assessment of rhythmicity in animals, as it makes the analysis of animal rhythmicity more accessible to researchers.
+The tool is easy to use, especially because it is delivered as an \proglang{R} package.
 The theoretical details and intuition behind the DFC is explained in section \ref{sec:dfc}.
@@
 
-At the date of writing this paper, the current version (v1.1) of the library does not offer the possibility to compute rhythmicity with other techniques such as cosinor analysis (\cite{refinetti2016}).
-Therefore, there is potential to expand the library's functionality in the future to include other techniques.
+The package enables the calculation of rhythmicity in, but not limited to, animals.
+To date, this is purely a computation of activity patterns that have been correlated with physiological welfare factors in previous studies.
+This indicator can be calculated from any type of regularly measured time series data.
+However, we have no evidence how these activity patterns are regulated by routine management practices, such as feeding or milking.
+Such routines could indirectly affect rhythmic activity patterns and blur physiological rhythmicity patterns.
+To date, the package provides the option to exclude known management time spans from the analysis.
+However, in order to increase the understanding of rhythmicity as a welfare indicator future studies should investigate how these management practices affect rhythmicity calculations and evaluate if this is a reasonable welfare indicator for animals depending on external management.
 
-Furthermore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various sources or techniques.
-In our study, we obtained rhythmicity measures from sensors on animals.
-It would be valuable to examine the correlation between the rhythmicity of these measures and those obtained from other sources within the same animals, such as blood samples.
+The required frequency of data collection is of utmost importance in animal-borne sensor technologies.
+Sensors are required to be small, so that they don't disturb the animal, and further expected to run without any sensor based maintenance, such as data reading or charging of the device.
+This in turn means that data collection frequency should be low.
+In the current study an extremely low data collection frequency is suggested (15 minutes) future studies should evaluate via downsampling, how low collection frequency can be set to create reliable rhythmicity calculations.
+Further it would be interesting to evaluate if activity data collected from different sources, such as access of feeding stations or watering troughs, compute similiar rhythmicity outputs in the same organism.
+
+Additionally, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
 For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
@@
 Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
+
+The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans.
+Therefore we highly recommend future research to evaluate the DigiRhythm package in other species.
+Many humans nowadays continuously collect activity data with their smart-watches or the phones.
+Thus, it would be interesting to evaluate the application of the DigiRhythm library in humans, offering an objective indicator of human welfare.
+A previous review highlights the potential of such rhythmicity computations in humans (https://journals.sagepub.com/doi/pdf/10.1177/0748730420940483).
+
+At the date of writing this paper, the current version (v1.1) of the library does not offer the possibility to compute rhythmicity with other techniques such as cosinor analysis (\cite{refinetti2016}).
+Therefore, there is potential to expand the library's functionality in the future to include other techniques.
 

### 2023-05-12 examples/HP_model.R

@@ -1,0 +1,396 @@
+#HP model f?r Paper 1- Verfahrensvergleich
+
+#ben?tigte Packages
+library(lme4)
+library(DHARMa)
+library(data.table)
+library(ggpubr)
+library(emmeans)
+library(flexplot)
+library(MuMIn)
+library(psych)
+
+#####
+#Daten einlesen --> Daten m?ssen alle Versuchstiere und Perioden enthalten
+#####
+data <- read.csv("/home/agsad.admin.ch/f80859433/projects/digiRhythm/examples/20230413_Gesamttabelle_Paper_1.csv")
+
+table(data$cow)
+
+#####
+#Daten ansehen
+#####
+
+hist(data$hp)
+#looks like an exponetial distribution but includes 0, that could be a problem
+
+median(data$hp)
+
+mean(data$hp)
+var(data$hp)
+#Mittelwert ist gr??er, als die Varianz, daher kein negative binomial model
+
+#####
+#Versuch verschiedener Datentransformationen
+#####
+hist(log(data$hp))
+
+data$hp2 <- data$hp
+
+#data$hp2[data$hp2 == 0] <- 0.0166123
+data$hp2[data$hp2 == 0] <- 0.001
+
+hist(log(data$hp2))
+
+library(lme4)
+
+# Check Heterosckedasticity
+# Check the repeated measurement settings
+# (1|row) Check if we need to add the id of each observation in case of repeated measurements
+
+# Ve = varExp(form = ~ day)
+# DFC.lme2e <- lme(
+#   dfc ~ breed * treatment * day,
+#   data = DataDFC.df,
+#   random = ~ barn|name, #Random effects have to be mentionned
+#   weights= Ve, control=ctrl)
+
+model <- glmer(formula = hp ~ 1 + factor(treat) + heat + bedding +
+                 ( 1 | herd/cow/lactnr) +
+               data = data,
+               family = poisson(link = 'log')
+               )
+residuals <- resid(model)
+weights <- 1 / residuals ^ 2 #Other options
+
+model <- glmer(formula = hp ~ 1 + factor(treat) + heat + bedding +
+                 ( 1 | herd/cow/lactnr) +
+                 data = data,
+               weights = weights,
+               family = poisson(link = 'log')
+)
+
+anova(model)
+plot(predict(model), residuals)
+
+
[…]

### 2023-05-12 jstatsoft/article.Rnw
Fixed the conflict but files got merged.

@@
 \author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
-   \And Joanna Stachowicz~\orcidlink{0000-0003-4921-8353}\\Agroscope - Digital Production
    \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institut für \"Okologischen Landbau
-   \And Christina Umst\"atter~\orcidlink{0000-0002-7638-3867}\\Johann Heinrich von Th\"unen\\ Institut für Agrartechnologie}
-\Plainauthor{Hassan-Roland Nasser, Joanna Stachowicz, Marie Schneider, Christina Umst\"atter}
+   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group}
+\Plainauthor{Hassan-Roland Nasser, Marie Schneider, Marianne Cockburn}
 
@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmic behavior assessment in livestock}
-\Plaintitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
-\Shorttitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
+\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmicity assessment using the degree of functional coupling}
+\Plaintitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
+\Shorttitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
 

### 2023-05-12 jstatsoft/runs.R

@@ -1,12 +1,12 @@
 #For runs
-library(devtools)
+library("devtools")
 library("tools")
 
 #install_github('https://github.com/nasserdr/digiRhythm')
 
 setwd("~/projects/digiRhythm/jstatsoft")
-Sweave('article_comr.Rnw')
-texi2pdf("article_comr.tex")
+Sweave('article.Rnw')
+texi2pdf("article.tex")
 
 # To rerun the pictures:
 # Check: ~/projects/digiRhythm/examples/code_snipets_statsoft_paper.R

### 2023-05-15 jstatsoft/runs.R
removed non-used personal versions

@@ -3,7 +3,6 @@ library("devtools")
 library("tools")
 
 #install_github('https://github.com/nasserdr/digiRhythm')
-
 setwd("~/projects/digiRhythm/jstatsoft")
 Sweave('article.Rnw')
 texi2pdf("article.tex")

### 2023-05-15 jstatsoft/article.Rnw
Added the article updated by Marianne.

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmicity assessment using the degree of functional coupling}
-\Plaintitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
-\Shorttitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
+\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmic behavior assessment in livestock}
+\Plaintitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
+\Shorttitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
 
@@
 \Abstract{
-Indicators that allow to reliably quantify the welfare of animals are scarce.
-One method that has been proposed to serve this purpose is measuring the organisms rhythmicity from activities acquired with sensors (e.g., accelerometers).
-The Degree of Functional Coupling (DFC) is one of these rhythmicity indicators.
+Measuring an organisms rhythmicity can provide information on its physiological and psychological state.
+What makes this measure so interesting, is that it can be computed from any activity related time series data (e.g., accelerometers, gps tracks, smart watches (step counts), etc.).
+The Degree of Functional Coupling (DFC) is an established method to calculate rhythmicity.
 Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks.
-In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from activity data.
-Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of a diurnality index are also integrated in this package.
+In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from times series data.
+Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of an diurnality index are also integrated in this package.
 The package is customized for integration in the research workflow, from raw data to creating publication-ready content and allows the simple application of these sophisticated methods.
@@
 
-\section[Introduction: The Degree of Functional Coupling as a welfare indicator]{Introduction: Degree of functional coupling and animal welfare} \label{sec:intro}
+\section[Introduction]{Introduction: The Degree of Functional Coupling in the context of livestock production} \label{sec:intro}
 
-The Degree of functional coupling (DFC) is a metric used to describe the extent to which a measured activity is significantly cyclic to 24 hours and thus harmonically synchronized with the periodicity of the environment.
-The activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal.
-The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep and used by \cite{scheibe1999comparative} to zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
-With the increasing availability of sensor systems to collect behavioral data continuously, more studies using the DFC were undertaken \citep{sarout2018assessment, fuchs2022detecting}.
-Throughout multiple studies (some refs), it has been proven that the DFC can be used as a meaningful indicator to assess an animal's welfare, thus the importance of having an open source package specifically for this type of computation.
-Therefore, an \proglang{R} package was developed and published with name \pkg{digiRhythm} \citep{dgm}.
+%new start
+The real-time monitoring of animal behaviour is becoming increasingly important in agriculture.
+It can provide new opportunities to enhance the efficiency and sustainability of livestock production and is used to enhance animal welfare (Wemelsfelder et al., 2001).
+Animal behaviour can be altered by stress, pain or disease, but monitoring these changes in welfare is still a challenge, since some behavioural definitions are unclear and not specifically linked to a unique adverse event (Theurer et al., 2013; Levitis et al., 2009).
+Relevant information about animal welfare can be obtained by studying individual differences in the capacity for behavioural adaptation (Cooper et al., 2005).
+Indeed, during the animals life, significant changes can occur due to internal and external stimuli; these alterations arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours (Berger et al., 2003).
+Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of animal behaviour data are regularly recorded, and can be used to calculate parameters describing biological rhythms (Theurer et al., 2013; Scheibe and Gromann, 2006; Berger et al.
+2003).
+Scheibe et al.
+(1999) developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding.
+The DFC provides an estimate of synchronization between an individual and its environment.
+It is calculated to each 24 h period within rolling seven day periods and predicts values between 0%
+and 100% (Scheibe et al., 1999, Berger et al., 2003).
+An organism with a high DFC is highly synchronized with its environment, while lower DFC's indicate a lower synchronization.
+Breakdowns in the DFC response can be caused by stressors (e.g.
+transport, disease, etc.) that affect the time structure of behaviour (Berger et al., 2003; Scheibe et al., 1999).
+In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i.
+e.
+24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations (Scheibe et al., 1999; Berger et al., 2003).
+Previous studies demonstrated that calves change their behaviour as illness develops (reduced feeding, less social behaviour and increased rest) as an adaptive response to cope with the disease (Owen-Ashley et al., 2006, Dantzer and Kelley, 2007; Borderas et al.
+2008).
+A better understanding of changes in behaviour due to illness or stress of handling is valuable to improve animal welfare and productivity as well as handler safety (Grandin et al., 1989; Kilgore et al., 2005; Borderas et al., 2008; Robért et al., 2011).
+Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).
+%new end
 
-Two design considerations were followed while developing the package to avoid errors and make the library's outputs easily integrated with third-party \proglang{R} packages and other analysis workflows.
+suggest to delete start:
+With the progression of digital technologies livestock production has changed significantly.
+Sensors are now regularly mounted on animals, with the intention of improving management practices through monitoring performance, such as daily milk yields.
+However, existing sensor systems are also used for the early detection of physiological changes.
+To date, the reliability of sensor based abnormality detection still proves potential for optimization.
+While processes on farms, including livestock production, are increasingly automated, it is still not possible to reliably assess the welfare state of a farm animals.
+The Degree of functional coupling (DFC) presents a unique approach to activity evaluation.
+The DFC is a metric used to describe the extent to which a measured activity is significantly cyclic to 24 hours and thus harmonically synchronized with the periodicity of the environment.
+Such activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal, but could also be the visiting times of feeding troughs or watering stations.
+It is assumed that organisms in a good welfare state have a rhythmicity close to 1, whereas the rhythmicity of the organism drops when its physiological state is compromised.
+The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep.
+Hereby, the authors confirmed that the DFC of sheep dropped before they died.
+\cite{scheibe1999comparative} further used the  DFC to evaluate the welfare of zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
+suggest to delete end:
+
+With the increasing availability of sensor technologies to collect behavioral data continuously, further studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}.
+These studies were able to confirm that the rhythm of animals is linked to their state of physical wellbeing.
+However, the application of the DFC was always limited to those with extensive knowledge of statistical programming.
+An open source package specifically for this type of computation opens up the opportunities for livestock research to easily apply these algorithms.
+Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}.
+
+% Two design considerations were followed while developing the package to avoid errors and make the library's outputs easily integrated with third-party \proglang{R} packages and other analysis workflows.
 The first consideration concerns the input data, whereas a specific tabular data structure is required to avoid function misuse and error propagation.
@@
 \section{Discussion} \label{sec:summary}
-The \pkg{digiRhythm} package can make a significant contribution to the animal welfare assessment as it makes the analysis of animal rhythmicity more accessible.
-The tools is easy to uses and comes in handy especially it is delivered as an \proglang{R} package.
-The theoretical details and intuition behind the DFC were explained in section \ref{sec:dfc}.
-The usage of the package was explained in the section \ref{sec:usage}.
-Further usage examples could be found in the library's vignettes.
+The \pkg{digiRhythm} package can make a significant contribution to the assessment of rhythmicity in animals, as it makes the analysis of animal rhythmicity more accessible to researchers.
+The tool is easy to use, especially because it is delivered as an \proglang{R} package.
+The theoretical details and intuition behind the DFC is explained in section \ref{sec:dfc}.
+The usage of the package is explained in the section \ref{sec:usage}.
+Further usage examples can be found in the library's vignettes.
+
+The package enables the calculation of rhythmicity in, but not limited to, animals.
+To date, this is purely a computation of activity patterns that have been correlated with physiological welfare factors in previous studies.
+This indicator can be calculated from any type of regularly measured time series data.
+However, we have no evidence how these activity patterns are regulated by routine management practices, such as feeding or milking.
+Such routines could indirectly affect rhythmic activity patterns and blur physiological rhythmicity patterns.
+To date, the package provides the option to exclude known management time spans from the analysis.
+However, in order to increase the understanding of rhythmicity as a welfare indicator future studies should investigate how these management practices affect rhythmicity calculations and evaluate if this is a reasonable welfare indicator for animals depending on external management.
+
+The required frequency of data collection is of utmost importance in animal-borne sensor technologies.
+Sensors are required to be small, so that they don't disturb the animal, and further expected to run without any sensor based maintenance, such as data reading or charging of the device.
+This in turn means that data collection frequency should be low.
+In the current study an extremely low data collection frequency is suggested (15 minutes) future studies should evaluate via downsampling, how low collection frequency can be set to create reliable rhythmicity calculations.
+Further it would be interesting to evaluate if activity data collected from different sources, such as access of feeding stations or watering troughs, compute similar rhythmicity outputs in the same organism.
+
+Include 7 days aspect.
+
+Additionally, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
+For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
+Additionally, \cite{plaza2022gps} used GPS data to estimate rhythmicity.
+Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
+
+The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans.
+Therefore we highly recommend future research to evaluate the DigiRhythm package in other species.
+Many humans nowadays continuously collect activity data with their smart-watches or the phones.
+Thus, it would be interesting to evaluate the application of the DigiRhythm library in humans, offering an objective indicator of human welfare.
+A previous review highlights the potential of such rhythmicity computations in humans (https://journals.sagepub.com/doi/pdf/10.1177/0748730420940483).
 
@@
 Therefore, there is potential to expand the library's functionality in the future to include other techniques.
-
-Furthermore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various sources or techniques.
-In our study, we obtained rhythmicity measures from sensors on animals.
-It would be valuable to examine the correlation between the rhythmicity of these measures and those obtained from other sources within the same animals, such as blood samples.
-For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
-Additionally, \cite{plaza2022gps} used GPS data to estimate rhythmicity.
-Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
 

### 2023-05-15 jstatsoft/cleanup.sh

@@ -1,0 +1,1 @@
+rm -rf *.pdf *.tpt *.tex *.out *.bbl *.log *.aux

### 2023-05-15 jstatsoft/article.Rnw
amendements

@@
 
-\section[Introduction]{Introduction: The Degree of Functional Coupling in the context of livestock production} \label{sec:intro}
+\section[Introduction]{Introduction} \label{sec:intro}
 
-%new start
 The real-time monitoring of animal behaviour is becoming increasingly important in agriculture.
@@
 The DFC provides an estimate of synchronization between an individual and its environment.
-It is calculated to each 24 h period within rolling seven day periods and predicts values between 0%
-and 100% (Scheibe et al., 1999, Berger et al., 2003).
+It is calculated to each 24 h period within rolling seven day periods and predicts values between $0\%$
+and 100 $\%$ (Scheibe et al., 1999, Berger et al., 2003).
 An organism with a high DFC is highly synchronized with its environment, while lower DFC's indicate a lower synchronization.
@@
 Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).
-%new end
-
-suggest to delete start:
-With the progression of digital technologies livestock production has changed significantly.
-Sensors are now regularly mounted on animals, with the intention of improving management practices through monitoring performance, such as daily milk yields.
-However, existing sensor systems are also used for the early detection of physiological changes.
-To date, the reliability of sensor based abnormality detection still proves potential for optimization.
-While processes on farms, including livestock production, are increasingly automated, it is still not possible to reliably assess the welfare state of a farm animals.
-The Degree of functional coupling (DFC) presents a unique approach to activity evaluation.
-The DFC is a metric used to describe the extent to which a measured activity is significantly cyclic to 24 hours and thus harmonically synchronized with the periodicity of the environment.
-Such activity can be, e.g., the number of steps during a time interval or an acceleration measured with an accelerometer mounted on the animal, but could also be the visiting times of feeding troughs or watering stations.
-It is assumed that organisms in a good welfare state have a rhythmicity close to 1, whereas the rhythmicity of the organism drops when its physiological state is compromised.
-The DFC was first introduced by \cite{sinz1976systemanalyse} to gain insight into the circadian rhythm of sheep.
-Hereby, the authors confirmed that the DFC of sheep dropped before they died.
-\cite{scheibe1999comparative} further used the  DFC to evaluate the welfare of zoo animals and wildlife by applying a sequence of mathematical operations on continuous signals collected via animal-borne sensors.
-suggest to delete end:
 
@@
 The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans.
-Therefore we highly recommend future research to evaluate the DigiRhythm package in other species.
+Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
 Many humans nowadays continuously collect activity data with their smart-watches or the phones.
-Thus, it would be interesting to evaluate the application of the DigiRhythm library in humans, offering an objective indicator of human welfare.
+Thus, it would be interesting to evaluate the application of the \pkg{digiRhythm} package in humans, offering an objective indicator of human welfare.
 A previous review highlights the potential of such rhythmicity computations in humans (https://journals.sagepub.com/doi/pdf/10.1177/0748730420940483).
 
-At the date of writing this paper, the current version (v1.1) of the library does not offer the possibility to compute rhythmicity with other techniques such as cosinor analysis (\cite{refinetti2016}).
-Therefore, there is potential to expand the library's functionality in the future to include other techniques.
-
-Finally, We invite other researchers to contribute to this library by submitting data, code, and techniques that can be shared with the scientific community.
-By collaborating and sharing these techniques, we can collectively advance our understanding of rhythmicity and its role in various biological systems.
+At the time of writing this paper, the current version (v1.1) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
+However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
+Researchers are encouraged to actively contribute by submitting new data, codes, or methods to the GitHub repository.
+These contributions will be reviewed by the package maintainers and, if approved, integrated into the \pkg{digiRhythm} package.
 

### 2023-05-15 jstatsoft/cleanup.sh

@@ -1,1 +1,2 @@
-rm -rf *.pdf *.tpt *.tex *.out *.bbl *.log *.aux
+cd jstatsoft
+rm -rf *.pdf *.tpt *.tex *.out *.bbl *.log *.aux *.blg

### 2023-05-15 jstatsoft/article.Rnw
Marie_added sections of her ToDoList

@@
 
+\subsection[sliding Diurnality Index (DI)]{sliding Diurnality Index (DI)}{\label{sub:sliding_di}}
+
+The function sliding_DI also computes the diurnality index as described above.
+The difference is that no single definition of day and night was used for the complete calculation period.
+It is possible to include a data frame with different definitions of day_start and day_end, as well as night_start and night_end for each day.
+The data frame should include five columns, one date column, and four columns of timestamps for night_end, day_start, day_end, and night_start.
+All of these columns should be in the POSIXct format.
+This method, which includes different definitions, could be helpful in various experiments.
+For example, in Europe, it could be used to reduce the bias emerging by converting summertime and wintertime to a time zone, such as GMT.
+An additional use could be includig sunset and sundown by using the same definition for night_end and day_start and for day_end and night_start.
+
 %% -- Summary/conclusions/discussion -------------------------------------------
@@
 
-Additionally, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
+Other metrics for calculating circadian rhythms have been previously published.
+For example, Interdaily Stability (IS) is mainly used in human research \cite{witting1990}.
+The Furier-Based Approximation with Threasholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021}.
+Their database depends on the 24h day, but none of them uses the 24h within the calculation.
+The calculation of harmonic frequencies, which sets the activity signal in relation to the external 24h day is unique.
+Therefore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
 For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
-Additionally, \cite{plaza2022gps} used GPS data to estimate rhythmicity.
+Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021} could be used to estimate rhythmicity.
 Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
+In addition, further studies will be conducted, to validate the DFC calculation, by using simulated data.
 

### 2023-05-16 jstatsoft/article.Rnw
Fixing bibs and some commands in the tex file. Modified the SDI a paragraph a bit.

@@
 
-% Two design considerations were followed while developing the package to avoid errors and make the library's outputs easily integrated with third-party \proglang{R} packages and other analysis workflows.
-The first consideration concerns the input data, whereas a specific tabular data structure is required to avoid function misuse and error propagation.
-The second consideration deals with the user-friendliness of the data interoperability, where the functions returns are designed to be easily exploited and integrated with other analysis workflows.
-
 This paper describes the functions of the package and is organized as follows.
@@
 The second step investigates the spectrum by looking at specific frequencies (harmonics and non-harmonics, further details in Figure \ref{fig:reallsp}).
-In a a next step, the statistical significance of frequency peaks is computed.
+In a a next step, the statistical significance of frequency peaks is computed (\ref{sub:significance}).
 Finally, a ratio of the sum of frequency powers is calculated (a detailed breakdown of the calculation is given in Figure \ref{dfc_sum}).
@@
 
-The function sliding_DI also computes the diurnality index as described above.
+The function \fct{sliding\_DI} also computes the diurnality index as described above.
 The difference is that no single definition of day and night was used for the complete calculation period.
-It is possible to include a data frame with different definitions of day_start and day_end, as well as night_start and night_end for each day.
-The data frame should include five columns, one date column, and four columns of timestamps for night_end, day_start, day_end, and night_start.
+It is possible to include a data frame with different definitions of the start/end of day/night (namely, the variable $day\_start$, $day\_end$, and $night\_start$ and $night\_end$).
+The data frame should include five columns, one date column, and four columns of timestamps for start/end of day/night.
 All of these columns should be in the POSIXct format.
+
 This method, which includes different definitions, could be helpful in various experiments.
 For example, in Europe, it could be used to reduce the bias emerging by converting summertime and wintertime to a time zone, such as GMT.
-An additional use could be includig sunset and sundown by using the same definition for night_end and day_start and for day_end and night_start.
+An additional use could be including sunrise and sunset by using the same definition for $day\_start$, $day\_end$, and $night\_start$ and $night\_end$.
 
@@
 Other metrics for calculating circadian rhythms have been previously published.
-For example, Interdaily Stability (IS) is mainly used in human research \cite{witting1990}.
-The Furier-Based Approximation with Threasholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021}.
+For example, Interdaily Stability (IS) is mainly used in human research \cite{witting1990alterations}.
+The Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
 Their database depends on the 24h day, but none of them uses the 24h within the calculation.
@@
 For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
-Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021} could be used to estimate rhythmicity.
+Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity.
 Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.

### 2023-05-17 jstatsoft/article.Rnw
changes marianne and marie

@@
 
-The real-time monitoring of animal behaviour is becoming increasingly important in agriculture.
-It can provide new opportunities to enhance the efficiency and sustainability of livestock production and is used to enhance animal welfare (Wemelsfelder et al., 2001).
-Animal behaviour can be altered by stress, pain or disease, but monitoring these changes in welfare is still a challenge, since some behavioural definitions are unclear and not specifically linked to a unique adverse event (Theurer et al., 2013; Levitis et al., 2009).
-Relevant information about animal welfare can be obtained by studying individual differences in the capacity for behavioural adaptation (Cooper et al., 2005).
-Indeed, during the animals life, significant changes can occur due to internal and external stimuli; these alterations arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours (Berger et al., 2003).
-Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of animal behaviour data are regularly recorded, and can be used to calculate parameters describing biological rhythms (Theurer et al., 2013; Scheibe and Gromann, 2006; Berger et al.
+Internal and external stimuli arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours (Berger et al., 2003).
+Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of behavioural data are regularly recorded, and can be used to calculate parameters describing biological rhythms (Theurer et al., 2013; Scheibe and Gromann, 2006; Berger et al.
 2003).
-Scheibe et al.
-(1999) developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding.
+\cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding of animals.
 The DFC provides an estimate of synchronization between an individual and its environment.
-It is calculated to each 24 h period within rolling seven day periods and predicts values between $0\%$
-and 100 $\%$ (Scheibe et al., 1999, Berger et al., 2003).
+It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative}, Berger et al., 2003).
+XXX suggested to use rolling seven day periods for this computation.
 An organism with a high DFC is highly synchronized with its environment, while lower DFC's indicate a lower synchronization.
 Breakdowns in the DFC response can be caused by stressors (e.g.
-transport, disease, etc.) that affect the time structure of behaviour (Berger et al., 2003; Scheibe et al., 1999).
+handling, disease, etc.) that affect the time structure of behaviour (Berger et al., 2003; Scheibe et al., 1999).
 In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i.
@@
 24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations (Scheibe et al., 1999; Berger et al., 2003).
-Previous studies demonstrated that calves change their behaviour as illness develops (reduced feeding, less social behaviour and increased rest) as an adaptive response to cope with the disease (Owen-Ashley et al., 2006, Dantzer and Kelley, 2007; Borderas et al.
-2008).
-A better understanding of changes in behaviour due to illness or stress of handling is valuable to improve animal welfare and productivity as well as handler safety (Grandin et al., 1989; Kilgore et al., 2005; Borderas et al., 2008; Robért et al., 2011).
+A better understanding of changes in behaviour due to illness or distress is valuable to get further inisight in the organisms affective state % checke reference
+(Grandin et al., 1989; Kilgore et al., 2005; Borderas et al., 2008; Robért et al., 2011).
 Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).
-
 With the increasing availability of sensor technologies to collect behavioral data continuously, further studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}.
-These studies were able to confirm that the rhythm of animals is linked to their state of physical wellbeing.
+These studies applied the DFC method in managed animal husbandry systems.
 However, the application of the DFC was always limited to those with extensive knowledge of statistical programming.
 An open source package specifically for this type of computation opens up the opportunities for livestock research to easily apply these algorithms.
-Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}.
+Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}, which was originally developed for livestock research, but has the potential to be applied in various domains.
 
@@
 
-The function \fct{sliding\_DI} also computes the diurnality index as described above.
-The difference is that no single definition of day and night was used for the complete calculation period.
-It is possible to include a data frame with different definitions of the start/end of day/night (namely, the variable $day\_start$, $day\_end$, and $night\_start$ and $night\_end$).
-The data frame should include five columns, one date column, and four columns of timestamps for start/end of day/night.
-All of these columns should be in the POSIXct format.
+The function \fct{sliding\_DI} computes the diurnality index as described above.
+However, instead of using a single definition of day and night for the complete computation period, a data frame defining different definitions of the start/end of day/night is included.
+The data frame must include five columns, one date column, and four columns of timestamps for start/end of day/night.
+All of these columns must be formatted as POSIXct.
 
-This method, which includes different definitions, could be helpful in various experiments.
-For example, in Europe, it could be used to reduce the bias emerging by converting summertime and wintertime to a time zone, such as GMT.
-An additional use could be including sunrise and sunset by using the same definition for $day\_start$, $day\_end$, and $night\_start$ and $night\_end$.
+This function, which includes different definitions, is useful in various settings, as it allows to improve the computation accuracy.
+For example, in Europe, it could be used to reduce the bias emerging by converting summertime and wintertime to a time zone, such as GMT, or by defining sunrise and sunset.
 
@@
 \section{Discussion} \label{sec:summary}
-The \pkg{digiRhythm} package can make a significant contribution to the assessment of rhythmicity in animals, as it makes the analysis of animal rhythmicity more accessible to researchers.
+The \pkg{digiRhythm} package can make a significant contribution to the assessment of rhythmicity, as it makes the analysis of rhythmicity more accessible to researchers.
 The tool is easy to use, especially because it is delivered as an \proglang{R} package.
@@
 The usage of the package is explained in the section \ref{sec:usage}.
-Further usage examples can be found in the library's vignettes.
+Further examples can be found in the library's vignettes.
 
-The package enables the calculation of rhythmicity in, but not limited to, animals.
-To date, this is purely a computation of activity patterns that have been correlated with physiological welfare factors in previous studies.
-This indicator can be calculated from any type of regularly measured time series data.
-However, we have no evidence how these activity patterns are regulated by routine management practices, such as feeding or milking.
-Such routines could indirectly affect rhythmic activity patterns and blur physiological rhythmicity patterns.
-To date, the package provides the option to exclude known management time spans from the analysis.
-However, in order to increase the understanding of rhythmicity as a welfare indicator future studies should investigate how these management practices affect rhythmicity calculations and evaluate if this is a reasonable welfare indicator for animals depending on external management.
+The package enables the calculation of rhythmicity from activity data.
+To date, this is purely a computation of activity patterns that have been correlated with physiological factors in previous studies.
+This indicator can be calculated from any type of regularly measured time series data, if they are brought in a suitable format, which can be assessed with the  \fct{is_dgm_friendly} function.
+However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
+Such extrinsic factors could affect rhythmic activity patterns and blur physiological rhythmicity patterns.
+To date, the package provides the option to exclude extrinsically structured time spans from the analysis.
+However, in order to increase the understanding of rhythmicity  future studies should investigate how these extrinsic factors affect rhythmicity calculations.
 
-The required frequency of data collection is of utmost importance in animal-borne sensor technologies.
-Sensors are required to be small, so that they don't disturb the animal, and further expected to run without any sensor based maintenance, such as data reading or charging of the device.
+The required frequency of data collection is of utmost importance in sensor technologies.
+Sensors are required to be small and further expected to run without any sensor based maintenance, such as data reading or charging of the device.
 This in turn means that data collection frequency should be low.
-In the current study an extremely low data collection frequency is suggested (15 minutes) future studies should evaluate via downsampling, how low collection frequency can be set to create reliable rhythmicity calculations.
-Further it would be interesting to evaluate if activity data collected from different sources, such as access of feeding stations or watering troughs, compute similar rhythmicity outputs in the same organism.
+In the current study an extremely low data collection frequency is suggested (15 minutes) future studies should evaluate how low collection frequency can be set to create reliable rhythmicity calculations.
+Further it would be interesting to evaluate if activity data collected from different sources, such as  access of feeding stations or coffee maschines, compute similar rhythmicity outputs in the same organism.
 
-Include 7 days aspect.
+Seven days of continuous data collection are required to compute one DFC value.
+This implies that very long data collection intervals are required to detect rhythmic patterns.
+Reducing the number of days required to compute one DFC value would make this rhythmicity computation much more accessible in practice an research.
+The impact of reducing the number of days required to compute the DFC should be investigated in future research.
 

### 2023-05-25 jstatsoft/article.Rnw
Mid day results of changes by Hassan. Added a new amendment tracker with the date of today, fixed some references in the intro and improved the language for some parts.

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmic behavior assessment in livestock}
-\Plaintitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
-\Shorttitle{digiRhythm: an R package for rhythmic behavior assessment in livestock}
+\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmicity assessment using the degree of functional coupling}
+\Plaintitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
+\Shorttitle{digiRhythm: an R package for rhythmicity assessment using the DFC}
 
@@
 \Abstract{
-Measuring an organisms rhythmicity can provide information on its physiological and psychological state.
-What makes this measure so interesting, is that it can be computed from any activity related time series data (e.g., accelerometers, gps tracks, smart watches (step counts), etc.).
-The Degree of Functional Coupling (DFC) is an established method to calculate rhythmicity.
-Computing the DFC involves several computational steps, such as computing a frequency spectrum and determining the statistical significance of its peaks.
-In this paper, we present the \proglang{R} package \pkg{digiRhythm}, which provides functions that allow calculating rhythmicity from times series data.
-Further, additional functions, such as creating actograms, computing and visualizing average activities, and the calculation of an diurnality index are also integrated in this package.
-The package is customized for integration in the research workflow, from raw data to creating publication-ready content and allows the simple application of these sophisticated methods.
-}
+Investigating an organism's rhythmicity yields valuable insights into its physiological and psychological state.
+What sets this measure apart is its applicability to various types of activity-related time series data, including accelerometers, GPS tracks, and smartwatches (e.g., step counts), among others.
+The Degree of Functional Coupling (DFC) represents a well-established method for assessing rhythmicity, encompassing several computational steps like frequency spectrum calculation and the determination of the statistical significance of spectral peaks.
+In this study, we introduce the \proglang{R} package \pkg{digiRhythm}, furnished with functions to compute rhythmicity from time series data.
+Complementary functions for actogram generation, computation and visualization of mean activities, and the calculation of a diurnality index, are also incorporated into this package.
+Tailored for seamless integration into the research workflow, \pkg{digiRhythm} facilitates the transformation of raw data into publication-ready content, thereby simplifying the application of these complex methods.}
 
@@
 
-Internal and external stimuli arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours (Berger et al., 2003).
-Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of behavioural data are regularly recorded, and can be used to calculate parameters describing biological rhythms (Theurer et al., 2013; Scheibe and Gromann, 2006; Berger et al.
-2003).
+Internal and external stimuli arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours \citep{Berger2003}.
+Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of behavioural data are regularly recorded, and can be used to calculate parameters describing biological rhythms \citep{theurer2013remote,scheibe2006application, Berger2003}.
 \cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding of animals.
 The DFC provides an estimate of synchronization between an individual and its environment.
-It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative}, Berger et al., 2003).
+It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}.
 XXX suggested to use rolling seven day periods for this computation.
@@
 Breakdowns in the DFC response can be caused by stressors (e.g.
-handling, disease, etc.) that affect the time structure of behaviour (Berger et al., 2003; Scheibe et al., 1999).
+handling, disease, etc.) that affect the time structure of behaviour \cite{Berger2003, scheibe1999comparative}.
 In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i.
 e.
-24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations (Scheibe et al., 1999; Berger et al., 2003).
+24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
 A better understanding of changes in behaviour due to illness or distress is valuable to get further inisight in the organisms affective state % checke reference
@@
 
-\section{Understanding the degree of functional coupling}{\label{sec:dfc}}
+\section{The degree of functional coupling}{\label{sec:dfc}}
 
-Computing the DFC involves several steps.
-First, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed.
-The second step investigates the spectrum by looking at specific frequencies (harmonics and non-harmonics, further details in Figure \ref{fig:reallsp}).
-In a a next step, the statistical significance of frequency peaks is computed (\ref{sub:significance}).
-Finally, a ratio of the sum of frequency powers is calculated (a detailed breakdown of the calculation is given in Figure \ref{dfc_sum}).
-We started by explaining the the ratio of frequency powers.
-Ultimately, we describe in detail how these frequency powers and their statistical significance can be computed.
+The computation of the Degree of Functional Coupling (DFC) necessitates several steps.
+Initially, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed.
+Following this, the spectrum is analyzed, focusing on specific frequencies (including both harmonics and non-harmonics, with further details in Figure \ref{fig:reallsp}).
+Subsequently, the statistical significance of frequency peaks is calculated (\ref{sub:significance}).
+Ultimately, a ratio of the sum of frequency powers is determined (a detailed breakdown of this calculation is illustrated in Figure \ref{dfc_sum}).
+We commence by elucidating the ratio of frequency powers and then provide a detailed explanation on the computation of these frequency powers and their statistical significance.
 
-According to  \cite{scheibe1999comparative}, the DFC can be computed as follows:
+As suggested by \cite{scheibe1999comparative}, the DFC can be computed using the following formula:
 %
@@
 %
-Another indicator, called the harmonic power, can also be computed similarly:
+In parallel, another indicator termed 'harmonic power' can be calculated similarly:
 %
@@
 %
-Where $ssh$ is the sum of powers of frequencies that are harmonic and statistically significant, $sumsig$ is the sum of powers of frequencies that are significant (harmonic and non-harmonic), and $sumall$ is the sum of the power of all frequencies in the spectrum.
+Here, $ssh$ represents the sum of powers of frequencies that are both harmonic and statistically significant, $sumsig$ refers to the sum of powers of significant frequencies (both harmonic and non-harmonic), and $sumall$ signifies the sum of the power of all frequencies in the spectrum.
 
-Both DFC and HP are used to study the rhythmic behaviors of animals. Both indicators can be computed using the \pkg{digiRhythm} package.
+Both DFC and HP are instrumental in studying the rhythmic behaviors of animals. The \pkg{digiRhythm} package facilitates the computation of both these indicators.
 
-\subsection{The Lomb-Scargle Periodogram (LSP)}{\label{sub:lsp}}
-The Lomb-Scargle Periodogram is the first step in computing the DFC and HP.
-The Lomb–Scargle periodogram (LSP) is widely used within the astronomy community and is well-known for detecting and characterizing periodicity in unevenly sampled time-series \citep{vanderplas2018understanding}.
+\subsection{The Lomb-Scargle Periodogram (LSP)}\label{sub:lsp}
+The Lomb-Scargle Periodogram (LSP) marks the initial step in calculating the DFC and HP.
+It is a well-regarded tool within the astronomy community, commonly used for detecting and characterizing periodicity in unevenly sampled time-series data \citep{vanderplas2018understanding}.
 
-Activity is in the time domain (Figure \ref{fig:signalVslsp}, left) and can be transformed to the frequency domain (Figure \ref{fig:signalVslsp}, right) by applying the LSP.
-In the time domain, we can see the magnitude of the activity at a particular time stamp.
-In the frequency domain we can see the magnitude (power) of oscillations of specific frequencies present in the activity.
-For example, a  sine wave of 24 hours (Figure \ref{fig:signalVslsp}, left, first row) period could be represented with a one-peak periodogram (Figure \ref{fig:signalVslsp}, right, first row); this peak represents the frequency of a sine wave of 24 hours period.
-The same idea could be illustrated for a sine wave of 6 hours period (Figure \ref{fig:signalVslsp}, second row), and so on.
-Moreover, an activity composed of the addition of 24 hours and 6 hour sine waves (Figure \ref{fig:signalVslsp}, left, third row) will result in having two peaks for the periodogram (Figure \ref{fig:signalVslsp}, right, third row).
-Consequently, a periodogram is a tool that allows us to infer the power of each possible sine wave present in the activity.
+In the time domain, we observe the magnitude of activity at specific time points (Figure \ref{fig:signalVslsp}, left).
+By applying the LSP, this activity can be transformed into the frequency domain (Figure \ref{fig:signalVslsp}, right).
+Here, the magnitude (power) of oscillations for specific frequencies present in the activity can be seen.
+For instance, a sine wave with a 24-hour period could be represented with a single-peak periodogram; this peak represents the frequency of a 24-hour sine wave.
+A similar principle can be illustrated for a sine wave with a 6-hour period, and so on.
+Moreover, activity composed of the summation of 24-hour and 6-hour sine waves will result in a periodogram with two peaks.
+Consequently, a periodogram allows us to infer the power of each possible sine wave present in the activity.
 
@@
 \includegraphics[width=1\columnwidth]{signalVslsp}
-  \caption{Examplary harmonics (left) and their corresponding Lomb-Scargle Periodogram (LSP) (right).
-The first two rows show regular sine waves activity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP.
-The bottom row shows the LSP (right) of two sine waves (of 24 hours period and 6 hours period) added (left) together.}
+\caption{Exemplary harmonics (left) and their corresponding Lomb-Scargle Periodogram (LSP) (right).
+The first two rows depict regular sine waves activity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP.
+The bottom row illustrates the LSP (right) of two sine waves (of 24 hours period and 6 hours period) summed (left) together.}
 \label{fig:signalVslsp}
@@
 
-Activity signals acquired from sensors on animals are irregular.
-Therefore, investigating periodic events inside these signals requires a conversion through LSP.
-Figure \ref{fig:realsignal} shows an example of activity data - the motion index (which is a proprietary metric of the
-overall animal activity measured in three dimensions (\cite{icetagiowa})) of a cow. The corresponding LSP is shown in Figure \ref{fig:reallsp}.
+Activity signals from animals tend to be irregular, requiring conversion via the LSP to investigate periodic events within these signals.
+Figure \ref{fig:realsignal} illustrates an example of such data - the motion index (a proprietary metric of the overall animal activity measured in three dimensions \cite{icetagiowa}) of a cow, and its corresponding LSP is depicted in Figure \ref{fig:reallsp}.
 
@@
 \centering
-<<realsignal, fig=TRUE, width=9>>==
+<<realsignal, fig=TRUE, width=9, cache=TRUE>>==
 data("df516b_2", package = 'digiRhythm')
@@
 
-Mathematically, the LSP is computed as follows:
+Mathematically, the LSP is computed using the following formula:
+
 %
@@
 
-Before computing the LSP, we have to determine the frequency grid on which we will apply the formula \ref{eq:lsp}.
-The frequency grid goes from 0 to $f_{max}$ which is the maximum frequency power that be computed from the activity $g_n$.
-$f_{max}$ is determined according to the Shanon-Nyquist theorem (\cite{shannon1928certain, shannon1949mathematical}).
+Before performing the LSP, we must define the frequency grid to which we will apply formula \ref{eq:lsp}.
+This frequency grid ranges from 0 to $f_{max}$, the maximum frequency power that can be computed from the activity $g_n$.
+According to the Shannon-Nyquist theorem, $f_{max}$ is determined \citep{shannon1928certain, shannon1949mathematical}.
 
-Figure \ref{fig:reallsp} represent the Lomb-Scargle periodogram computed from the activity present in Figure \ref{fig:realsignal}.
+Figure \ref{fig:reallsp} represents the Lomb-Scargle periodogram computed from the activity shown in Figure \ref{fig:realsignal}.
 
@@
 \centering
-<<reallsp, fig=TRUE, width=9>>==
+<<reallsp, fig=TRUE, width=9, cache=TRUE>>==
 lsp <- lomb_scargle_periodogram(df[1:672 , c(1,2)],
@@
 \caption{The Lomb-Scargle Periodogram of the activity data shown in Figure \ref{fig:realsignal}.
-Red bars show the amplitude of harmonic frequencies (with the corresponding harmonic period), while green bars represent non-harmonic frequencies.
-The dashed line represents the threshold of significance (frequencies with power less than this threshold are considered non-significant).
-In this particular example, there are only two frequencies with significant power; one of them - which corresponds to a 12-hours period - is harmonic.}
-\label{fig:reallsp}
+Red bars show the amplitude of harmonic frequencies (with corresponding harmonic periods), while green bars represent non-harmonic frequencies.
+The dashed line indicates the significance threshold; frequencies with power below this threshold are considered non-significant.
+In this particular example, only two frequencies with significant power are evident; one of them - corresponding to a 12-hour period - is harmonic.}\label{fig:reallsp}
 \end{figure}
 
-\subsection{Harmonic frequencies and the intuition behind}{\label{sub:harmonic}}
-Animal species exhibit a circadian rhythm reflected in behavioral changes following a 24 hours cycle.
-Therefore, rhythmic behavioral events are reflected in the power of harmonic frequencies of the primary cycle (24 hours).
-These harmonics are the frequencies that correspond to 24 hour periods divided by any integer smaller than 24 (24/1, 24/2, …).
-At this point, it is important to recall the subset of frequency powers previously described.
-These subsets are illustrated in Figure \ref{dfc_sum}.
+\subsection{Understanding Harmonic Frequencies and the Underlying Intuition}\label{sub:harmonic}
+
+Various animal species are characterized by circadian rhythms, signifying their behaviors align with a 24-hour cycle.
+This rhythmicity is manifested as power within the harmonic frequencies of the primary cycle (24 hours).
+Importantly, these harmonics refer to the frequencies linked to periods of 24 hours divided by integers less than 24 (e.g., 24/1, 24/2, …).
+At this juncture, it is crucial to revisit the concept of frequency power subsets that have been discussed earlier.
+These subsets are graphically represented in Figure \ref{dfc_sum}.
 
@@
 \includegraphics[width=1\columnwidth]{sum_dfc}
-  \caption{Frequency peaks found in a Lomb Scargle Periodogram in the context of animals' rhythmicity.
-The whole frequency space consists of frequencies with significant and non-significant peaks, whose powers sum up to $sumall$.
-The powers of the frequencies with significant peaks sum up to $sumsig$.
-A subset of the significant peaks represents harmonic frequencies (24h, 24h/2, 24h/3, \dots) whose powers sum up to $ssh$.}
+\caption{Significant and non-significant frequency peaks depicted in a Lomb-Scargle Periodogram, contextualizing the rhythmicity in animal behaviors.
+The total frequency space includes frequencies with significant and non-significant peaks, the combined powers of which amount to $sumall$.
+The cumulative power of frequencies with significant peaks equates to $sumsig$.
+A subset of these significant peaks corresponds to harmonic frequencies (24h, 24h/2, 24h/3, \dots), and their summed power is represented by $ssh$.}
 \label{dfc_sum}
@@
 
-The relationship between rhythmicity and frequencies would now be clear for the reader.
-If the significant frequencies are only the harmonic ones (i.e., ssh = sumsig), then the DFC = 1, and therefore the animal exhibits a highly rhythmic behavior.
-On the counterpart, if the significant frequencies are not harmonic, then the DFC < 1; therefore, the animal exhibits a low rhythmic behavior., which in turn would indicate a poor physiological state.
+The connection between rhythmicity and frequencies should now be clear to the reader.
+If the frequencies deemed significant are exclusively harmonic (i.e., $ssh = sumsig$), the DFC equals 1, indicative of strong rhythmic behavior in the animal.
+Conversely, if the significant frequencies are not purely harmonic, the DFC value is less than 1, suggesting relatively low rhythmic behavior.
+This phenomenon could be indicative of sub-optimal physiological states.
 
-\subsection{Measuring the statistical significance of a frequency peak}{\label{sub:significance}}
-After computing the periodogram, measuring the statistical significance of  frequency peaks is the second step in the DFC computation.
-The latter would show a peak per frequency.
-Determining whether a certain peak reflects a valid activity component or a noise component is a crucial step.
-In fact, for a small number of samples and a low signal-to-noise ratio, the noise peaks become comparable to the activity peaks.
-Hence, it is important to verify the data by applying a tool that measures the statistical significance of a peak.
-In this regard, the False Alarm Probability (FAP) is the measure that is typically used. The FAP reflects the probability that data with no activity would lead to a peak of similar magnitude.
+\subsection{Evaluating the Statistical Significance of Frequency Peaks}\label{sub:significance}
 
-According to \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom.
-This implies that the probability of observing periodogram values less than a given $Z=P(f_0)$ in an activity formed by only Gaussian noise is:
-%
+The computation of the Degree of Functional Coupling (DFC) involves two critical steps.
+The first is the calculation of the periodogram, and the second is determining the statistical significance of the frequency peaks.
+A periodogram typically displays a distinct peak for each frequency, making the identification of valid activity components from noise components a fundamental task.
+This distinction becomes particularly challenging when dealing with a small sample size and a low signal-to-noise ratio, where noise peaks may mimic the magnitude of activity peaks.
+Consequently, it is imperative to utilize tools that can accurately measure the statistical significance of a peak.
+
+Commonly, the False Alarm Probability (FAP) is employed to serve this purpose. FAP quantifies the likelihood that data devoid of any activity would generate a peak of comparable magnitude.
+
+As reported by Scargle \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom.
+Therefore, the probability of observing periodogram values less than a certain $Z=P(f_0)$ in a data series constituted exclusively by Gaussian noise is expressed as:
+
 \begin{equation}
-  P_{single} = 1 - \exp{Z}
+P_{single} = 1 - \exp{-Z}
 \end{equation}
-%
-To compute the FAP, multiple approaches are available.
-According to \citep{suveges2015comparative}, the Baluev method \citep{baluev2008assessing} is one of the top performing ones.
-This method was chosen to be implemented in \pkg{digiRhythm}.
 
-According to Baluev \citep{baluev2008assessing}, the FAP could be written as:
+Multiple methodologies can be used to calculate the FAP.
+Among these, the Baluev method, according to comparative studies by \citep{suveges2015comparative}, stands as one of the most effective and hence, has been implemented in the \pkg{digiRhythm} package.
+
+As per Baluev \citep{baluev2008assessing}, the FAP can be formalized as follows:
 
@@
 
-\section{digiRhythm functionalities and usage}\label{sec:usage}
-In addition to the ability to compute the DFC and HP, \pkg{digiRhythm} contains much more functions.
-These functions were complied with animal behavior scientists to provide the commonly needed tools for rhythmicity analysis.
-Currently, the library offers the following tools:
+\section{Functionalities and Usage of digiRhythm}\label{sec:usage}
+
+The \pkg{digiRhythm} package offers a range of functionalities beyond merely calculating the Degree of Functional Coupling (DFC) and Harmonic Part (HP).
+In collaboration with animal behavior scientists, we have incorporated a variety of tools typically required for rhythmicity analysis.
+At present, the \pkg{digiRhythm} library includes the following features:
 
 \begin{itemize}
-  \item \fct{lomb\_scargle\_periodogram}: Computing and plotting the Lomb Scargle Periodogram. This function also offers comprehensive visualization of the LSP.
-  \item \fct{dfc}: Computing and plotting the Degree of Functional Coupling and the harmonic part.
-  \item \fct{daily\_average\_activity}: Computing and plotting the daily average activity.
-  \item \fct{diurnality}: Computing and plotting the diurnality index.
-  \item \fct{actogram}: Actogram Visualization.
+\item \fct{lomb\_scargle\_periodogram}: This function not only computes the Lomb Scargle Periodogram (LSP) but also provides an intricate visualization of it.
 
+\item \fct{dfc}: This function allows for both computation and graphical representation of the Degree of Functional Coupling and its harmonic part.
+
+\item \fct{daily\_average\_activity}: This feature facilitates the calculation and plotting of the daily average activity.
+
+\item \fct{diurnality}: With this function, users can compute and graphically present the diurnality index.
+
+\item \fct{actogram}: This tool is designed for actogram visualization.
 \end{itemize}
 
-The \fct{dfc} function is a wrapper of the !MISSING WORD! that users could also use should they wish to investigate more about!MISSING WORD!
-\subsection{The input data format}
-As mentioned in section \ref{sec:intro}, the library accepts input Data frames with a specific structure.
-A typical data frame sampled at 15 min intervals is shown below.
-This data frame is provided with the \pkg{digiRhythm} package and could be fetched using:
+The \fct{dfc} function acts as a wrapper for the \fct{lomb\_scargle\_periodogram}, providing users with the opportunity to delve further into the Lomb-Scargle Periodogram, should they wish to explore it in greater detail.
 
-<<input_data>>==
+\subsection{Format of Input Data}
+
+As referenced in Section \ref{sec:intro}, the \pkg{digiRhythm} package necessitates that input data frames adhere to a specific structure.
+An example of a typically structured data frame, sampled at 15-minute intervals, is displayed below.
+This representative data frame is bundled with the \pkg{digiRhythm} package and can be retrieved using:
+
+<<input_data, cache=TRUE>>==
 data("df516b_2", package = 'digiRhythm')
@@
 
-The above snippet shows a sample activity dataframe where columns represent date-time, the motion index and the number of steps measured during a 15 minutes time interval.
+This excerpt presents a sample activity dataframe. The columns represent date-time, the motion index, and the number of steps measured during each 15-minute time interval.
 
-To make sure that users are using the right format, a utility routine that can check whether the data is digiRhythm friendly or not was included.
-This could be done using the function \fct{is\_dgm\_friendly} as follows:
+To confirm that the data adheres to the required format, we've included a utility function capable of verifying its compatibility with \pkg{digiRhythm}.
+This check can be performed using the \fct{is\_dgm\_friendly} function, as demonstrated below:
 
-<<is_dgm>>==
+<<is_dgm, cache=TRUE>>==
 is_dgm_friendly(data = df, verbose = TRUE)
@@
 
-This function takes a data frame as an argument and checks whether the data is digiRhythm friendly or not by returning a boolean.
-The data is considered \pkg{digiRhythm} friendly if it fulfills all these conditions:
+This function takes a data frame as an argument and verifies its \pkg{digiRhythm} compatibility, returning a boolean.
+The data is considered \pkg{digiRhythm}-friendly if it satisfies the following conditions:
 
 \begin{itemize}
-  \item A data frame is not a NULL object.
-  \item The first column follows the POSIX standard.
+\item The data frame is not a NULL object.
+\item The first column adheres to the POSIX standard.
 \item The data frame contains at least two columns.
-  \item All columns - except the first one - are numeric.
+\item All columns, excluding the first one, are numeric.
 \end{itemize}
 
-The verbose argument helps understand why (or why not) the data is considered (or not) to be digiRhythm friendly. Verbose = TRUE will output information about each condition checked.
-A warning message will appear if the data contains less than seven days because at least seven days are required to compute the DFC and HP, yet other functionalities could be performed.
+The verbose argument aids in understanding why the data is or isn't considered \pkg{digiRhythm}-friendly. If verbose = TRUE, the function will provide information about each condition checked.
+
+A warning message will be generated if the data encompasses less than seven days, since at least a week's worth of data is needed to compute the DFC and HP.
+However, other functionalities may still be executable.
 
@@
 
-<<dfc>>==
+<<dfc, cache=TRUE>>==
 <<dfc_snippet, results=hide>>=
@@
 
-<<print_dfc>>==
+<<print_dfc, cache=TRUE>>==
 head(my_dfc$data)
@@
 \centering
-<<plot_dfc, fig=TRUE, cache = TRUE, width=9>>==
+<<plot_dfc, fig=TRUE, cache = TRUE, width=9, cache=TRUE>>==
 library("ggplot2")
@@
 \centering
-<<dfc_plot_modified, fig=TRUE, width=9>>==
+<<dfc_plot_modified, fig=TRUE, width=9, cache=TRUE>>==
 new_plot <- my_dfc +
@@
 \centering
-<<actogram, fig=TRUE, width=9>>==
+<<actogram, fig=TRUE, width=9, cache=TRUE>>==
 start <- "2020-05-01" #year-month-day
@@
 \centering
-<<avg_activity, fig=TRUE, width=9>>==
+<<avg_activity, fig=TRUE, width=9, cache=TRUE>>==
 daily_average_activity(df, activity, activity_alias,
@@
 
-<<setting_times>>==
+<<setting_times, cache=TRUE>>==
 day_time <- c("06:30:00", "16:30:00")
@@
 \centering
-<<diurnality, fig=TRUE>>==
+<<diurnality, fig=TRUE, cache=TRUE>>==
 my_di <- diurnality(df, activity, day_time, night_time)
@@
 To date, this is purely a computation of activity patterns that have been correlated with physiological factors in previous studies.
-This indicator can be calculated from any type of regularly measured time series data, if they are brought in a suitable format, which can be assessed with the  \fct{is_dgm_friendly} function.
+This indicator can be calculated from any type of regularly measured time series data, if they are brought in a suitable format, which can be assessed with the  \fct{is\_dgm\_friendly} function.
 However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.

### 2023-05-25 jstatsoft/article.Rnw
Second pass of the day. All improvements done.

@@
 
-\subsection[The dfc function]{The \fct{dfc} function}{\label{sub:dgm_fct}}
-The DFC and HP are both implemented (\ref{alg:dfc}) in the \fct{dfc} function.
-The function evaluates a \pkg{digiRhythm} friendly data frame, by analyzing the start, and end date, and the variable for which the user wants to run the analysis.
-It then applies a Lomb Scargle Periodogram over a sliding window of seven days at a time.
-For instance, the DFC and HP are computed for the days one to seven, then for the days two to eight, and so on.
-The last six days of the data set would not be considered because it is not recommended to use the LSP (in the context of computing the DFC) for less than seven days.
-Figure \ref{fig:dfc} shows a typical DFC/HP output.
+\subsection[The dfc Function]{The \fct{dfc} Function}\label{sub:dgm_fct}
+
+The Degree of Functional Coupling (DFC) and Harmonic Part (HP) are both implemented in the \fct{dfc} function, as per Algorithm \ref{alg:dfc}.
+This function processes a \pkg{digiRhythm}-compatible data frame by calculating the DFC for the chosen variable targeted for analysis.
+Initially, it evaluates the start and end dates to ensure a sufficient number of days (a minimum of seven by default) are available for the computation.
+The function then applies a Lomb Scargle Periodogram to a seven-day sliding window.
+For instance, the DFC and HP are computed for the first seven days, then for days two to eight, and so on.
+The final six days of the dataset are excluded, as applying the LSP to periods shorter than seven days is not recommended when computing the DFC.
+A typical DFC/HP output is illustrated in Figure \ref{fig:dfc}.
 
 \begin{algorithm}
-\caption{The Degree of Functional Coupling algorithm}\label{alg:dfc}
+\caption{The Degree of Functional Coupling Algorithm}\label{alg:dfc}
 \begin{algorithmic}
@@
 \For{\texttt{day in 7:days}}
-        \State \texttt{Select data from the day ($day-7$) to day ($day$)}
+        \State \texttt{Select data from day ($day-7$) to day ($day$)}
         \State \texttt{Compute the Lomb Scargle Periodogram of the selected data}
@@
 \end{algorithm}
+\subsection[Utilizing digiRhythm for DFC and HP Computation]{Utilizing \pkg{digiRhythm} for DFC and HP Computation}\label{sub:dgm_dfc_hp}
 
-\subsection[Using digiRhythm to compute DFC and HP]{Using \pkg{digiRhythm} to compute DFC and HP}{\label{sub:dgm_dfc_hp}}
-
-The DFC and HP can be computed using the \fct{dfc} function.
-A recommended workflow is shown below.
-After reading the data and verfying its compatibility with the library, the data (originally sampled with 1 min) is resampled to 15 minutes.
-In this context, resampling means simple summation.
-For instance, if the dataset contains the number of steps with 1 minute sampling period, resampling to 15 minutes means adding the steps during each 15 minutes interval.
-Further, an activity is chosen (in this example, we choose the activity from the second column (motion index)), then the DFC is computed using the \fct{dfc} function as shown below:
+The DFC and HP can be calculated employing the \fct{dfc} function.
+A suggested workflow is outlined below.
+Once the data is loaded and its compatibility with the library is verified, the data—originally sampled at 1-minute intervals—is resampled to 15-minute intervals.
+In this context, 'resampling' refers to a simple summation process.
+For instance, if the dataset includes the number of steps recorded at 1-minute intervals, resampling to 15 minutes entails summing the steps taken during each 15-minute interval.
+Afterward, an activity is selected (in this example, we select the activity from the second column, i.e., motion index), and the DFC is calculated using the \fct{dfc} function as illustrated below:
 
@@
 
-The returned data frame  contains the DFC and HP measured along a sliding window of seven days as follows:
+The returned data frame, \code{my_dfc}, contains the DFC and HP measured within a sliding window of seven days as follows:
 
@@
 
-The returned variable \code{my_dfc} contains both the data and the plot as it is a \code{ggplot2} object.
+The returned variable \code{my_dfc} is a \code{ggplot2} object, comprising both the data and the plot.
 
@@
 
-It is possible to access the computed data and the user can adjust the plot features.
-Calling the data field from the variable \code{my_dfc} will allow accessing the computed DFC and HP as shown below:
+You can access the computed data directly and modify the plot features according to your preference.
+Accessing the data field from the \code{my_dfc} variable provides the computed DFC and HP as shown below:
 
-It is, therefore, easy to access the resulting DFC and HP and use them in further analysis like statistical modeling or more customized workflows.
-Because the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, it is also possible to change the plot directly by adding a layer on the aesthetics of the \code{my_dfc} object as shown below:
+Thus, it's straightforward to access the resulting DFC and HP for further analysis such as statistical modeling or other tailored workflows.
+As the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, you can directly modify the plot by adding a layer to the aesthetics of the \code{my_dfc} object as demonstrated below:
 
@@
 
-\caption{\label{fig:dfc_modified} Output plot from Figure \ref{fig:dfc} modified by adding a layer of aesthetics causing some modifications to the \fct{ggplot} output returned by the \fct{dfc}.
-In particular, the font family was changed to \emph{Times}, the font size to 8 points, and the y-axis text was changed to a bigger size (15) and colored red.}
+\caption{\label{fig:dfc_modified} Output plot from Figure \ref{fig:dfc} modified by augmenting a layer of aesthetics to the \fct{ggplot} output returned by the \fct{dfc} function.
+Specifically, the font family was updated to \emph{Times}, the font size adjusted to 8 points, and the y-axis text was enlarged (15 points) and colorized in red.}
 \end{figure}
 
-This is an illustrative yet straightforward example to sketch the feature of reusing the graphical outputs of the library, but the application is useful when users need to produce publication-quality pictures in a specific format.
-It is important to mention that this feature is possible because the functions were designed in a way they return \fct{ggplot} objects rather than base \proglang{R} objects.
+This serves as a simple, yet illustrative example of how the graphical outputs of the library can be repurposed.
+Such a feature becomes especially beneficial when users need to generate publication-quality figures adhering to specific formats.
+Importantly, this capability is made possible because the functions in the \pkg{digiRhythm} library are designed to return \fct{ggplot} objects instead of standard \proglang{R} objects.
 
-\subsection[Recommendations to use the dfc function]{Recommendations to use the dfc function}{\label{sub:recom_dfc}}
+Although the \fct{dfc} function can accommodate data with any sampling frequency, several authors \citep{sinz1976systemanalyse, sarout2018assessment, fuchs2022detecting, scheibe1999comparative} have suggested that optimal insights can be achieved using a seven-day sliding window with a 15-minute sampling period.
+To maintain simplicity, the current version does not offer an option to alter the length of the sliding window (e.g., to 6, 8, or more days).
+However, this could potentially be implemented in future versions, if found to be useful.
 
-Although the \fct{dfc} function can accept data with any sampling frequency, it is suggested by several authors \citep{sinz1976systemanalyse, sarout2018assessment, fuchs2022detecting, scheibe1999comparative} that optimal insights could be obtained using seven days sliding window with 15 minutes sampling period.
-For the sake of simplicity, an option for changing the sliding window length (for example, 6 or 8, or more days) was not implemented.
-However, this could be done in future versions if proven useful.
+The actogram is typically the first tool used when working with animal activity.
+It is a 2D tiled representation where the x-axis represents the time of the day, the y-axis represents the number of days (or date), and the tiles are color-coded according to the activity intensity at each sample.
 
-\subsection[Actogram visualization]{Actogram visualization}{\label{sub:acto}}
-
-The actogram is often the first-to-go tool when working with animal activity.
-It is a tiled 2D representation where the x-axis represents the time of the day, the y-axis represents the number of days (or date), and the tiles are color-coded according to the activity intensity at each sample.
-
-Figure \ref{fig:actogram} shows a code snippet for visualizing the actogram of the dataset \fct{df516b\_2} and its corresponding actogram.
+The code snippet in Figure \ref{fig:actogram} illustrates how to visualize the actogram of the dataset \fct{df516b\_2}, and the corresponding actogram is displayed alongside.
 
@@
 
-The average activity is an indicator of how active an animal is on average during times of the day. The average activity is computed using the following formula:
+The average activity provides an indication of how active an animal is on average at different times of the day. This average activity is calculated using the formula:
 
@@
 
-Where $T_i$ is a specific time of the day, $d$ is the number of days, and $A(T_{id})$ is the activity value at the time $T_i$ during the day $d$.
+In this formula, $T_i$ represents a specific time of day, $d$ stands for the total number of days, and $A(T_{id})$ denotes the activity level at time $T_i$ on day $d$.
 
-The function \fct{daily\_average\_activity} computes and plots this indicator as shown in Figure \ref{fig:avg_act}.
+The function \fct{daily\_average\_activity} is used to calculate and plot this average activity indicator, as shown in Figure \ref{fig:avg_act}.
 
@@
 
-The Diurnality Index is also an important measure when it comes to assessing animals' rhythmicity.
-This measure indicates whether the animal exercises its activity more during diurnal (daytime) or nocturnal (nighttime) periods.
-Using the Diurnality function, the user can specify day and night time.
-The user can also omit some periods during the day, such as milking time, where the milking process controls the animal's activity rather than its behavior.
-The Diurnality index is computed according to \cite{hoogenboom1984seasonal}:
+The Diurnality Index (DI) is another crucial measure used to assess animal rhythmicity.
+This index provides insights into whether an animal tends to be more active during the day (diurnal) or at night (nocturnal).
+When using the Diurnality function, users can define specific daytime and nighttime periods.
+Additionally, certain periods during the day can be excluded from the computation, such as milking time for dairy animals, during which the animal's activity is controlled more by the milking process than by its natural behaviors.
+
+The Diurnality Index is computed according to the formula proposed by \cite{hoogenboom1984seasonal}:
 
@@
 
-Where $C_d$ and $C_n$ are the summations of activity during the daytime and nighttime, respectively.
-$T_d$ and $T_n$ are the numbers of activity samples during the daytime and nighttime.
-The number of samples is usually measured from the sampling time.
-For instance, if the activity is measured with 15 min time sample, and the daytime is considered to be between 06h00 AM and 04h00 PM, then $T_d$ is considered equal to 4 samples per hour × 10 hours = 40 samples.
-The Diurnality function takes the daytime and nighttime as arguments, with default daytime as 06h30 AM till 04h30 PM and default nighttime as 18h00 pm till 05 AM the next day.
-To change these times, the configuration should follow an HH:MM:SS format as shown below:
+In this formula, $C_d$ and $C_n$ represent the sum of activity during the daytime and nighttime, respectively.
+$T_d$ and $T_n$ are the number of activity samples recorded during the daytime and nighttime.
+The number of samples is typically dependent on the sampling frequency.
+For instance, if activity is measured at 15-minute intervals, and daytime is defined as the period between 06:00 AM and 04:00 PM, then $T_d$ would be calculated as 4 samples per hour × 10 hours = 40 samples.
+
+By default, the Diurnality function defines daytime as the period from 06:30 AM to 04:30 PM (local time) and nighttime as the period from 06:00 PM to 05:00 AM (local time) the next day.
+To adjust these periods, users need to input their desired times in the format HH:MM:SS, as demonstrated below:
 
@@
 
-Figure \ref{fig:di} shows how to compute and visualize the diurnality index for an example dataset.
+The \fct{diurnality} function computes the Diurnality Index (DI) for the given dataset, and the resultant plot provides a visual representation of the computed DI.
+As shown in Figure \ref{fig:di}, the plot allows users to assess the temporal activity distribution of the dataset, providing a valuable overview of the dataset's diurnal and nocturnal activity patterns.
+
+The computation process and the graphical representation of the Diurnality Index for a sample dataset are demonstrated below:
 
@@
 \end{figure}
+As seen in the output figure, the calculated Diurnality Index is displayed along with the daytime and nighttime activity distributions.
+This output can provide valuable insights into the temporal behavioral patterns of the animal or group of animals under study.
 
@@
 
-The function \fct{sliding\_DI} computes the diurnality index as described above.
-However, instead of using a single definition of day and night for the complete computation period, a data frame defining different definitions of the start/end of day/night is included.
-The data frame must include five columns, one date column, and four columns of timestamps for start/end of day/night.
-All of these columns must be formatted as POSIXct.
+The \fct{sliding\_DI} function provides a more nuanced approach to calculating the Diurnality Index (DI).
+It allows for dynamic definitions of day and night time periods across the dataset, rather than using a static, single definition.
+This is particularly useful when dealing with data that spans across different seasons, daylight saving time changes, or different geographical locations where the length of the day varies significantly.
 
-This function, which includes different definitions, is useful in various settings, as it allows to improve the computation accuracy.
-For example, in Europe, it could be used to reduce the bias emerging by converting summertime and wintertime to a time zone, such as GMT, or by defining sunrise and sunset.
+As mentioned, the \fct{sliding\_DI} function requires a data frame that includes a date column and four timestamp columns, indicating the start and end times of the day and night.
+These timestamps must be in POSIXct format.
+
+An example of using the \fct{sliding\_DI} function might look like this:
+
+**MARIE: ADD CODE HERE**
+
+In the above example, the day starts at 6 AM and ends at 6 PM, while the night starts at 6 PM and ends at 6 AM.
+Note that the \fct{sliding\_DI} function would then compute the DI for each day in the dataset based on the specified day/night periods.
+This could provide a more accurate reflection of the animal's behavior, as it accounts for changes in daylight hours throughout the year.
 
@@
 
+\section{Discussion (New proposal Hassan)}
+
+The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of Rhythmicity, animal behavior, and beyond.
+The package offers a versatile approach to assessing rhythmicity based on digital activity records.
+By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
+
+While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
+This opens up exciting avenues for future research in various domains.
+The unique utilization of harmonic frequencies related to the 24-hour day in the package provides a novel approach for rhythmicity analysis.
+
+The package is designed with flexibility and expansion in mind.
+The current version allows for the computation of different metrics like DFC, Diurnality Index, and visualization tools such as actograms.
+Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
+As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become a an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
+
+Lastly, we welcome contributions to the \pkg{digiRhythm} package.
+The open-source nature of the package fosters collaboration and continuous improvement.
+Researchers and developers are encouraged to contribute their codes, methodologies, and data to the project, thereby supporting the scientific community's collaborative spirit.
+By doing so, we hope to further our understanding of rhythmicity, its implications, and how we can leverage this knowledge for the welfare of animals and humans alike.
+
 \bibliography{refs}

### 2023-06-01 jstatsoft/article.Rnw
Marianne's changes

@@
 
-%% -- Summary/conclusions/discussion -------------------------------------------
+% Summary/conclusions/discussion
 
@@
 
-The package enables the calculation of rhythmicity from activity data.
-To date, this is purely a computation of activity patterns that have been correlated with physiological factors in previous studies.
-This indicator can be calculated from any type of regularly measured time series data, if they are brought in a suitable format, which can be assessed with the  \fct{is\_dgm\_friendly} function.
+%pacckage enables computation of any activity data
+The \pkg{digiRhythm} package enables the calculation of rhythmicity from any activity data.
+This indicator can be calculated from any type of regularly measured time series data, if the data are prepared in a suitable format.
+The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function.
+To date, this is purely a computation of activity patterns, resulting the in the DFC.
+This measure has been correlated with physiological functions in previous studies.
+
+% Scope of DFC application
+The computation of rhythmicity allows to gain further insight into an organisms harmonic patterns.
+Organisms in a good physiological state will present harmonic locomotory patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
+Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions.
+In the past these rhythmic patterns have been discussed with regard to the psychological state of a humans, where psychological challenges, such as bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
+%Limitation of DFC computation
+
 However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
@@
 
+<<>>=
+
+@
+
 While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.

### 2023-06-02 jstatsoft/article.Rnw
Commented a buging line
Included Sliding_DI function and worked on the complete text

@@
 What sets this measure apart is its applicability to various types of activity-related time series data, including accelerometers, GPS tracks, and smartwatches (e.g., step counts), among others.
-The Degree of Functional Coupling (DFC) represents a well-established method for assessing rhythmicity, encompassing several computational steps like frequency spectrum calculation and the determination of the statistical significance of spectral peaks.
+The Degree of Functional Coupling (DFC) represents a method for assessing rhythmicity, encompassing several computational steps like frequency spectrum calculation and the determination of the statistical significance of spectral peaks.
 In this study, we introduce the \proglang{R} package \pkg{digiRhythm}, furnished with functions to compute rhythmicity from time series data.
@@
   Marie Schneider\\
-  Johann Heinrich von Th\"unen Institut für \"Okologischen Landbau\\
+  Johann Heinrich von Th\"unen Institute of Organic Farming\\
   Trenthorst 32\\
   23847 Westerau\\
-  Deutschland \\
+  Germany \\
   \emph{and}\\
-  Universit\"at Hohenheim – Institut f\"ur Verfahrenstechnik der Tierhaltungssysteme \\
+  University of Hohenheim –  Department of Livestock Systems Engineering\\
   Garbenstra{\ss}e 9 \\
   70599 Stuttgart\\
-  Deutschland\\
+  Germany\\
   E-mail: \email{marie.schneider@thuenen.de}\\
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE} %For some reason, this gives an error!
 
@@
 \cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding of animals.
-The DFC provides an estimate of synchronization between an individual and its environment.
+The DFC provides an estimate of synchronization between an individual and the 24-hour day of the environment.
 It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}.
 XXX suggested to use rolling seven day periods for this computation.
-An organism with a high DFC is highly synchronized with its environment, while lower DFC's indicate a lower synchronization.
+An organism with a high DFC is highly synchronized with the 24-hour day of its environment, while lower DFC's indicate a lower synchronization.
 Breakdowns in the DFC response can be caused by stressors (e.g.
@@
 
-Both DFC and HP are instrumental in studying the rhythmic behaviors of animals. The \pkg{digiRhythm} package facilitates the computation of both these indicators.
+Both DFC and HP are instrumental in studying the rhythmic behaviors of animals.
+The \pkg{digiRhythm} package facilitates the computation of both these indicators.
+A further study showed, that the DFC reduces during an adaption process, whereas HP reduces for a longer time during diseases.
 
@@
 Here, the magnitude (power) of oscillations for specific frequencies present in the activity can be seen.
-For instance, a sine wave with a 24-hour period could be represented with a single-peak periodogram; this peak represents the frequency of a 24-hour sine wave.
+For instance, a sine wave with a 24-hour period could be represented by a single-peak periodogram, with a peak representing the frequency of a 24-hour sine wave.
 A similar principle can be illustrated for a sine wave with a 6-hour period, and so on.
@@
 
-Activity signals from animals tend to be irregular, requiring conversion via the LSP to investigate periodic events within these signals.
+Activity signals from individuals tend to be irregular, requiring conversion via the LSP to investigate periodic events within these signals.
 Figure \ref{fig:realsignal} illustrates an example of such data - the motion index (a proprietary metric of the overall animal activity measured in three dimensions \cite{icetagiowa}) of a cow, and its corresponding LSP is depicted in Figure \ref{fig:reallsp}.
@@
 \caption{The Lomb-Scargle Periodogram of the activity data shown in Figure \ref{fig:realsignal}.
-Red bars show the amplitude of harmonic frequencies (with corresponding harmonic periods), while green bars represent non-harmonic frequencies.
+Red bars show the amplitude of harmonic frequencies (with corresponding harmonic periods).
+While green bars represent non-harmonic frequencies.
 The dashed line indicates the significance threshold; frequencies with power below this threshold are considered non-significant.
@@
 
-Various animal species are characterized by circadian rhythms, signifying their behaviors align with a 24-hour cycle.
+As mentioned before, most individuals are characterized by circadian rhythms, signifying their behaviors align with a 24-hour cycle.
 This rhythmicity is manifested as power within the harmonic frequencies of the primary cycle (24 hours).
@@
 Conversely, if the significant frequencies are not purely harmonic, the DFC value is less than 1, suggesting relatively low rhythmic behavior.
-This phenomenon could be indicative of sub-optimal physiological states.
+This phenomenon could be indicative of sub-optimal physiological or psychological states.
 
@@
 
+\item \fct{sliding\_DI}: With this function, users can compute and graphically present the diurnality index, using a specific definitions of day/night per day.
+
 \item \fct{actogram}: This tool is designed for actogram visualization.
@@
 
-A warning message will be generated if the data encompasses less than seven days, since at least a week's worth of data is needed to compute the DFC and HP.
+A warning message will be generated if the data encompasses less than seven days, since at the moment, at least a week's worth of data is needed to compute the DFC and HP.
 However, other functionalities may still be executable.
@@
 
-You can access the computed data directly and modify the plot features according to your preference.
-Accessing the data field from the \code{my_dfc} variable provides the computed DFC and HP as shown below:
+The computed data can be accessed directly and the plot features can be modified. Accessing the data field from the \code{my_dfc} variable provides the computed DFC and HP as shown below:
 
 Thus, it's straightforward to access the resulting DFC and HP for further analysis such as statistical modeling or other tailored workflows.
-As the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, you can directly modify the plot by adding a layer to the aesthetics of the \code{my_dfc} object as demonstrated below:
+As the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, the plot can directly be modified, by adding a layer to the aesthetics of the \code{my_dfc} object as demonstrated below:
 
@@
 
-The actogram is typically the first tool used when working with animal activity.
+The actogram is typically the first tool used when working with activity data.
 It is a 2D tiled representation where the x-axis represents the time of the day, the y-axis represents the number of days (or date), and the tiles are color-coded according to the activity intensity at each sample.
@@
 
-The average activity provides an indication of how active an animal is on average at different times of the day. This average activity is calculated using the formula:
+The average activity provides an indication of how active an individual is on average at different times of the day. This average activity is calculated using the formula:
 
@@
 
-The Diurnality Index (DI) is another crucial measure used to assess animal rhythmicity.
-This index provides insights into whether an animal tends to be more active during the day (diurnal) or at night (nocturnal).
-When using the Diurnality function, users can define specific daytime and nighttime periods.
+The Diurnality Index (DI) is another crucial measure used to assess an individual's rhythmicity.
+This index provides insights into whether an individual tends to be more active during the day (diurnal) or at night (nocturnal).
+When using the \fct{diurnality} function, users can define specific daytime and nighttime periods.
 Additionally, certain periods during the day can be excluded from the computation, such as milking time for dairy animals, during which the animal's activity is controlled more by the milking process than by its natural behaviors.
@@
 
-By default, the Diurnality function defines daytime as the period from 06:30 AM to 04:30 PM (local time) and nighttime as the period from 06:00 PM to 05:00 AM (local time) the next day.
+By default, the \fct{diurnality} function defines daytime as the period from 06:30 AM to 04:30 PM (local time) and nighttime as the period from 06:00 PM to 05:00 AM (local time) the next day.
 To adjust these periods, users need to input their desired times in the format HH:MM:SS, as demonstrated below:
@@
 %
-\caption{\label{fig:di} The Diurnality index of the dataset \fct{df516b\_2}.}
+\caption{\label{fig:di} The Diurnality Index of the dataset \fct{df516b\_2}.}
 \end{figure}
@@
 These timestamps must be in POSIXct format.
+An example dataset of a typically definition of daytime and nighttime, by excluding milking times in dairy farms is displayed below.
+This data frame is bundled with the \pkg{digiRhythm} package and can be retrieved using:
 
-An example of using the \fct{sliding\_DI} function might look like this:
+<<input_data, cache=TRUE>>==
+data("timedata", package = 'digiRhythm')
+td <- timedata
+head(td)
+str(td)
+@
 
-**MARIE: ADD CODE HERE**
+An example of using the \fct{sliding\_DI} function with the example activity data and time data might look like this:
 
-In the above example, the day starts at 6 AM and ends at 6 PM, while the night starts at 6 PM and ends at 6 AM.
-Note that the \fct{sliding\_DI} function would then compute the DI for each day in the dataset based on the specified day/night periods.
-This could provide a more accurate reflection of the animal's behavior, as it accounts for changes in daylight hours throughout the year.
+\begin{figure}[h!]
+\centering
+<<sliding\_DI, fig=TRUE, cache=TRUE>>==
+my_sliding_di <- sliding_DI(df, activity, td)
+@
+%
+\caption{\label{fig:di} The Diurnality Index of the dataset \fct{df516b\_2}, calculated using the \fct{sliding\_DI} function.}
+\end{figure}
+
+In the above example for the \fct{diurnality} function, the day starts at 6:30 AM and ends at 4:30 PM, while the night starts at 6 PM and ends at 5 AM.
+Note that the \fct{sliding\_DI} function would compute the DI for each day in the dataset based on the specified day/night periods, of that day.
+This could provide a more accurate reflection of the individual's behavior, as it accounts for changes in daylight hours throughout the year.
 
@@
 The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function.
-To date, this is purely a computation of activity patterns, resulting the in the DFC.
+To date, this is purely a computation of activity patterns, resulting in the DFC.
 This measure has been correlated with physiological functions in previous studies.
@@
 % Scope of DFC application
-The computation of rhythmicity allows to gain further insight into an organisms harmonic patterns.
-Organisms in a good physiological state will present harmonic locomotory patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
+The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns.
+Individuals in a good physiological state will present harmonic locomotory patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
 Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions.
@@
 This in turn means that data collection frequency should be low.
-In the current study an extremely low data collection frequency is suggested (15 minutes) future studies should evaluate how low collection frequency can be set to create reliable rhythmicity calculations.
+In the current study a low data collection frequency is suggested (15 minutes) future studies should evaluate how low collection frequency can be set to create reliable rhythmicity calculations.
 Further it would be interesting to evaluate if activity data collected from different sources, such as  access of feeding stations or coffee maschines, compute similar rhythmicity outputs in the same organism.
-
 Seven days of continuous data collection are required to compute one DFC value.
 This implies that very long data collection intervals are required to detect rhythmic patterns.
-Reducing the number of days required to compute one DFC value would make this rhythmicity computation much more accessible in practice an research.
-The impact of reducing the number of days required to compute the DFC should be investigated in future research.
+Reducing the number of days required to compute one DFC value would make this rhythmicity computation much more accessible in practice and research.
+However, the affect of reducing the intervals are not investigated, yet.
+Therefore, the impact of reducing the number of days required to compute the DFC should be investigated in future research.
 
-Other metrics for calculating circadian rhythms have been previously published.
-For example, Interdaily Stability (IS) is mainly used in human research \cite{witting1990alterations}.
-The Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
-Their database depends on the 24h day, but none of them uses the 24h within the calculation.
-The calculation of harmonic frequencies, which sets the activity signal in relation to the external 24h day is unique.
+In addition to the open questions about the data collection frequency and intervals, a comparison to other metrics will be needed in further studies.
+In the past, other metrics for calculating circadian rhythms have been published.
+For example, Interdaily Stability (IS), which is mainly used in human research \cite{witting1990alterations}.
+And the Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
+Their database depends on the 24 h day, but none of them uses the 24 h within the calculation.
+The calculation of harmonic frequencies within the DFC calculation, which sets the activity signal in relation to the external 24h day is unique.
 Therefore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
-For instance, \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
+In addition to the mentioned metrics (IS and FBAT), \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
 Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity.
@@
 Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
-Many humans nowadays continuously collect activity data with their smart-watches or the phones.
+Many humans nowadays continuously collect activity data with their smart-watches or their phones.
 Thus, it would be interesting to evaluate the application of the \pkg{digiRhythm} package in humans, offering an objective indicator of human welfare.
@@
 Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
-As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become a an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
+As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
 

### 2023-06-02 jstatsoft/article.Rnw
included references for the 7 days and 15 min

@@
 In this study, we introduce the \proglang{R} package \pkg{digiRhythm}, furnished with functions to compute rhythmicity from time series data.
-Complementary functions for actogram generation, computation and visualization of mean activities, and the calculation of a diurnality index, are also incorporated into this package.
+Complementary functions for actogram generation, computation and visualization of mean activities, and the calculation of a Diurnality Index, are also incorporated into this package.
 Tailored for seamless integration into the research workflow, \pkg{digiRhythm} facilitates the transformation of raw data into publication-ready content, thereby simplifying the application of these complex methods.}
@@
 It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}.
-XXX suggested to use rolling seven day periods for this computation.
+\cite{sarout2018assessment} and \cite{fuchs2022detecting} suggested to use rolling seven day periods for this computation.
 An organism with a high DFC is highly synchronized with the 24-hour day of its environment, while lower DFC's indicate a lower synchronization.
@@
 
-The computation of the Degree of Functional Coupling (DFC) necessitates several steps.
+The computation of the DFC necessitates several steps.
 Initially, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed.
@@
 
-The computation of the Degree of Functional Coupling (DFC) involves two critical steps.
+The computation of the DFC involves two critical steps.
 The first is the calculation of the periodogram, and the second is determining the statistical significance of the frequency peaks.
@@
 
-The \pkg{digiRhythm} package offers a range of functionalities beyond merely calculating the Degree of Functional Coupling (DFC) and Harmonic Part (HP).
+The \pkg{digiRhythm} package offers a range of functionalities beyond merely calculating the DFC and Harmonic Part (HP).
 In collaboration with animal behavior scientists, we have incorporated a variety of tools typically required for rhythmicity analysis.
@@
 
-The Degree of Functional Coupling (DFC) and Harmonic Part (HP) are both implemented in the \fct{dfc} function, as per Algorithm \ref{alg:dfc}.
+The DFC and HP are both implemented in the \fct{dfc} function, as per Algorithm \ref{alg:dfc}.
 This function processes a \pkg{digiRhythm}-compatible data frame by calculating the DFC for the chosen variable targeted for analysis.
@@
 
-The Diurnality Index is computed according to the formula proposed by \cite{hoogenboom1984seasonal}:
+The DI is computed according to the formula proposed by \cite{hoogenboom1984seasonal}:
 
@@
 
-The \fct{diurnality} function computes the Diurnality Index (DI) for the given dataset, and the resultant plot provides a visual representation of the computed DI.
+The \fct{diurnality} function computes the DI for the given dataset, and the resultant plot provides a visual representation of the computed DI.
 As shown in Figure \ref{fig:di}, the plot allows users to assess the temporal activity distribution of the dataset, providing a valuable overview of the dataset's diurnal and nocturnal activity patterns.
 
-The computation process and the graphical representation of the Diurnality Index for a sample dataset are demonstrated below:
+The computation process and the graphical representation of the DI for a sample dataset are demonstrated below:
 
@@
 \end{figure}
-As seen in the output figure, the calculated Diurnality Index is displayed along with the daytime and nighttime activity distributions.
+As seen in the output figure, the calculated DI is displayed along with the daytime and nighttime activity distributions.
 This output can provide valuable insights into the temporal behavioral patterns of the animal or group of animals under study.
@@
 
-The \fct{sliding\_DI} function provides a more nuanced approach to calculating the Diurnality Index (DI).
+The \fct{sliding\_DI} function provides a more nuanced approach to calculating the DI.
 It allows for dynamic definitions of day and night time periods across the dataset, rather than using a static, single definition.
@@
 The package is designed with flexibility and expansion in mind.
-The current version allows for the computation of different metrics like DFC, Diurnality Index, and visualization tools such as actograms.
+The current version allows for the computation of different metrics like DFC, DI, and visualization tools such as actograms.
 Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.

### 2023-06-02 jstatsoft/article.Rnw
fixed few compile erros, added a ref in the bib

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE} %For some reason, this gives an error!
@@
 data("timedata", package = 'digiRhythm')
-td <- timedata
+td <- as.data.frame(timedata)
 head(td)
@@
 \centering
-<<sliding\_DI, fig=TRUE, cache=TRUE>>==
+<<sliding_DI, fig=TRUE, cache=TRUE>>==
 my_sliding_di <- sliding_DI(df, activity, td)
@@
 
-%pacckage enables computation of any activity data
+%package enables computation of any activity data
 The \pkg{digiRhythm} package enables the calculation of rhythmicity from any activity data.
@@
 Thus, it would be interesting to evaluate the application of the \pkg{digiRhythm} package in humans, offering an objective indicator of human welfare.
-A previous review highlights the potential of such rhythmicity computations in humans (https://journals.sagepub.com/doi/pdf/10.1177/0748730420940483).
+A previous review highlights the potential of such rhythmicity computations in humans \citep{dijk2020novel}.
 
-At the time of writing this paper, the current version (v1.1) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
+At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
 However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
@@
 By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
-
-<<>>=
-
-@
 

### 2023-06-02 jstatsoft/runs.R

@@ -2,8 +2,8 @@
 library("devtools")
 library("tools")
 
-#install_github('https://github.com/nasserdr/digiRhythm')
 setwd("~/projects/digiRhythm/jstatsoft")
+#install_github('https://github.com/nasserdr/digiRhythm')
 Sweave('article.Rnw')
 texi2pdf("article.tex")
 

### 2023-06-06 jstatsoft/article.Rnw
only little wording changes

@@
 \cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding of animals.
-The DFC provides an estimate of synchronization between an individual and the 24-hour day of the environment.
+The DFC provides an estimate of synchronization between an organism and the 24-hour day of the environment.
 It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}.
-\cite{sarout2018assessment} and \cite{fuchs2022detecting} suggested to use rolling seven day periods for this computation.
-An organism with a high DFC is highly synchronized with the 24-hour day of its environment, while lower DFC's indicate a lower synchronization.
+\cite{sarout2018assessment} and \cite{fuchs2022detecting} suggested to use sliding seven day periods for this computation.
+An organism with a high DFC is highly synchronized with the 24-hour day of its environment, while lower DFC indicate a lower synchronization.
 Breakdowns in the DFC response can be caused by stressors (e.g.
 handling, disease, etc.) that affect the time structure of behaviour \cite{Berger2003, scheibe1999comparative}.
-In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i.
+In fact, animals that live in stress-free conditions show activities all tuned to one Zeitgeber (i.
 e.
-24 hour periodicity), while their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
+24 hour periodicity), while their activity rhythm will deviate from the Zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
 A better understanding of changes in behaviour due to illness or distress is valuable to get further inisight in the organisms affective state % checke reference

### 2023-06-06 jstatsoft/article.Rnw
new proposal for the discussion

@@
 
+\section{Discussion (New proposal Marie)}
+
+The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of rhythmicity, animal behavior, and beyond.
+The package offers a versatile approach to assessing rhythmicity based on digital activity records.
+By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
+
+While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
+This opens up exciting avenues for future research in various domains.
+The \pkg{digiRhythm} package enables the calculation of rhythmicity from any activity data, if the data are prepared in a suitable format.It offers the propability, to compute the rhythmicity and therefore allows to gain further insight into the organism's harmonic patterns.
+As further studies showed, an organism in a good physiological state will present harmonic locomotory patterns, whereas physiological events (e.
+g.
+calving in cows), physiological or psychological diseases desynchronize these rhythmic patterns.
+However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
+Such extrinsic factors could affect rhythmic activity patterns and blur physiological rhythmicity patterns.
+To date, the package provides the option to exclude extrinsically structured time spans from the analysis.
+However, in order to increase the understanding of rhythmicity, future studies should investigate how these extrinsic factors affect rhythmicity calculations.
+
+The DFC calculation within the package includes a unique utilization of harmonic frequencies related to the 24-hour day, which provides a novel approach for rhythmicity analysis.
+However, further research is needed, to completely validate the value.
+These studies should evaluate how low collection frequency (currently 15 minutes) can be set to create reliable rhythmicity calculations.
+Additionally, the sliding seven day period, the DFC calculation depends on, and its reduction should be investigated further.
+Finally, the DFC should be compared with other metrics to calculate rhythmicity, as the Interdaily Stability (IS) and the Fourier-Based Approximation with Thresholding (FBAT).
+As well as validating the DFC, using simulated data.
+
+The package is designed with flexibility and expansion in mind.
+The current version allows for the computation of different metrics like DFC, DI, and visualization tools such as actograms.
+Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
+As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
+
+Lastly, we welcome contributions to the \pkg{digiRhythm} package.
+The open-source nature of the package fosters collaboration and continuous improvement.
+Researchers and developers are encouraged to contribute their codes, methodologies, and data to the project, thereby supporting the scientific community's collaborative spirit.
+By doing so, we hope to further our understanding of rhythmicity, its implications, and how we can leverage this knowledge for the welfare of animals and humans alike.
+
 \bibliography{refs}

### 2023-06-06 jstatsoft/Discussion_comr.R
Added new article comr

@@ -1,0 +1,26 @@
+\section{Discussion} \label{sec:summary}
+% Scope of DFC application
+The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of rhythmicity, animal behavior, and beyond. The package offers a versatile approach to assessing rhythmicity based on digital activity records. By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
+
+While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans. This opens up exciting avenues for future research in various domains. The unique utilization of harmonic frequencies related to the 24-hour day in the package provides a novel approach for rhythmicity analysis.
+
+The package is designed with flexibility and expansion in mind. The current version allows for the computation of different metrics like DFC, DI, and visualization tools such as actograms. Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future. As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
+
+%package enables computation of any activity data
+Rhythmicity can be calculated in the \pkg{digiRhythm} package from any type of regularly measured time series data, if the data are prepared in a suitable format. The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function. To date, this is purely a computation of activity patterns, resulting in the DFC. This measure has been correlated with physiological functions in previous studies. The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns. Individuals in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events. Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions. In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
+
+%Limitation of DFC computation
[…]

### 2023-06-06 jstatsoft/runs.R

@@ -3,7 +3,7 @@ library("devtools")
 library("tools")
 
 setwd("~/projects/digiRhythm/jstatsoft")
-#install_github('https://github.com/nasserdr/digiRhythm')
+install_github('https://github.com/nasserdr/digiRhythm')
 Sweave('article.Rnw')
 texi2pdf("article.tex")
 

### 2023-06-06 jstatsoft/article.Rnw
New version with Marianne/Hassan's and Marie discussions in seperate sections

@@
 \section{Discussion} \label{sec:summary}
-The \pkg{digiRhythm} package can make a significant contribution to the assessment of rhythmicity, as it makes the analysis of rhythmicity more accessible to researchers.
-The tool is easy to use, especially because it is delivered as an \proglang{R} package.
-The theoretical details and intuition behind the DFC is explained in section \ref{sec:dfc}.
-The usage of the package is explained in the section \ref{sec:usage}.
-Further examples can be found in the library's vignettes.
-
-%package enables computation of any activity data
-The \pkg{digiRhythm} package enables the calculation of rhythmicity from any activity data.
-This indicator can be calculated from any type of regularly measured time series data, if the data are prepared in a suitable format.
-The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function.
-To date, this is purely a computation of activity patterns, resulting in the DFC.
-This measure has been correlated with physiological functions in previous studies.
-
 % Scope of DFC application
-The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns.
-Individuals in a good physiological state will present harmonic locomotory patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
-Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions.
-In the past these rhythmic patterns have been discussed with regard to the psychological state of a humans, where psychological challenges, such as bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
-%Limitation of DFC computation
-
-However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
-Such extrinsic factors could affect rhythmic activity patterns and blur physiological rhythmicity patterns.
-To date, the package provides the option to exclude extrinsically structured time spans from the analysis.
-However, in order to increase the understanding of rhythmicity  future studies should investigate how these extrinsic factors affect rhythmicity calculations.
-
-The required frequency of data collection is of utmost importance in sensor technologies.
-Sensors are required to be small and further expected to run without any sensor based maintenance, such as data reading or charging of the device.
-This in turn means that data collection frequency should be low.
-In the current study a low data collection frequency is suggested (15 minutes) future studies should evaluate how low collection frequency can be set to create reliable rhythmicity calculations.
-Further it would be interesting to evaluate if activity data collected from different sources, such as  access of feeding stations or coffee maschines, compute similar rhythmicity outputs in the same organism.
-Seven days of continuous data collection are required to compute one DFC value.
-This implies that very long data collection intervals are required to detect rhythmic patterns.
-Reducing the number of days required to compute one DFC value would make this rhythmicity computation much more accessible in practice and research.
-However, the affect of reducing the intervals are not investigated, yet.
-Therefore, the impact of reducing the number of days required to compute the DFC should be investigated in future research.
-
-In addition to the open questions about the data collection frequency and intervals, a comparison to other metrics will be needed in further studies.
-In the past, other metrics for calculating circadian rhythms have been published.
-For example, Interdaily Stability (IS), which is mainly used in human research \cite{witting1990alterations}.
-And the Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
-Their database depends on the 24 h day, but none of them uses the 24 h within the calculation.
-The calculation of harmonic frequencies within the DFC calculation, which sets the activity signal in relation to the external 24h day is unique.
-Therefore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
-In addition to the mentioned metrics (IS and FBAT), \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
-Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity.
-Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
-In addition, further studies will be conducted, to validate the DFC calculation, by using simulated data.
-
-The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans.
-Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
-Many humans nowadays continuously collect activity data with their smart-watches or their phones.
-Thus, it would be interesting to evaluate the application of the \pkg{digiRhythm} package in humans, offering an objective indicator of human welfare.
-A previous review highlights the potential of such rhythmicity computations in humans \citep{dijk2020novel}.
-
-At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
-However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
-Researchers are encouraged to actively contribute by submitting new data, codes, or methods to the GitHub repository.
-These contributions will be reviewed by the package maintainers and, if approved, integrated into the \pkg{digiRhythm} package.
-
-\section{Discussion (New proposal Hassan)}
-
-The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of Rhythmicity, animal behavior, and beyond.
+The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of rhythmicity, animal behavior, and beyond.
 The package offers a versatile approach to assessing rhythmicity based on digital activity records.
@@
 As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
+
+%package enables computation of any activity data
+Rhythmicity can be calculated in the \pkg{digiRhythm} package from any type of regularly measured time series data, if the data are prepared in a suitable format.
+The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function.
+To date, this is purely a computation of activity patterns, resulting in the DFC.
+This measure has been correlated with physiological functions in previous studies.
+The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns.
+Individuals in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
+Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions.
+In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
+
+%Limitation of DFC computation
+However, there is no evidence to whether these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
+Such extrinsic factors could affect rhythmic activity patterns and blur the organisms intrinsic rhythmicity.
+To date, the package provides the option to exclude extrinsically structured time spans from the analysis.
+However, in order to increase the understanding of rhythmicity  future studies should investigate how these extrinsic factors affect rhythmicity calculations.
+Future research should also address data collection frequencies.
+Frequency of data collection is of great importance in sensor technology, because sensors are required to be small and expected to run without any sensor based maintenance, such as data reading or charging.
+This in turn implies that data collection frequencies should be set to a minimum.
+Although the current package used low data collection frequency in its example data (15 minutes), this time was given by the preset of the sensor data evaluated with this package and has no scientific basis.
+Future studies should therefore evaluate how low data collection frequencies can be set to compute reliable rhythmicity indicators.
+
+Additionally, seven days of continuous data collection are currently required to compute one DFC value, implying that long data collection intervals are necessary to detect rhythmic patterns.
+Hereby, the basis for choosing a 7 day window is unclear.
+Reducing the number of days required to compute one DFC value would make this rhythmicity computation more accessible.
+Yet, the effect of reducing the intervals has not been investigated.
+Therefore, we encourage future studies to investigate the impact of reducing the number of days required to compute the DFC.
+
+In addition to the open questions about the data collection frequency and intervals, a comparison to other metrics will be needed in further studies.
+In the past, other metrics for calculating circadian rhythms have been published.
+For example, Interdaily Stability (IS), which is mainly used in human research \cite{witting1990alterations}.
+And the Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
+Their database depends on the 24 h day, but none of them use the 24 h within the calculation.
+The calculation of harmonic frequencies within the DFC calculation, which sets the activity signal in relation to the external 24h day is unique.
+Therefore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
+In addition to the mentioned metrics (IS and FBAT), \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
+Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity.
+Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
+Further, research should evaluate if activity data collected from different sources, such as tri-axial activity data, access of feeding stations or coffee machines, compute similar rhythmicity outputs in the same organism.
+In addition, further studies usinf simulated data should be conducted, to validate the DFC calculation.
+
+The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans.
+Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
+Many humans nowadays continuously collect activity data with their smart-watches or their phones.
+A previous review highlighted the potential of such rhythmicity computations in humans \citep{dijk2020novel}.
+Thus, it would be interesting to implement the \pkg{digiRhythm} package in this domain.
+F
+
+At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
+However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
 

### 2023-06-06 jstatsoft/Discussion_comr.R

@@ -1,26 +1,0 @@
-\section{Discussion} \label{sec:summary}
-% Scope of DFC application
-The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of rhythmicity, animal behavior, and beyond. The package offers a versatile approach to assessing rhythmicity based on digital activity records. By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
-
-While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans. This opens up exciting avenues for future research in various domains. The unique utilization of harmonic frequencies related to the 24-hour day in the package provides a novel approach for rhythmicity analysis.
-
-The package is designed with flexibility and expansion in mind. The current version allows for the computation of different metrics like DFC, DI, and visualization tools such as actograms. Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future. As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
-
-%package enables computation of any activity data
-Rhythmicity can be calculated in the \pkg{digiRhythm} package from any type of regularly measured time series data, if the data are prepared in a suitable format. The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function. To date, this is purely a computation of activity patterns, resulting in the DFC. This measure has been correlated with physiological functions in previous studies. The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns. Individuals in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events. Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions. In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
-
-%Limitation of DFC computation
[…]

### 2023-06-09 jstatsoft/article.Rnw
New version with Marianne and Marie's discussion

@@
 \author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
-   \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institut für \"Okologischen Landbau
+   \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institute of Organic Farming
    \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group}
@@
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE} %For some reason, this gives an error!
@@
 By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
-
 While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
@@
 The unique utilization of harmonic frequencies related to the 24-hour day in the package provides a novel approach for rhythmicity analysis.
-
 The package is designed with flexibility and expansion in mind.
@@
 As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
-
-%package enables computation of any activity data
 Rhythmicity can be calculated in the \pkg{digiRhythm} package from any type of regularly measured time series data, if the data are prepared in a suitable format.
 The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function.
-To date, this is purely a computation of activity patterns, resulting in the DFC.
-This measure has been correlated with physiological functions in previous studies.
+At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
+However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
+
+To date, this is a computation of activity patterns, resulting in the DFC.
+This measure has been correlated with physiological functions in previous studies (\cite{Berger, Saraout}).
 The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns.
-Individuals in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
-Simply spoken, an organism will follow the same patterns of locomotion as long as they are in a good physiological and psychological state, whereas these patterns will deviate in the face of dysfunctions.
+Organisms in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
 In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
-
-%Limitation of DFC computation
-However, there is no evidence to whether these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
+However, there is no evidence to whether these activity patterns are intrinsically or extrinsically motivated.
 Such extrinsic factors could affect rhythmic activity patterns and blur the organisms intrinsic rhythmicity.
@@
 However, in order to increase the understanding of rhythmicity  future studies should investigate how these extrinsic factors affect rhythmicity calculations.
-Future research should also address data collection frequencies.
-Frequency of data collection is of great importance in sensor technology, because sensors are required to be small and expected to run without any sensor based maintenance, such as data reading or charging.
-This in turn implies that data collection frequencies should be set to a minimum.
-Although the current package used low data collection frequency in its example data (15 minutes), this time was given by the preset of the sensor data evaluated with this package and has no scientific basis.
-Future studies should therefore evaluate how low data collection frequencies can be set to compute reliable rhythmicity indicators.
 
-Additionally, seven days of continuous data collection are currently required to compute one DFC value, implying that long data collection intervals are necessary to detect rhythmic patterns.
-Hereby, the basis for choosing a 7 day window is unclear.
-Reducing the number of days required to compute one DFC value would make this rhythmicity computation more accessible.
-Yet, the effect of reducing the intervals has not been investigated.
-Therefore, we encourage future studies to investigate the impact of reducing the number of days required to compute the DFC.
-
+The DFC calculation within the package includes a unique utilization of harmonic frequencies related to the 24-hour day, which provides a novel approach for rhythmicity analysis.
+However, further research is needed, to completely validate the value.
+These studies should evaluate how low collection frequency (currently 15 minutes) can be set to create reliable rhythmicity calculations.
+Additionally, the sliding seven day period, the DFC calculation depends on, and its reduction should be investigated further.
 In addition to the open questions about the data collection frequency and intervals, a comparison to other metrics will be needed in further studies.
@@
 Further, research should evaluate if activity data collected from different sources, such as tri-axial activity data, access of feeding stations or coffee machines, compute similar rhythmicity outputs in the same organism.
-In addition, further studies usinf simulated data should be conducted, to validate the DFC calculation.
+In addition, further studies using simulated data should be conducted, to validate the DFC calculation.
 
@@
 Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
-Many humans nowadays continuously collect activity data with their smart-watches or their phones.
 A previous review highlighted the potential of such rhythmicity computations in humans \citep{dijk2020novel}.
 Thus, it would be interesting to implement the \pkg{digiRhythm} package in this domain.
-F
-
-At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
-However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
-
-Lastly, we welcome contributions to the \pkg{digiRhythm} package.
-The open-source nature of the package fosters collaboration and continuous improvement.
-Researchers and developers are encouraged to contribute their codes, methodologies, and data to the project, thereby supporting the scientific community's collaborative spirit.
-By doing so, we hope to further our understanding of rhythmicity, its implications, and how we can leverage this knowledge for the welfare of animals and humans alike.
-
-\section{Discussion (New proposal Marie)}
-
-The \pkg{digiRhythm} package presents a novel, user-friendly, and accessible tool for researchers in the field of rhythmicity, animal behavior, and beyond.
-The package offers a versatile approach to assessing rhythmicity based on digital activity records.
-By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
-
-While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
-This opens up exciting avenues for future research in various domains.
-The \pkg{digiRhythm} package enables the calculation of rhythmicity from any activity data, if the data are prepared in a suitable format.It offers the propability, to compute the rhythmicity and therefore allows to gain further insight into the organism's harmonic patterns.
-As further studies showed, an organism in a good physiological state will present harmonic locomotory patterns, whereas physiological events (e.
-g.
-calving in cows), physiological or psychological diseases desynchronize these rhythmic patterns.
-However, we have no evidence if these activity patterns are intrinsically (such as going to sleep) or extrinsically (such as waking up to an alarm) motivated.
-Such extrinsic factors could affect rhythmic activity patterns and blur physiological rhythmicity patterns.
-To date, the package provides the option to exclude extrinsically structured time spans from the analysis.
-However, in order to increase the understanding of rhythmicity, future studies should investigate how these extrinsic factors affect rhythmicity calculations.
-
-The DFC calculation within the package includes a unique utilization of harmonic frequencies related to the 24-hour day, which provides a novel approach for rhythmicity analysis.
-However, further research is needed, to completely validate the value.
-These studies should evaluate how low collection frequency (currently 15 minutes) can be set to create reliable rhythmicity calculations.
-Additionally, the sliding seven day period, the DFC calculation depends on, and its reduction should be investigated further.
-Finally, the DFC should be compared with other metrics to calculate rhythmicity, as the Interdaily Stability (IS) and the Fourier-Based Approximation with Thresholding (FBAT).
-As well as validating the DFC, using simulated data.
-
-The package is designed with flexibility and expansion in mind.
-The current version allows for the computation of different metrics like DFC, DI, and visualization tools such as actograms.
-Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
-As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
 

### 2023-06-14 jstatsoft/article.Rnw
Discussion done

@@
 While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
-This opens up exciting avenues for future research in various domains.
+This opens up new opportunities for future research in various domains.
 The unique utilization of harmonic frequencies related to the 24-hour day in the package provides a novel approach for rhythmicity analysis.
 The package is designed with flexibility and expansion in mind.
-The current version allows for the computation of different metrics like DFC, DI, and visualization tools such as actograms.
+The current version allows for the computation of different metrics, such as DFC, DI, and visualization tools such as actograms.
 Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
@@
 
-To date, this is a computation of activity patterns, resulting in the DFC.
-This measure has been correlated with physiological functions in previous studies (\cite{Berger, Saraout}).
-The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns.
+The package calculates the DFC, by computing harmonic frequencies of activity patterns.
+This measure has been correlated with physiological functions in previous studies (\cite{Berger2003, sarout2018assessment}).
+The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns (\cite{trigonum1993circadian}).
 Organisms in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
-In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the}).
-However, there is no evidence to whether these activity patterns are intrinsically or extrinsically motivated.
+\cite{rufener2018finding} found that individual laying hens followed a consistent individual pattern, but this pattern would differ between individuals within the flock.
+(\cite{opperhuizen2016feeding} further showed that altered feeding schedules during resting phases cause changes in physiology, as well as desynchronization of liver and muscle rhythms in rats.
+In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression, as well as delirium development have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the, ANGELESCASTELLANOS2016285}).
+However, the effect of intrinsic or extrinsic motivation of these activity patterns is not yet fully understood.
 Such extrinsic factors could affect rhythmic activity patterns and blur the organisms intrinsic rhythmicity.
@@
 A previous review highlighted the potential of such rhythmicity computations in humans \citep{dijk2020novel}.
+\citep{sani2015daily} were able to distinguish differences in 24 hrs activity patterns between humans in different sociocultural backrounds.
 Thus, it would be interesting to implement the \pkg{digiRhythm} package in this domain.

### 2023-06-14 jstatsoft/article.Rnw
small changes in the complete text

@@
 \author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
-   \AND Marie Schneider~\orcidlink{0000-0003-4921-8353}\\Johann Heinrich von Th\"unen\\ Institute of Organic Farming
+   \AND Marie Schneider~\orcidlink{0000-0003-1073-7754}\\Johann Heinrich von Th\"unen\\ Institute of Organic Farming
    \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group}
@@
   Switzerland\\
-  E-mail: \email{roland.nasser@agroscope.admin.ch, joa.stachowi@gmail.com}\\
-  URL: \url{https://ch.linkedin.com/in/nasserha/, https://orcid.org/0000-0003-4921-8353}\\
+  E-mail: \email{roland.nasser@agroscope.admin.ch}\\
+  URL: \url{https://ch.linkedin.com/in/nasserha/, https://orcid.org/0000-0003-1821-3234}\\
   \newline
@@
   E-mail: \email{marie.schneider@thuenen.de}\\
+  URL: \url{https://orcid.org/0000-0003-1073-7754}
   \newline
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE} %For some reason, this gives an error!
@@
 
-\item \fct{sliding\_DI}: With this function, users can compute and graphically present the diurnality index, using a specific definitions of day/night per day.
+\item \fct{sliding\_DI}: With this function, users can compute and graphically present the diurnality index, using specific definitions of day/night per day.
 
@@
 For instance, if the dataset includes the number of steps recorded at 1-minute intervals, resampling to 15 minutes entails summing the steps taken during each 15-minute interval.
-Afterward, an activity is selected (in this example, we select the activity from the second column, i.e., motion index), and the DFC is calculated using the \fct{dfc} function as illustrated below:
+It is additionally possible to convert the timezone, if needed (not in this excample).
+Afterwards, an activity is selected (in this example, we select the activity from the second column, i.e., motion index), and the DFC is calculated using the \fct{dfc} function as illustrated below:
 
@@
 
-Although the \fct{dfc} function can accommodate data with any sampling frequency, several authors \citep{sinz1976systemanalyse, sarout2018assessment, fuchs2022detecting, scheibe1999comparative} have suggested that optimal insights can be achieved using a seven-day sliding window with a 15-minute sampling period.
+Although the \fct{dfc} function can accommodate data with any sampling frequency, several authors \citep{sinz1976systemanalyse, sarout2018assessment, fuchs2022detecting} have suggested that optimal insights can be achieved using a seven-day sliding window with a 15-minute sampling period.
 To maintain simplicity, the current version does not offer an option to alter the length of the sliding window (e.g., to 6, 8, or more days).
@@
 The number of samples is typically dependent on the sampling frequency.
-For instance, if activity is measured at 15-minute intervals, and daytime is defined as the period between 06:00 AM and 04:00 PM, then $T_d$ would be calculated as 4 samples per hour × 10 hours = 40 samples.
+For instance, if activity is measured at 15-minute intervals, and daytime is defined as the period between 06:30 AM and 04:30 PM, then $T_d$ would be calculated as 4 samples per hour × 10 hours = 40 samples.
 
@@
 Note that the \fct{sliding\_DI} function would compute the DI for each day in the dataset based on the specified day/night periods, of that day.
-This could provide a more accurate reflection of the individual's behavior, as it accounts for changes in daylight hours throughout the year.
+This could provide a more accurate reflection of the individual's behavior, as it accounts for changes daily routines, that externally control the organisms activity or changes in daylight hours throughout the year.
 

### 2023-06-15 jstatsoft/article.Rnw
Changes of Marianne.

--- 2023-06-14 jstatsoft/article.Rnw
@@
 24 hour periodicity), while their activity rhythm will deviate from the Zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
-A better understanding of changes in behaviour due to illness or distress is valuable to get further inisight in the organisms affective state % checke reference
-(Grandin et al., 1989; Kilgore et al., 2005; Borderas et al., 2008; Robért et al., 2011).
+A better understanding of changes in behaviour due to illness or distress is valuable to get further insight in the organisms affective state (GRANDIN19891; Robert2011Determination).
 Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).

### 2023-06-15 jstatsoft/runs.R

@@ -3,7 +3,7 @@ library("devtools")
 library("tools")
 
 setwd("~/projects/digiRhythm/jstatsoft")
-install_github('https://github.com/nasserdr/digiRhythm')
+#install_github('https://github.com/nasserdr/digiRhythm')
 Sweave('article.Rnw')
 texi2pdf("article.tex")
 

### 2023-07-03 jstatsoft/article.Rnw
Fixed the sweaveopts
housekeepng

@@
 \author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
+   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group
       \AND Marie Schneider~\orcidlink{0000-0003-1073-7754}\\Johann Heinrich von Th\"unen\\ Institute of Organic Farming
-   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group}
-\Plainauthor{Hassan-Roland Nasser, Marie Schneider, Marianne Cockburn}
+}
+\Plainauthor{Hassan-Roland Nasser, Marianne Cockburn, Marie Schneider}
 
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE} %For some reason, this gives an error!

### 2023-07-03 jstatsoft/article.Rnw
Proofreading

@@
 %% - \AND starts a new line, \And does not.
-\author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope - Digital Production
-   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope - Equine Research Group
+\author{Hassan-Roland Nasser~\orcidlink{0000-0003-1821-3234}\\Agroscope\\Digital Production Group
+   \And Marianne Cockburn~\orcidlink{0000-0003-3193-0641}\\Agroscope\\Equine Research Group
    \AND Marie Schneider~\orcidlink{0000-0003-1073-7754}\\Johann Heinrich von Th\"unen\\ Institute of Organic Farming
 }
-\Plainauthor{Hassan-Roland Nasser, Marianne Cockburn, Marie Schneider}
+\Plainauthor{Hassan-Roland Nasser, Marie Schneider, Marianne Cockburn}
 
@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{\pkg{digiRhythm}: an \proglang{R} package for rhythmicity assessment using the degree of functional coupling}
-\Plaintitle{digiRhythm: an R package for rhythmicity assessment using the degree of functional coupling}
-\Shorttitle{digiRhythm: an R package for rhythmicity assessment using the DFC}
+\title{\pkg{digiRhythm}: An \proglang{R} Package for Rhythmicity Assessment Using the Degree of Functional Coupling}
+\Plaintitle{digiRhythm: an R Package for Rhythmicity Assessment Using the Degree of Functional Coupling}
+\Shorttitle{digiRhythm: an R Package for Rhythmicity Assessment Using the DFC}
 
@@
 What sets this measure apart is its applicability to various types of activity-related time series data, including accelerometers, GPS tracks, and smartwatches (e.g., step counts), among others.
-The Degree of Functional Coupling (DFC) represents a method for assessing rhythmicity, encompassing several computational steps like frequency spectrum calculation and the determination of the statistical significance of spectral peaks.
+The degree of functional coupling (DFC) represents a method for assessing rhythmicity, encompassing several computational steps such as frequency spectrum calculation and the determination of the statistical significance of spectral peaks.
 In this study, we introduce the \proglang{R} package \pkg{digiRhythm}, furnished with functions to compute rhythmicity from time series data.
-Complementary functions for actogram generation, computation and visualization of mean activities, and the calculation of a Diurnality Index, are also incorporated into this package.
+Complementary functions for actogram generation, computation and visualization of mean activities, and the calculation of a diurnality index, are also incorporated into this package.
 Tailored for seamless integration into the research workflow, \pkg{digiRhythm} facilitates the transformation of raw data into publication-ready content, thereby simplifying the application of these complex methods.}
@@
 %% - Should be comma-separated and in sentence case.
-\Keywords{Lomb-Scargle periodogram, Degree of Functional Coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
-\Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
+\Keywords{Lomb-Scargle periodogram, degree of functional coupling, periodicity, circadian rhythm, harmonic frequencies, actogram, diurnality index, harmonic power}
+\Plainkeywords{Lomb-Scargle periodogram, Degree of functional coupling, periodicity, circadian rhythm, harmonic frequency, actogram, diurnality index, harmonic power}
 
@@
 
+ Marianne Cockburn\\
+  Agroscope, Equine Research Group \\
+  Swiss National Stud \\
+  Les Longs-Prés\\
+  1580, Avenches \\
+  Switzerland\\
+  E-mail: \email{marianne.cockburn@agroscope.admin.ch}\\
+  URL: \url{https://orcid.org/0000-0003-3193-0641}\\
+  \newline
+
   Marie Schneider\\
@@
 
- Marianne Cockburn\\
-  Agroscope, Equine Research Group \\
-  Swiss National Stud \\
-  Les Longs-Prés\\
-  1580, Avenches \\
-  Switzerland\\
-  E-mail: \email{marianne.cockburn@agroscope.admin.ch}\\
-  URL: \url{https://orcid.org/0000-0003-3193-0641}\\
-  \newline
-
 }
@@
 
-Internal and external stimuli arouse adaptive behavioural patterns that can be monitored by analysing the rhythmicity of behaviours \citep{Berger2003}.
-Due to the increasing development of ICT (Information and Communication Technology) based on small, non-invasive sensors, large amounts of behavioural data are regularly recorded, and can be used to calculate parameters describing biological rhythms \citep{theurer2013remote,scheibe2006application, Berger2003}.
-\cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioural patterns, the Degree of Functional Coupling (DFC), which is based on telemetric measurements of activity and/or feeding of animals.
+Internal and external stimuli arouse adaptive behavioral patterns that can be monitored by analyzing the rhythmicity of behaviors \citep{Berger2003}.
+Owing to the increasing development of information and communication technology based on small, non-invasive sensors, large amounts of behavioral data are regularly recorded, and can be used to calculate parameters describing biological rhythms \citep{theurer2013remote,scheibe2006application, Berger2003}.
+\cite{sinz1976systemanalyse} developed a parameter able to quantify and investigate the rhythmicity of behavioral patterns, the degree of functional coupling (DFC), which is based on telemetric measurements of activity or feeding or both in animals.
 The DFC provides an estimate of synchronization between an organism and the 24-hour day of the environment.
-It is calculated to each 24 h period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}.
-\cite{sarout2018assessment} and \cite{fuchs2022detecting} suggested to use sliding seven day periods for this computation.
-An organism with a high DFC is highly synchronized with the 24-hour day of its environment, while lower DFC indicate a lower synchronization.
-Breakdowns in the DFC response can be caused by stressors (e.g.
-handling, disease, etc.) that affect the time structure of behaviour \cite{Berger2003, scheibe1999comparative}.
-In fact, animals that live in stress-free conditions show activities all tuned to one Zeitgeber (i.
-e.
-24 hour periodicity), while their activity rhythm will deviate from the Zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
-A better understanding of changes in behaviour due to illness or distress is valuable to get further insight in the organisms affective state (GRANDIN19891; Robert2011Determination).
-Bio-rhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).
+It is calculated to each 24-hour period within a timespan of a few days and predicts values between $0\%$ and 100 $\%$ \citep{scheibe1999comparative, Berger2003}.
+\cite{sarout2018assessment} and \cite{fuchs2022detecting} suggested using sliding 7-day periods for this computation.
+An organism with a high DFC is highly synchronized with the 24-hour day of its environment, whereas a low DFC indicates a low synchronization.
+Breakdowns in the DFC response can be caused by stressors (e.g., handling, disease, etc.) that affect the time structure of behavior \citep{Berger2003, scheibe1999comparative}.
+In fact, animals that live in stress-free conditions show activities all tuned to one zeitgeber (i.e., 24-hour periodicity), wheareas their activity rhythm will deviate from the zeitgeber frequency if they are subjected to stressful situations \citep{scheibe1999comparative, Berger2003}.
+A better understanding of changes in behaviour due to illness or distress is valuable to get further insight into the organism's affective state \citep{grandin19891, Robert2011Determination}.
+Biorhythmic analysis has the potential to provide the basis for the early detection of psychological and physiological disorders (Berger et al., 1999).
 With the increasing availability of sensor technologies to collect behavioral data continuously, further studies applying the DFC were performed \citep{sarout2018assessment, fuchs2022detecting}.
@@
 An open source package specifically for this type of computation opens up the opportunities for livestock research to easily apply these algorithms.
-Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}, which was originally developed for livestock research, but has the potential to be applied in various domains.
+Therefore, we developed the \proglang{R} package \pkg{digiRhythm} \citep{dgm}, which was originally developed for livestock research but has the potential to be applied in various domains.
 
@@
 The computation of the DFC necessitates several steps.
-Initially, the Lomb-Scargle Periodogram \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed.
-Following this, the spectrum is analyzed, focusing on specific frequencies (including both harmonics and non-harmonics, with further details in Figure \ref{fig:reallsp}).
+Initially, the Lomb-Scargle periodogram (LSP) \citep{lomb1976least, scargle1982studies, vanderplas2018understanding} of the activity is computed.
+Next, the spectrum is analyzed, focusing on specific frequencies (including both harmonics and non-harmonics, with further details in Figure \ref{fig:reallsp}).
 Subsequently, the statistical significance of frequency peaks is calculated (\ref{sub:significance}).
 Ultimately, a ratio of the sum of frequency powers is determined (a detailed breakdown of this calculation is illustrated in Figure \ref{dfc_sum}).
-We commence by elucidating the ratio of frequency powers and then provide a detailed explanation on the computation of these frequency powers and their statistical significance.
+We commence by elucidating the ratio of frequency powers and then provide a detailed explanation of the computation of these frequency powers and their statistical significance.
 
@@
 %
-In parallel, another indicator termed 'harmonic power' can be calculated similarly:
+In parallel, another indicator termed harmonic power (HP) can be calculated similarly:
 %
@@
 The \pkg{digiRhythm} package facilitates the computation of both these indicators.
-A further study showed, that the DFC reduces during an adaption process, whereas HP reduces for a longer time during diseases.
+A further study showed that the DFC decreases during an adaption process, whereas HP decreases for a longer time during diseases \cite{Berger2003}.
 
-\subsection{The Lomb-Scargle Periodogram (LSP)}\label{sub:lsp}
-The Lomb-Scargle Periodogram (LSP) marks the initial step in calculating the DFC and HP.
+\subsection{The Lomb-Scargle periodogram (LSP)}\label{sub:lsp}
+The Lomb-Scargle periodogram (LSP) marks the initial step in calculating the DFC and HP.
 It is a well-regarded tool within the astronomy community, commonly used for detecting and characterizing periodicity in unevenly sampled time-series data \citep{vanderplas2018understanding}.
@@
 \includegraphics[width=1\columnwidth]{signalVslsp}
-\caption{Exemplary harmonics (left) and their corresponding Lomb-Scargle Periodogram (LSP) (right).
-The first two rows depict regular sine waves activity signals with 24 hours and 6 hours periods, respectively, and their corresponding LSP.
-The bottom row illustrates the LSP (right) of two sine waves (of 24 hours period and 6 hours period) summed (left) together.}
+\caption{Exemplary harmonics (left) and their corresponding Lomb-Scargle periodogram (LSP) (right).
+The first two rows depict regular sine waves activity signals with 24-hour and 6-hour periods, respectively, and their corresponding LSP.
+The bottom row illustrates the LSP (right) of two sine waves (of 24-hour period and 6-hour period) summed (left) together.}
 \label{fig:signalVslsp}
@@
 Activity signals from individuals tend to be irregular, requiring conversion via the LSP to investigate periodic events within these signals.
-Figure \ref{fig:realsignal} illustrates an example of such data - the motion index (a proprietary metric of the overall animal activity measured in three dimensions \cite{icetagiowa}) of a cow, and its corresponding LSP is depicted in Figure \ref{fig:reallsp}.
+Figure \ref{fig:realsignal} illustrates an example of such data - the motion index (a proprietary metric of the overall animal activity measured in three dimensions \citep{icetagiowa} of a cow, and its corresponding LSP is depicted in Figure \ref{fig:reallsp}.
 
@@
 @
-\caption{\label{fig:realsignal} The \emph{Motion Index} activity column in the dataset \fct{df516b\_2}.}
+\caption{\label{fig:realsignal} The \emph{motion index} activity column in the dataset \fct{df516b\_2}.}
 \end{figure}
@@
 This frequency grid ranges from 0 to $f_{max}$, the maximum frequency power that can be computed from the activity $g_n$.
-According to the Shannon-Nyquist theorem, $f_{max}$ is determined \citep{shannon1928certain, shannon1949mathematical}.
+According to the Nyquist-Shannon theorem, $f_{max}$ is determined \citep{shannon1928certain, shannon1949mathematical}.
 
@@
 @
-\caption{The Lomb-Scargle Periodogram of the activity data shown in Figure \ref{fig:realsignal}.
+\caption{The Lomb-Scargle periodogram of the activity data shown in Figure \ref{fig:realsignal}.
 Red bars show the amplitude of harmonic frequencies (with corresponding harmonic periods).
-While green bars represent non-harmonic frequencies.
+Green bars represent non-harmonic frequencies.
 The dashed line indicates the significance threshold; frequencies with power below this threshold are considered non-significant.
-In this particular example, only two frequencies with significant power are evident; one of them - corresponding to a 12-hour period - is harmonic.}\label{fig:reallsp}
+In this particular example, only two frequencies with significant power are evident; one of them - corresponding to a 12-hour period is harmonic.}\label{fig:reallsp}
 \end{figure}
 
-\subsection{Understanding Harmonic Frequencies and the Underlying Intuition}\label{sub:harmonic}
+\subsection{Understanding harmonic frequencies and the underlying intuition}\label{sub:harmonic}
 
@@
 \includegraphics[width=1\columnwidth]{sum_dfc}
-\caption{Significant and non-significant frequency peaks depicted in a Lomb-Scargle Periodogram, contextualizing the rhythmicity in animal behaviors.
+\caption{Significant and non-significant frequency peaks depicted in a Lomb-Scargle periodogram, contextualizing the rhythmicity in animal behaviors.
 The total frequency space includes frequencies with significant and non-significant peaks, the combined powers of which amount to $sumall$.
@@
 
-\subsection{Evaluating the Statistical Significance of Frequency Peaks}\label{sub:significance}
+\subsection{Evaluating the statistical significance of frequency peaks}\label{sub:significance}
 
@@
 
-Commonly, the False Alarm Probability (FAP) is employed to serve this purpose. FAP quantifies the likelihood that data devoid of any activity would generate a peak of comparable magnitude.
+Commonly, the false alarm probability (FAP) is employed to serve this purpose. FAP quantifies the likelihood that data devoid of any activity would generate a peak of comparable magnitude.
 
-As reported by Scargle \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom.
+As reported by \cite{scargle1982studies}, the values of the un-normalized periodogram follow a $\chi^2$ distribution with two degrees of freedom.
 Therefore, the probability of observing periodogram values less than a certain $Z=P(f_0)$ in a data series constituted exclusively by Gaussian noise is expressed as:
@@
 
-Multiple methodologies can be used to calculate the FAP.
-Among these, the Baluev method, according to comparative studies by \citep{suveges2015comparative}, stands as one of the most effective and hence, has been implemented in the \pkg{digiRhythm} package.
+Multiple methods can be used to calculate the FAP.
+Among these, the Baluev method, according to comparative studies by \cite{suveges2015comparative}, stands as one of the most effective and hence, has been implemented in the \pkg{digiRhythm} package.
 
-As per Baluev \citep{baluev2008assessing}, the FAP can be formalized as follows:
+As per \citep{baluev2008assessing}, the FAP can be formalized as follows:
 
@@
 
-\section{Functionalities and Usage of digiRhythm}\label{sec:usage}
+\section{Functionalities and usage of digiRhythm}\label{sec:usage}
 
-The \pkg{digiRhythm} package offers a range of functionalities beyond merely calculating the DFC and Harmonic Part (HP).
+The \pkg{digiRhythm} package offers a range of functionalities beyond merely calculating the DFC and harmonic part (HP).
 In collaboration with animal behavior scientists, we have incorporated a variety of tools typically required for rhythmicity analysis.
@@
 \begin{itemize}
-\item \fct{lomb\_scargle\_periodogram}: This function not only computes the Lomb Scargle Periodogram (LSP) but also provides an intricate visualization of it.
+\item \fct{lomb\_scargle\_periodogram}: This function not only computes the Lomb-Scargle periodogram (LSP) but also provides an intricate visualization of it.
 
-\item \fct{dfc}: This function allows for both computation and graphical representation of the Degree of Functional Coupling and its harmonic part.
+\item \fct{dfc}: This function allows for both computation and graphical representation of the DFC and the HP.
 
@@
 
-\item \fct{sliding\_DI}: With this function, users can compute and graphically present the diurnality index, using specific definitions of day/night per day.
+\item \fct{sliding\_DI}: With this function, users can compute and graphically present the diurnality index, using specific definitions of daytime and nightime per day.
 
@@
 
-The \fct{dfc} function acts as a wrapper for the \fct{lomb\_scargle\_periodogram}, providing users with the opportunity to delve further into the Lomb-Scargle Periodogram, should they wish to explore it in greater detail.
+The \fct{dfc} function acts as a wrapper for the \fct{lomb\_scargle\_periodogram}, providing users with the opportunity to delve further into the LSP, should they wish to explore it in greater detail.
 
-\subsection{Format of Input Data}
+\subsection{Format of input data}
 
-As referenced in Section \ref{sec:intro}, the \pkg{digiRhythm} package necessitates that input data frames adhere to a specific structure.
+As referenced in section \ref{sec:intro}, the \pkg{digiRhythm} package necessitates that input data frames adhere to a specific structure.
 An example of a typically structured data frame, sampled at 15-minute intervals, is displayed below.
@@
 
-To confirm that the data adheres to the required format, we've included a utility function capable of verifying its compatibility with \pkg{digiRhythm}.
+To confirm that the data adheres to the required format, we have included a utility function capable of verifying its compatibility with \pkg{digiRhythm}.
 This check can be performed using the \fct{is\_dgm\_friendly} function, as demonstrated below:
@@
 
-This function takes a data frame as an argument and verifies its \pkg{digiRhythm} compatibility, returning a boolean.
+This function takes a data frame as an argument and verifies its \pkg{digiRhythm} compatibility, returning a Boolean.
 The data is considered \pkg{digiRhythm}-friendly if it satisfies the following conditions:
@@
 
-The verbose argument aids in understanding why the data is or isn't considered \pkg{digiRhythm}-friendly. If verbose = TRUE, the function will provide information about each condition checked.
+The verbose argument aids in understanding why the data is or is not considered \pkg{digiRhythm}-friendly. If verbose = TRUE, the function will provide information about each condition checked.
 
-A warning message will be generated if the data encompasses less than seven days, since at the moment, at least a week's worth of data is needed to compute the DFC and HP.
+A warning message will be generated if the data encompasses less than 7 days because at the moment, at least a week's worth of data is needed to compute the DFC and HP.
 However, other functionalities may still be executable.
 
-\subsection[The dfc Function]{The \fct{dfc} Function}\label{sub:dgm_fct}
+\subsection[The dfc function]{The \fct{dfc} function}\label{sub:dgm_fct}
 
-The DFC and HP are both implemented in the \fct{dfc} function, as per Algorithm \ref{alg:dfc}.
+The DFC and HP are both implemented in the \fct{dfc} function, as per algorithm \ref{alg:dfc}.
 This function processes a \pkg{digiRhythm}-compatible data frame by calculating the DFC for the chosen variable targeted for analysis.
-Initially, it evaluates the start and end dates to ensure a sufficient number of days (a minimum of seven by default) are available for the computation.
-The function then applies a Lomb Scargle Periodogram to a seven-day sliding window.
-For instance, the DFC and HP are computed for the first seven days, then for days two to eight, and so on.
-The final six days of the dataset are excluded, as applying the LSP to periods shorter than seven days is not recommended when computing the DFC.
-A typical DFC/HP output is illustrated in Figure \ref{fig:dfc}.
+Initially, it evaluates the start and end dates to ensure a sufficient number of days (a minimum of 7 by default) are available for the computation.
+The function then applies an LSP to a seven-day sliding window.
+For instance, the DFC and HP are computed for the first 7 days, then for days 2 to 8, and so on.
+The final 6 days of the dataset are excluded, because applying the LSP to periods shorter than seven days is not recommended when computing the DFC.
+A typical combined DFC and HP output is illustrated in Figure \ref{fig:dfc}.
 
@@
 \end{algorithm}
-\subsection[Utilizing digiRhythm for DFC and HP Computation]{Utilizing \pkg{digiRhythm} for DFC and HP Computation}\label{sub:dgm_dfc_hp}
+\subsection[Utilizing digiRhythm for DFC and HP Computation]{Utilizing \pkg{digiRhythm} for DFC and HP computation}\label{sub:dgm_dfc_hp}
 
@@
 @
-\caption{\label{fig:dfc} The DFC and HP of the dataset \fct{df516b\_2}.}
+\caption{\label{fig:dfc} The degree of functional coupling and the harmonic part of the dataset \fct{df516b\_2}.}
 \end{figure}
@@
 
-Thus, it's straightforward to access the resulting DFC and HP for further analysis such as statistical modeling or other tailored workflows.
-As the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, the plot can directly be modified, by adding a layer to the aesthetics of the \code{my_dfc} object as demonstrated below:
+Thus, it is straightforward to access the resulting DFC and HP for further analysis such as statistical modeling or other tailored workflows.
+Because the returned object is a \pkg{ggplot2} \citep{wickham2010layered}, the plot can directly be modified, by adding a layer to the aesthetics of the \code{my_dfc} object as demonstrated below in the code snipet below.
 
@@
 %
-\caption{\label{fig:actogram} The actogram of the of the dataset \fct{df516b\_2}.}
+\caption{\label{fig:actogram} The actogram of the dataset \fct{df516b\_2}.}
 \end{figure}
@@
 
-The average activity provides an indication of how active an individual is on average at different times of the day. This average activity is calculated using the formula:
+The average activity provides an indication of how active an individual is on average at different times of the day. This average activity is calculated using the following formula:
 
@@
 
-\subsection[Diurnality Index (DI)]{Diurnality Index (DI)}{\label{sub:di}}
+\subsection[Diurnality index (DI)]{Diurnality index (DI)}{\label{sub:di}}
 
-The Diurnality Index (DI) is another crucial measure used to assess an individual's rhythmicity.
+The diurnality index (DI) is another crucial measure used to assess an individual's rhythmicity.
 This index provides insights into whether an individual tends to be more active during the day (diurnal) or at night (nocturnal).
@@
 
-\subsection[sliding Diurnality Index (DI)]{sliding Diurnality Index (DI)}{\label{sub:sliding_di}}
+\subsection[sliding diurnality index (DI)]{sliding diurnality index (DI)}{\label{sub:sliding_di}}
 
 The \fct{sliding\_DI} function provides a more nuanced approach to calculating the DI.
-It allows for dynamic definitions of day and night time periods across the dataset, rather than using a static, single definition.
+It allows for dynamic definitions of daytime and nighttime periods across the dataset, rather than using a static, single definition.
 This is particularly useful when dealing with data that spans across different seasons, daylight saving time changes, or different geographical locations where the length of the day varies significantly.
@@
 These timestamps must be in POSIXct format.
-An example dataset of a typically definition of daytime and nighttime, by excluding milking times in dairy farms is displayed below.
+An example dataset of a typical definition of daytime and nighttime by excluding milking times on dairy farms is displayed below.
 This data frame is bundled with the \pkg{digiRhythm} package and can be retrieved using:
@@
 %
-\caption{\label{fig:di} The Diurnality Index of the dataset \fct{df516b\_2}, calculated using the \fct{sliding\_DI} function.}
+\caption{\label{fig:di} The diurnality index of the dataset \fct{df516b\_2}, calculated using the \fct{sliding\_DI} function.}
 \end{figure}
 
-In the above example for the \fct{diurnality} function, the day starts at 6:30 AM and ends at 4:30 PM, while the night starts at 6 PM and ends at 5 AM.
-Note that the \fct{sliding\_DI} function would compute the DI for each day in the dataset based on the specified day/night periods, of that day.
-This could provide a more accurate reflection of the individual's behavior, as it accounts for changes daily routines, that externally control the organisms activity or changes in daylight hours throughout the year.
+In the above example for the \fct{diurnality} function, the day starts at 6:30 AM and ends at 4:30 PM, and the night starts at 6:00 PM and ends at 5:00 AM.
+Note that the \fct{sliding\_DI} function would compute the DI for each day in the dataset based on the specified daytime and nightime periods of that day.
+This could provide a more accurate reflection of the individual's behavior because it accounts for changes daily routines that externally control the organisms activity or changes in daylight hours throughout the year.
 
@@
 By leveraging the power of \proglang{R}, it brings robust statistical capabilities and intuitive visualization features into the hands of researchers.
-While the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
+Although the current implementation of the \pkg{digiRhythm} package focuses on rhythmicity in farm animals, the underlying principles and methodology can be extended to other organisms, including humans.
 This opens up new opportunities for future research in various domains.
@@
 The current version allows for the computation of different metrics, such as DFC, DI, and visualization tools such as actograms.
-Still, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
+Moreover, the modular structure of the package enables the incorporation of additional techniques and tools in the future.
 As the \pkg{digiRhythm} package continues to develop and evolve, we expect it to become an off-the-shelf tool for researchers interested in studying rhythmicity patterns and their implications.
@@
 The suitability of the format can be assessed with the  \fct{is\_dgm\_friendly} function.
-At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis (\cite{refinetti2016}).
+At the time of writing this paper, the current version (v1.2) of the library does not provide the capability to compute rhythmicity using techniques such as cosinor analysis \cite{refinetti2016}.
 However, there is potential for future expansion of the library's functionality to incorporate additional techniques.
@@
 This measure has been correlated with physiological functions in previous studies (\cite{Berger2003, sarout2018assessment}).
-The computation of rhythmicity allows to gain further insight into an individual's harmonic patterns (\cite{trigonum1993circadian}).
+The computation of rhythmicity allows gaining further insight into an individual's harmonic patterns \cite{trigonum1993circadian}.
 Organisms in a good physiological state will present harmonic locomotion patterns, whereas this rhythmicity desynchronizes as a result of physiological events.
 \cite{rufener2018finding} found that individual laying hens followed a consistent individual pattern, but this pattern would differ between individuals within the flock.
-(\cite{opperhuizen2016feeding} further showed that altered feeding schedules during resting phases cause changes in physiology, as well as desynchronization of liver and muscle rhythms in rats.
-In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression, as well as delirium development have been associated with the desynchronization of rhythmic patterns (\cite{sabet2021the, ANGELESCASTELLANOS2016285}).
+\cite{opperhuizen2016feeding} further showed that altered feeding schedules during resting phases cause changes in physiology, as well as desynchronization of liver and muscle rhythms in rats.
+In the past, these rhythmic patterns have been discussed with regard to the psychological state of humans, where  bipolar disorders or depression, as well as delirium development have been associated with the desynchronization of rhythmic patterns \citep{sabet2021the, ANGELESCASTELLANOS2016285}.
 However, the effect of intrinsic or extrinsic motivation of these activity patterns is not yet fully understood.
@@
 The DFC calculation within the package includes a unique utilization of harmonic frequencies related to the 24-hour day, which provides a novel approach for rhythmicity analysis.
-However, further research is needed, to completely validate the value.
-These studies should evaluate how low collection frequency (currently 15 minutes) can be set to create reliable rhythmicity calculations.
-Additionally, the sliding seven day period, the DFC calculation depends on, and its reduction should be investigated further.
-In addition to the open questions about the data collection frequency and intervals, a comparison to other metrics will be needed in further studies.
-In the past, other metrics for calculating circadian rhythms have been published.
-For example, Interdaily Stability (IS), which is mainly used in human research \cite{witting1990alterations}.
-And the Fourier-Based Approximation with Thresholding (FBAT) was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
-Their database depends on the 24 h day, but none of them use the 24 h within the calculation.
-The calculation of harmonic frequencies within the DFC calculation, which sets the activity signal in relation to the external 24h day is unique.
+However, further research is needed to completely validate the value.
+These studies should evaluate how low the sampling periods (currently 15 minutes) can be set to create reliable rhythmicity calculations.
+Additionally, the sliding 7-day period, on which DFC calculation depends, and its reduction should be investigated further.
+In addition to the open questions about the data collection frequency and intervals, a comparison with other metrics will be needed in further studies.
+In the past, other metrics for calculating circadian rhythms have been published, for example, interdaily stability, which is mainly used in human research \cite{witting1990alterations}.
+and the Fourier-based approximation with thresholding which was designed to calculate the circadian rhythms of dairy cows \cite{wagner2021detection}.
+Their databases depends on the 24-hour day, but none of them use the 24 hours within the calculation.
+The calculation of harmonic frequencies within the DFC calculation, which sets the activity signal in relation to the external 24-hour day is unique.
 Therefore, it would be interesting to investigate the correlation between the DFC and rhythmicity computed on data from various computation techniques.
-In addition to the mentioned metrics (IS and FBAT), \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
-Additionally, GPS data of free ranging animals \cite{plaza2022gps} or data of real time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity.
+In addition to the mentioned metrics (interdaily stability and Fourier-based approximation with thresholding), \cite{moller2022diurnal} utilized a range of techniques, including set logic, circular cross-correlation, circular clustering, functional enrichment analyses, and least squares regression, on transcriptomic time series data from humans and baboons.
+
+Additionally, GPS data of free-ranging animals \cite{plaza2022gps} or data of real-time positioning systems in a barn \cite{wagner2021detection} could be used to estimate rhythmicity.
 Comparing the correlations between rhythmicity measures derived from different types of data and between various techniques could be a useful research area.
-Further, research should evaluate if activity data collected from different sources, such as tri-axial activity data, access of feeding stations or coffee machines, compute similar rhythmicity outputs in the same organism.
-In addition, further studies using simulated data should be conducted, to validate the DFC calculation.
+Furtheremore, research should evaluate if activity data collected from different sources, such as tri-axial activity data, access of feeding stations, or use of coffee machines, compute similar rhythmicity outputs in the same organism.
+In addition, further studies using simulated data should be conducted to validate the DFC calculation.
 
 The current package has been developed for farm animals, but the principles of rhythmicity apply for any organism, including humans.
-Therefore we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
-A previous review highlighted the potential of such rhythmicity computations in humans \citep{dijk2020novel}.
-\citep{sani2015daily} were able to distinguish differences in 24 hrs activity patterns between humans in different sociocultural backrounds.
+Therefore, we highly recommend future research to evaluate the \pkg{digiRhythm} package in other species.
+A previous review highlighted the potential of such rhythmicity computations in humans \cite{dijk2020novel}.
+\cite{sani2015daily} were able to distinguish differences in 24-hour activity patterns between humans in different sociocultural backrounds.
 Thus, it would be interesting to implement the \pkg{digiRhythm} package in this domain.

### 2024-04-05 examples/code_snipets_manuscript.R
Fixed a glitch in the vignettes. Finalized the code for the paper.

@@ -1,21 +1,32 @@
+#' General info
+#' This code is to reproduce the results in the digiRhythm paper.
+#' The code was test with R 4.3.3
+#' For the below imported libraries, if you get an error, please install the library using the command install.packages("library_name")
+#' Before you start, click on session ->  set working directory -> to source file location
 library(ggplot2)
 library(magrittr)
 library(lubridate)
 library(dplyr)
-library(digiRhythm)
 library(lomb)
 library(patchwork)
-
 
 ###############################################################################
 ####################### Installing DigiRhythm ##################################
 
 # Install the devtools package if you haven't already
+
+# To install the development version from GitHub
 #install.packages("devtools")
-devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
+#devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
+
+# To install the stable version from CRAN
+install.packages('digiRhythm')
+library(digiRhythm)
 
 ###############################################################################
-####################### FIGURE: 2 ###################################
+################################## FIGURE: 2 ##################################
+###############################################################################
+
 # LSP on purely synthetic data ----------------------------
 #Creating time grid
 sampling = 15*60 #15 minutes
@@ -56,7 +67,7 @@ for (period in c(24, 6)){
       axis.text = element_text(color = "#000000"),
       text = element_text(size = 15),
       panel.background = element_rect(fill = "white"),
-      axis.line = element_line(size = 0.5),
+      axis.line = element_line(linewidth = 0.5),
       legend.key = element_rect(fill = "white"),
       legend.key.width = unit(0.5, "cm"),
       legend.justification ="right",
@@ -64,9 +75,6 @@ for (period in c(24, 6)){
       legend.position = c(1,0.89),
       plot.margin = margin(t = 50)) +
       scale_x_continuous(n.breaks = 8)
-
-      # axis.text  = element_blank(),
-      # axis.ticks  = element_blank())
 
   name <- paste0('figures/sig', period, '.pdf')
   ggsave(
@@ -80,7 +88,7 @@ for (period in c(24, 6)){
     limitsize = TRUE)
 
   df$num <- NULL
[…]

### 2024-07-16 README.Rmd
Prettify all the codes with the package styler

@@
 ```{r}
-#Uncomment the below two lines if you use the library for the first time of if
-#you want to update the library
-#install.packages("devtools")
-#devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
+# Uncomment the below two lines if you use the library for the first time of if
+# you want to update the library
+# install.packages("devtools")
+# devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
 ```
@@
 
-#The file name with the path
-url <- 'https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv'
-download.file(url, destfile = '516b_2.csv')
-filename <- file.path(getwd(), '516b_2.csv')
+# The file name with the path
+url <- "https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv"
+download.file(url, destfile = "516b_2.csv")
+filename <- file.path(getwd(), "516b_2.csv")
 
-#The columns that we are interested in
-colstoread <- c("Date", "Time", "Motion Index", 'Steps')
+# The columns that we are interested in
+colstoread <- c("Date", "Time", "Motion Index", "Steps")
 
-#Reading the activity data from the csv file
+# Reading the activity data from the csv file
 data <- import_raw_icetag_data(filename = filename, skipLines = , act.cols.names = colstoread, sampling = 15)
@@
 print(head(data))
-
 ```
@@
 df_act_info(df)
-activity = names(df)[2]
+activity <- names(df)[2]
 start <- "2020-08-25"
@@
 save <- TRUE
-outputdir <- 'testresults'
-outplotname <- 'myplot'
+outputdir <- "testresults"
+outplotname <- "myplot"
 width <- 10
-device <- 'tiff'
+device <- "tiff"
 height <- 5
-actogram(df, activity, start, end, save = FALSE,
-     outputdir = 'testresults', outplotname = 'actoplot', width = 10,
-     height =  5, device = 'tiff')
+actogram(df, activity, start, end,
+  save = FALSE,
+  outputdir = "testresults", outplotname = "actoplot", width = 10,
+  height = 5, device = "tiff"
+)
 ```
@@
 df_act_info(df)
-activity = names(df)[2]
-my_dfc <- dfc(df, activity , sampling = 15, show_lsp_plot = FALSE)
+activity <- names(df)[2]
+my_dfc <- dfc(df, activity, sampling = 15, show_lsp_plot = FALSE)
 
-#You may want to explore the two list inside my_dfc.
-#DFC and SPECTRUM are saved inside my_dfc, each as a list
+# You may want to explore the two list inside my_dfc.
+# DFC and SPECTRUM are saved inside my_dfc, each as a list
 ```
@@
 df_act_info(df)
-activity = names(df)[2]
+activity <- names(df)[2]
 d_index <- diurnality(data, activity, plot = TRUE)
-
 ```
@@
 new_dgm <- resample_dgm(df, new_sampling)
-
 ```

### 2024-07-16 R/actogram.R

@@ -26,24 +26,22 @@
 #' data("df516b_2")
 #' df <- df516b_2
 #' activity <- names(df)[2]
-#' start <- "2020-05-01" #year-month-day
-#' end <- "2020-08-13" #year-month-day
-#' activity_alias <- 'Motion Index'
+#' start <- "2020-05-01" # year-month-day
+#' end <- "2020-08-13" # year-month-day
+#' activity_alias <- "Motion Index"
 #' my_actogram <- actogram(df, activity, activity_alias, start, end, save = NULL)
 #' print(my_actogram)
-
 actogram <- function(
-  df,
-  activity,
-  activity_alias,
-  start,
-  end,
-  save = 'actogram'
-){
+    df,
+    activity,
+    activity_alias,
+    start,
+    end,
+    save = "actogram") {
   start <- lubridate::date(start)
   end <- lubridate::date(end)
 
-  names(df)[1] <- 'datetime'
+  names(df)[1] <- "datetime"
 
   df$date <- lubridate::date(df$datetime)
   data_to_plot <- df %>%
@@ -52,7 +50,7 @@ actogram <- function(
   data_to_plot$time <- format(data_to_plot$datetime, format = "%H:%M", tz = "CET")
   data_to_plot <- data_to_plot %>% select(-datetime)
 
-  data_to_plot$date_numeric <- xtfrm(as.Date(data_to_plot$date, format='%Y-%m-%d'))
+  data_to_plot$date_numeric <- xtfrm(as.Date(data_to_plot$date, format = "%Y-%m-%d"))
 
 
   equally_spaced_select <- function(x) {
@@ -61,43 +59,50 @@ actogram <- function(
   }
 
   if (length(unique(data_to_plot$date_numeric)) > 15) {
-    breaks_for_y_axis = equally_spaced_select(unique(data_to_plot$date_numeric))
-    labels_for_y_axis = equally_spaced_select(unique(data_to_plot$date))
+    breaks_for_y_axis <- equally_spaced_select(unique(data_to_plot$date_numeric))
+    labels_for_y_axis <- equally_spaced_select(unique(data_to_plot$date))
   } else {
-    breaks_for_y_axis = unique(data_to_plot$date_numeric)
-    labels_for_y_axis = unique(data_to_plot$date)
+    breaks_for_y_axis <- unique(data_to_plot$date_numeric)
+    labels_for_y_axis <- unique(data_to_plot$date)
   }
 
-  act_plot <- ggplot(data_to_plot,
[…]

### 2024-07-16 R/daily_acticity_wrap_plot.R

@@ -28,31 +28,31 @@
 #' data("df516b_2")
 #' df <- df516b_2
 #' activity <- names(df)[2]
-#' activity_alias <- 'Motion Index'
-#' start <- "2020-05-01" #year-month-day
-#' end <- "2020-05-07" #year-month-day
+#' activity_alias <- "Motion Index"
+#' start <- "2020-05-01" # year-month-day
+#' end <- "2020-05-07" # year-month-day
 #' ncols <- 3
 #' sampling_rate <- 30
-#' my_dwp <- daily_activity_wrap_plot(df, activity, activity_alias, start, end, sampling_rate,
-#' ncols)
-
-
+#' my_dwp <- daily_activity_wrap_plot(
+#'   df, activity, activity_alias, start, end, sampling_rate,
+#'   ncols
+#' )
 daily_activity_wrap_plot <- function(
-  df,
-  activity,
-  activity_alias,
-  start,
-  end,
-  sampling_rate,
-  ncols,
-  save = 'daily_wrap_plot'
-
-){
+    df,
+    activity,
+    activity_alias,
+    start,
+    end,
+    sampling_rate,
+    ncols,
+    save = "daily_wrap_plot") {
   start <- lubridate::date(start)
   end <- lubridate::date(end)
 
   df$date <- lubridate::date(df$datetime)
-  data_to_plot <- df %>% filter(date >= start) %>% filter(date <= end)
+  data_to_plot <- df %>%
+    filter(date >= start) %>%
+    filter(date <= end)
 
   data_to_plot$time <- format(data_to_plot$datetime, format = "%H:%M", tz = "CET")
 
@@ -64,33 +64,35 @@ daily_activity_wrap_plot <- function(
   # batch1 <- data_to_plot %>% filter(data_to_plot$date <= days[l])
   # batch2 <- data_to_plot %>% filter(data_to_plot$date > days[l])
 
-  time2 = rep(1:(24*60/sampling_rate), length(unique(data_to_plot$date)))
+  time2 <- rep(1:(24 * 60 / sampling_rate), length(unique(data_to_plot$date)))
   data_to_plot$time2 <- time2[1:dim(data_to_plot)[1]]
 
   avg_plot <- ggplot(data_to_plot, aes(
     x = time2,
-    y = !!as.name(activity)))+
+    y = !!as.name(activity)
+  )) +
     geom_area() +
     facet_wrap(~date, ncol = ncols) +
[…]

### 2024-07-16 R/daily_average_activity.R

@@ -26,22 +26,18 @@
 #' data("df516b_2")
 #' df <- df516b_2
 #' activity <- names(df)[2]
-#' start <- "2020-05-01" #year-month-day
-#' end <- "2020-08-13" #year-month-day
-#' activity_alias <- 'Motion Index'
+#' start <- "2020-05-01" # year-month-day
+#' end <- "2020-08-13" # year-month-day
+#' activity_alias <- "Motion Index"
 #' my_daa <- daily_average_activity(df, activity, activity_alias, start, end, save = NULL)
 #' print(my_daa)
-
-
 daily_average_activity <- function(
-  df,
-  activity,
-  activity_alias,
-  start,
-  end,
-  save
-){
-
+    df,
+    activity,
+    activity_alias,
+    start,
+    end,
+    save) {
   df$date <- lubridate::date(df$datetime)
   data_to_plot <- df %>%
     filter(lubridate::date(df$datetime) >= start) %>%
@@ -51,20 +47,21 @@ daily_average_activity <- function(
   start <- lubridate::date(start)
   end <- lubridate::date(end)
 
-  sum_of_activity_over_all_days_per_sample = NULL
-  sum_of_activity_over_all_days_per_sample =  data.frame(
+  sum_of_activity_over_all_days_per_sample <- NULL
+  sum_of_activity_over_all_days_per_sample <- data.frame(
     time = as.character(),
     average = as.numeric()
   )
 
-  for(t in unique(data_to_plot$time)){
+  for (t in unique(data_to_plot$time)) {
     tdf <- data_to_plot %>% filter(time == t)
-    mean = mean(tdf[[activity]])
+    mean <- mean(tdf[[activity]])
     sum_of_activity_over_all_days_per_sample <- rbind(
       sum_of_activity_over_all_days_per_sample,
       data.frame(
         time = t,
-        average = mean)
+        average = mean
+      )
     )
   }
 
@@ -75,11 +72,13 @@ daily_average_activity <- function(
 
   s <- s %>% select(datetime, average)
 
-  avg_act_plot <- ggplot(s,
-                         aes(
-                           x = datetime,
-                           y = average
-                         )) +
+  avg_act_plot <- ggplot(
+    s,
+    aes(
+      x = datetime,
+      y = average
+    )
+  ) +
     geom_line() +
     xlab("Time") +
     ylab(paste0("Daily Average of ", activity_alias)) +
@@ -87,19 +86,19 @@ daily_average_activity <- function(
     theme_classic() +
[…]

### 2024-07-16 R/dfc.R

@@ -1,4 +1,3 @@
-
 #' Computes the Degree of Function coupling (DFC), Harmonic Part (HP) and Weekly
 #' Lomb-Scargle Spectrum (LSP Spec) for one variable in an activity dataset.
 #' The dataset should be digiRhythm friendly.
@@ -52,69 +51,71 @@
 #' @export
 #' @examples
 #' data("df516b_2", package = "digiRhythm")
-#' df <- df516b_2[, c(1,2)]
+#' df <- df516b_2[, c(1, 2)]
 #' df <- remove_activity_outliers(df)
 #' df_act_info(df)
-#' activity = names(df)[2]
+#' activity <- names(df)[2]
 #' my_dfc_1 <- dfc(df, activity, sampling = 15)
 #' my_dfc_2 <- dfc(df, activity, sampling = 15, rolling_window = 4)
-
 #######################################################
 dfc <- function(
     data,
     activity,
-    sampling = 15, #in minutes
+    sampling = 15, # in minutes
     sig = 0.05,
     rolling_window = 7,
     plot = TRUE,
     plot_harmonic_part = TRUE,
     verbose = TRUE,
-    plot_lsp = TRUE
-)
-{
-
+    plot_lsp = TRUE) {
   data <- as.data.frame(data, row.names = NULL)
 
   print(head(data))
   if (!is_dgm_friendly(data)) {
-    stop('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
+    stop("The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information")
   }
 
 
   data$date <- date(data$datetime)
 
-  days <- seq(data$date[1],
-              last(data$date),
-              1)
+  days <- seq(
+    data$date[1],
+    last(data$date),
+    1
+  )
   if (length(days) < 2) {
-    stop('You need at least 2 days of data to run the Degree of Functional Coupling algorithm')
+    stop("You need at least 2 days of data to run the Degree of Functional Coupling algorithm")
   }
 
 
   if (length(days) < rolling_window) {
-    stop('The number of rolling widown should be bigger than the number of days in the dataset.')
+    stop("The number of rolling widown should be bigger than the number of days in the dataset.")
   }
 
   dfc <- NULL
   spec <- NULL
   from <- NULL
 
-  dfc <- data.frame(from = character(),
-                    to = character(),
-                    dfc = numeric(),
[…]

### 2024-07-16 R/diurnality.R

@@ -19,7 +19,7 @@
 #' data("df516b_2", package = "digiRhythm")
 #' data <- df516b_2
 #' data <- remove_activity_outliers(data)
-#' activity = names(data)[2]
+#' activity <- names(data)[2]
 #' d_index <- diurnality(data, activity)
 #'
 #' @export
@@ -28,141 +28,146 @@ diurnality <- function(data,
                        activity,
                        day_time = c("06:30:00", "16:30:00"),
                        night_time = c("18:00:00", "T05:00:00"),
-                       save = NULL){
+                       save = NULL) {
+  # di = (cd/td - cn/tn)/(cd/td + cn/tn)
 
-  #di = (cd/td - cn/tn)/(cd/td + cn/tn)
-
-  dates <- unique(lubridate::date(data[,1]))
+  dates <- unique(lubridate::date(data[, 1]))
   X <- xts(
     x = data[[activity]],
-    order.by = data[,1]
+    order.by = data[, 1]
   )
 
   sampling <- dgm_periodicity(data)[["frequency"]]
 
-  #Code Addition 1
-  #Formatting the time range of the day
+  # Code Addition 1
+  # Formatting the time range of the day
   start_day <- paste0(
-    'T',
-    sprintf("%02d",attr(hms(day_time[1]),'hour')),
-    ':',
-    sprintf("%02d",attr(hms(day_time[1]),'minute')),
-    ':00')
+    "T",
+    sprintf("%02d", attr(hms(day_time[1]), "hour")),
+    ":",
+    sprintf("%02d", attr(hms(day_time[1]), "minute")),
+    ":00"
+  )
 
   end_day <- paste0(
-    'T',
-    sprintf("%02d",attr(hms(day_time[2]),'hour')),
-    ':',
-    sprintf("%02d",attr(hms(day_time[2]),'minute')),
-    ':00')
+    "T",
+    sprintf("%02d", attr(hms(day_time[2]), "hour")),
+    ":",
+    sprintf("%02d", attr(hms(day_time[2]), "minute")),
+    ":00"
+  )
 
-  day_range = paste0(start_day, '/', end_day)
+  day_range <- paste0(start_day, "/", end_day)
 
-  #Formatting the time range of the night
+  # Formatting the time range of the night
   start_night <- paste0(
-    'T',
-    sprintf("%02d",attr(hms(night_time[1]),'hour')),
-    ':',
[…]

### 2024-07-16 R/globals.R

@@ -1,20 +1,22 @@
 utils::globalVariables(
-  c("datetime",
-    'average',
-    'hp',
-    'where',
-    'na.omit',
-    'Date',
-    'Diurnality',
-    'fibsearch',
-    'frequency_hz',
-    'period_seconds',
-    'status_harmonic',
-    'power',
-    'period_hours',
-    'new_h',
-    'to',
-    'tz',
-    'df',
-    'date_numeric')
+  c(
+    "datetime",
+    "average",
+    "hp",
+    "where",
+    "na.omit",
+    "Date",
+    "Diurnality",
+    "fibsearch",
+    "frequency_hz",
+    "period_seconds",
+    "status_harmonic",
+    "power",
+    "period_hours",
+    "new_h",
+    "to",
+    "tz",
+    "df",
+    "date_numeric"
   )
+)

### 2024-07-16 R/import_raw_activity_data.R

@@ -57,96 +57,96 @@
 #'
 #' filename <- system.file("extdata", "sample_data.csv", package = "digiRhythm")
 #' data <- import_raw_activity_data(
-#'     filename,
-#'     skipLines = 7,
-#'     act.cols.names = c("Date", "Time", "Motion Index", 'Steps'),
-#'     sep = ',',
-#'     original_tz = 'CET',
-#'     target_tz = 'CET',
-#'     date_format = "%d.%m.%Y",
-#'     time_format = "%H:%M:%S",
-#'     sampling = 15,
-#'     trim_first_day = TRUE,
-#'     trim_middle_days = TRUE,
-#'     trim_last_day = TRUE,
-#'     verbose = TRUE)
+#'   filename,
+#'   skipLines = 7,
+#'   act.cols.names = c("Date", "Time", "Motion Index", "Steps"),
+#'   sep = ",",
+#'   original_tz = "CET",
+#'   target_tz = "CET",
+#'   date_format = "%d.%m.%Y",
+#'   time_format = "%H:%M:%S",
+#'   sampling = 15,
+#'   trim_first_day = TRUE,
+#'   trim_middle_days = TRUE,
+#'   trim_last_day = TRUE,
+#'   verbose = TRUE
+#' )
 #' print(head(data))
 #'
 #' @export
 
 
 import_raw_activity_data <- function(filename,
-                              skipLines = 0,
-                              act.cols.names = c("Date", "Time", "Motion Index", 'Steps'),
-                              date_format = "%d.%m.%Y",
-                              time_format = "%H:%M:%S",
-                              sep = ',',
-                              original_tz = 'CET',
-                              target_tz = 'CET',
-                              sampling = 15,
-                              trim_first_day = TRUE,
-                              trim_middle_days = TRUE,
-                              trim_last_day = TRUE,
-                              verbose = FALSE){
-
-
+                                     skipLines = 0,
+                                     act.cols.names = c("Date", "Time", "Motion Index", "Steps"),
+                                     date_format = "%d.%m.%Y",
+                                     time_format = "%H:%M:%S",
+                                     sep = ",",
+                                     original_tz = "CET",
+                                     target_tz = "CET",
+                                     sampling = 15,
+                                     trim_first_day = TRUE,
+                                     trim_middle_days = TRUE,
[…]

### 2024-07-16 R/is_dgm_friendly.R

@@ -16,51 +16,48 @@
 #' data("df516b_2", package = "digiRhythm")
 #' d <- df516b_2
 #' is_dgm_friendly(data = d, verbose = TRUE)
-
-
-is_dgm_friendly <- function(data, verbose = FALSE){
-
+is_dgm_friendly <- function(data, verbose = FALSE) {
   is_dgm <- TRUE
 
-  if(is.null(data)){
-    message <- red('Data cannot be NULL')
+  if (is.null(data)) {
+    message <- red("Data cannot be NULL")
     print_v(message, verbose)
     is_dgm <- FALSE
-  } else{
-    #Checking if the first column meets the requirements
-    if("POSIXct" %in%  class(data[,1])  | "POSIXt" %in%  class(data[,1])){
-      message <- paste0(green('v Correct time format: '), 'First column has a POSIXct Format')
+  } else {
+    # Checking if the first column meets the requirements
+    if ("POSIXct" %in% class(data[, 1]) | "POSIXt" %in% class(data[, 1])) {
+      message <- paste0(green("v Correct time format: "), "First column has a POSIXct Format")
       print_v(message, verbose)
-    }else {
-      message <- paste0(red('x Inorrect time format: '), 'First column does not a POSIXct Format')
-      print_v(message, verbose)
-      is_dgm <- FALSE
-    }
-
-
-    #Checking if data contains at least 7 days of data
-    if(length(unique(as.Date(data[,1]))) >= 2){
-      message <- paste0(green('v Number of days: '), 'Bigger or equal to 2')
-      print_v(message, verbose)
-    }else {
-      message <- paste0(red('x Number of days: '), 'Less than 2 (minimum of 2 days is recommended)')
+    } else {
+      message <- paste0(red("x Inorrect time format: "), "First column does not a POSIXct Format")
       print_v(message, verbose)
       is_dgm <- FALSE
     }
 
-    #Checking if we have more than one column
-    if(ncol(data) == 1){
-      message <- paste0(red('x Illogical number of columns: '), 'The dataset has only one column. Minimum number of columns is 2')
+
+    # Checking if data contains at least 7 days of data
+    if (length(unique(as.Date(data[, 1]))) >= 2) {
+      message <- paste0(green("v Number of days: "), "Bigger or equal to 2")
+      print_v(message, verbose)
+    } else {
+      message <- paste0(red("x Number of days: "), "Less than 2 (minimum of 2 days is recommended)")
       print_v(message, verbose)
[…]

### 2024-07-16 R/lomb_scargle_periodogram.R

@@ -26,15 +26,12 @@
 #'
 #' @examples
 #' data("df516b_2", package = "digiRhythm")
-#' data <- df516b_2[1:672, c(1,2)]
+#' data <- df516b_2[1:672, c(1, 2)]
 #' sig <- 0.01
 #' lomb_scargle_periodogram(data, alpha = sig, plot = TRUE)
-
-
-lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = TRUE, extra_info_plot = TRUE) {
-
+lomb_scargle_periodogram <- function(data, alpha = 0.01, sampling = 15, plot = TRUE, extra_info_plot = TRUE) {
   if (!is_dgm_friendly(data, verbose = TRUE)) {
-    stop('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
+    stop("The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information")
   }
 
   # if (length(unique(as.Date(data[,1]))) != 7 ) {
@@ -43,15 +40,17 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = 
   # }
 
   x <- data
-  start <- as.Date(x[1,1])
+  start <- as.Date(x[1, 1])
   # print(start)
-  end <- as.Date(x[nrow(x),1])
+  end <- as.Date(x[nrow(x), 1])
   # print(end)
 
-  from_to = paste('From',
-                  start,
-                  'To',
-                  end)
+  from_to <- paste(
+    "From",
+    start,
+    "To",
+    end
+  )
 
   ofac <- 1
   names <- colnames(x)
@@ -82,8 +81,8 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = 
 
   n <- length(y)
   tspan <- t[n] - t[1]
-  fr.d <- 1/tspan
-  step <- 1/(tspan * ofac)
+  fr.d <- 1 / tspan
+  step <- 1 / (tspan * ofac)
 
 
   # if (is.null(to)) {
@@ -97,13 +96,13 @@ lomb_scargle_periodogram <- function (data, alpha = 0.01, sampling = 15, plot = 
 
   x <- t * 2 * pi
   y <- y - mean(y)
-  norm = 1/sum(y^2)
+  norm <- 1 / sum(y^2)
 
   w <- 2 * pi * freq
[…]

### 2024-07-16 R/resample_dgm.R

@@ -18,40 +18,38 @@
 #' df <- remove_activity_outliers(df)
 #' new_sampling <- 30
 #' new_dgm <- resample_dgm(df, new_sampling)
-
-
-resample_dgm <- function(data, new_sampling){
-
+resample_dgm <- function(data, new_sampling) {
   xts_data <- xts(
-    x = data[,c(2:ncol(data))],
-    order.by = data[,1]
+    x = data[, c(2:ncol(data))],
+    order.by = data[, 1]
   )
 
 
   original_sampling <- xts::periodicity(xts_data)$frequency
 
-  if((new_sampling %% original_sampling) != 0){
+  if ((new_sampling %% original_sampling) != 0) {
     stop("The new sampling should be a multiple of the current sampling in minutes")
   }
 
-  if(new_sampling < original_sampling){
+  if (new_sampling < original_sampling) {
     stop("The new sampling should be bigger than the current sampling")
   }
 
-  if(is.null(new_sampling)){
-    stop('The new sampling cannot be NULL')
+  if (is.null(new_sampling)) {
+    stop("The new sampling cannot be NULL")
   }
 
-  if(new_sampling <= 0 ){
-    stop('The new sampling cannot be non-positive')
+  if (new_sampling <= 0) {
+    stop("The new sampling cannot be non-positive")
   }
 
   sampled_xts <- NULL
-  for(var in names(xts_data)){
+  for (var in names(xts_data)) {
     xts_var <- period.apply(
-      xts_data[,var],
-      endpoints(xts_data, on = 'minutes', k = new_sampling),
-      FUN = sum)
+      xts_data[, var],
+      endpoints(xts_data, on = "minutes", k = new_sampling),
+      FUN = sum
+    )
     sampled_xts <- cbind(sampled_xts, xts_var)
   }
 

### 2024-07-16 R/sliding_DI.R

@@ -18,7 +18,7 @@
 #' data("df516b_2", package = "digiRhythm")
 #' data <- df516b_2
 #' data <- remove_activity_outliers(data)
-#' activity = names(data)[2]
+#' activity <- names(data)[2]
 #' data("timedata", package = "digiRhythm")
 #' timedata <- timedata
 #' d_index <- sliding_DI(data, activity, timedata)
@@ -26,96 +26,106 @@
 #' @export
 
 sliding_DI <- function(data,
-                        activity,
-                        timedata,
-                        save = NULL){
+                       activity,
+                       timedata,
+                       save = NULL) {
+  start_date <- lubridate::date(data[, 1])[1]
+  end_date <- last(lubridate::date(data[, 1]))
 
-start_date <- lubridate::date(data[, 1])[1]
-end_date <- last(lubridate::date(data[, 1]))
+  dates <- unique(lubridate::date(data[, 1]))
+  X <- xts(x = data[[activity]], order.by = data[, 1])
+  sampling <- dgm_periodicity(data)[["frequency"]]
 
-dates <- unique(lubridate::date(data[, 1]))
-X <- xts(x = data[[activity]], order.by = data[, 1])
-sampling <- dgm_periodicity(data)[["frequency"]]
+  # Formatting time range of day and night
+  timedata <- subset(timedata, lubridate::date(timedata$night_end) >= start_date & lubridate::date(timedata$night_end) <= end_date)
+  day_range <- paste0(timedata$day_start, "/", timedata$day_end)
+  night_range <- paste0(timedata$night_start, "/", timedata$night_end[2:nrow(timedata)])
 
-#Formatting time range of day and night
-timedata <- subset(timedata, lubridate::date(timedata$night_end) >= start_date & lubridate::date(timedata$night_end) <= end_date)
-day_range = paste0(timedata$day_start, "/", timedata$day_end)
-night_range = paste0(timedata$night_start, "/", timedata$night_end[2:nrow(timedata)])
+  # Compute the range of samples for the day and the night (Td & Tn)
+  hms_day_start <- hms(substr(timedata$day_start, 12, 19))
+  hms_day_end <- hms(substr(timedata$day_end, 12, 19))
+  sample_size <- hms(paste0("00:", sampling, ":00"))
+  Td <- abs((hms_day_end - hms_day_start) / sample_size)
 
[…]

### 2024-07-16 R/utils.R

@@ -7,23 +7,21 @@
 #' @export
 #'
 
-df_act_info <- function(df){
-
-  print('First days of the data set: ')
+df_act_info <- function(df) {
+  print("First days of the data set: ")
   print(head(df))
 
-  print('Last days of the data set: ')
+  print("Last days of the data set: ")
   print(tail(df))
 
-  print(paste('The dataset contains', length(unique(as.Date(df[,1]))), 'Days'))
+  print(paste("The dataset contains", length(unique(as.Date(df[, 1]))), "Days"))
 
-  first <- as.Date(df[1,1])
+  first <- as.Date(df[1, 1])
 
-  print(paste('Starting date is:', first))
+  print(paste("Starting date is:", first))
 
-  last <- as.Date(df[nrow(df),1])
-  print(paste('Last date is:', last))
-
+  last <- as.Date(df[nrow(df), 1])
+  print(paste("Last date is:", last))
 }
 
 #' Remove outliers from the data
@@ -37,17 +35,16 @@ df_act_info <- function(df){
 #'
 
 
-remove_activity_outliers <- function(df){
-
+remove_activity_outliers <- function(df) {
   data <- as.data.frame(df)
   for (i in 2:ncol(df)) {
-    Q <- quantile(data[,i], probs = c(.25, .75), na.rm = TRUE)
-    iqr <- IQR(data[,i])
-    up <-  Q[2] + 1.5*iqr # Upper Range
-    low <- Q[1] - 1.5*iqr # Lower Range
-    non_outliers <- which(data[,i] <= up | data[,i] >= low)
-    outliers <- which(data[,i] > up | data[,i] < low)
-    mean_without_outliers <- mean(data[non_outliers,i], na.rm = TRUE)
+    Q <- quantile(data[, i], probs = c(.25, .75), na.rm = TRUE)
+    iqr <- IQR(data[, i])
+    up <- Q[2] + 1.5 * iqr # Upper Range
+    low <- Q[1] - 1.5 * iqr # Lower Range
+    non_outliers <- which(data[, i] <= up | data[, i] >= low)
+    outliers <- which(data[, i] > up | data[, i] < low)
+    mean_without_outliers <- mean(data[non_outliers, i], na.rm = TRUE)
[…]

### 2024-07-16 build_cran_package.sh

@@ -28,6 +28,5 @@ cp -R README.md ../dgm
 
 cd ..
 
-/opt/local/R/4.2.3/bin/R CMD check dgm
 /opt/local/R/4.2.3/bin/R CMD build dgm
 /opt/local/R/4.2.3/bin/R CMD check digirhythm_1.2.tar.gz --as-cran dgm

### 2024-07-16 tests/testthat.R

@@ -1,2 +1,1 @@
 library(testthat)
-

### 2024-07-16 R/utils.R
Removed 'import from' from the utils.R source code

@@ -1,7 +1,6 @@
 #' Outputs some information about the activity dataframe
 #'
 #' @param df The dataframe containing the activity data
-#' @importFrom utils head tail
 #' @return No return value. Prints the head and tail as well as the starting and
 #'  end date of a digiRhythm friendly dataframe.
 #' @export
@@ -9,10 +8,10 @@
 
 df_act_info <- function(df) {
   print("First days of the data set: ")
-  print(head(df))
+  print(utils::head(df))
 
   print("Last days of the data set: ")
-  print(tail(df))
+  print(utils::tail(df))
 
   print(paste("The dataset contains", length(unique(as.Date(df[, 1]))), "Days"))
 
@@ -28,7 +27,6 @@ df_act_info <- function(df) {
 #'
 #' @param df The dataframe containing the activity data
 #'
-#' @importFrom stats IQR quantile
 #' @return return a dataframe where columns start the second one have undergone
 #' an outlier removal.
 #' @export
@@ -38,8 +36,8 @@ df_act_info <- function(df) {
 remove_activity_outliers <- function(df) {
   data <- as.data.frame(df)
   for (i in 2:ncol(df)) {
-    Q <- quantile(data[, i], probs = c(.25, .75), na.rm = TRUE)
-    iqr <- IQR(data[, i])
+    Q <- stats::quantile(data[, i], probs = c(.25, .75), na.rm = TRUE)
+    iqr <- stats::IQR(data[, i])
     up <- Q[2] + 1.5 * iqr # Upper Range
     low <- Q[1] - 1.5 * iqr # Lower Range
     non_outliers <- which(data[, i] <= up | data[, i] >= low)
@@ -55,8 +53,6 @@ remove_activity_outliers <- function(df) {
 #'
 #' @param string The string to print
 #' @param verbose if TRUE, print the string
-#'
-#' @importFrom crayon red green
 #'
 #' @return No return value. Prints the string concatenated with a verbose if the
 #'  latter is not NULL.
@@ -74,8 +70,6 @@ print_v <- function(
 #'
 #' @param data a digiRhythm friendly dataframe
 #'
-#' @importFrom xts periodicity as.xts
-#'
 #' @return returns a periodicity object of type xts.
 #' @export
 #' @examples
@@ -84,18 +78,10 @@ print_v <- function(
 #' dgm_periodicity(df)
 #'
 dgm_periodicity <- function(data) {
[…]

### 2024-07-16 R/sliding_DI.R
removed 'import from' from the source code 'sliding_DI.R'

@@ -11,9 +11,6 @@
 #'
 #' @return A ggplot2 object that contains the Sliding diurnality plot in addition to a dataframe with 2 col: date and sliding diurnality index
 #'
-#' @importFrom lubridate date hms hour minute
-#' @importFrom xts xts period.apply merge.xts
-#'
 #' @examples
 #' data("df516b_2", package = "digiRhythm")
 #' data <- df516b_2
@@ -33,7 +30,7 @@ sliding_DI <- function(data,
   end_date <- last(lubridate::date(data[, 1]))
 
   dates <- unique(lubridate::date(data[, 1]))
-  X <- xts(x = data[[activity]], order.by = data[, 1])
+  X <- xts::xts(x = data[[activity]], order.by = data[, 1])
   sampling <- dgm_periodicity(data)[["frequency"]]
 
   # Formatting time range of day and night
@@ -42,15 +39,15 @@ sliding_DI <- function(data,
   night_range <- paste0(timedata$night_start, "/", timedata$night_end[2:nrow(timedata)])
 
   # Compute the range of samples for the day and the night (Td & Tn)
-  hms_day_start <- hms(substr(timedata$day_start, 12, 19))
-  hms_day_end <- hms(substr(timedata$day_end, 12, 19))
-  sample_size <- hms(paste0("00:", sampling, ":00"))
+  hms_day_start <- lubridate::hms(substr(timedata$day_start, 12, 19))
+  hms_day_end <- lubridate::hms(substr(timedata$day_end, 12, 19))
+  sample_size <- lubridate::hms(paste0("00:", sampling, ":00"))
   Td <- abs((hms_day_end - hms_day_start) / sample_size)
 
-  hms_night_start <- hms(substr(timedata$night_start[1:nrow(timedata) - 1], 12, 19))
-  hms_midnight <- hms("00:00:00")
+  hms_night_start <- lubridate::hms(substr(timedata$night_start[1:nrow(timedata) - 1], 12, 19))
+  hms_midnight <- lubridate::hms("00:00:00")
   hms_24t <- hms("24:00:00")
-  hms_night_end <- hms(substr(timedata$night_end[2:nrow(timedata)], 12, 19))
+  hms_night_end <- lubridate::hms(substr(timedata$night_end[2:nrow(timedata)], 12, 19))
   Tn <- (hms_24t - hms_night_start + hms_night_end - hms_midnight) / sample_size
 
[…]

### 2024-07-16 R/resample_dgm.R
Removed the 'import from' from the source code 'resample.R'

@@ -8,8 +8,6 @@
 #'
 #' @return A digiRhythm friendly dataset with the new sampling
 #'
-#' @importFrom xts xts period.apply
-#' @importFrom zoo index coredata
 #' @export
 #'
 #' @examples
@@ -19,7 +17,7 @@
 #' new_sampling <- 30
 #' new_dgm <- resample_dgm(df, new_sampling)
 resample_dgm <- function(data, new_sampling) {
-  xts_data <- xts(
+  xts_data <- xts::xts(
     x = data[, c(2:ncol(data))],
     order.by = data[, 1]
   )
@@ -45,7 +43,7 @@ resample_dgm <- function(data, new_sampling) {
 
   sampled_xts <- NULL
   for (var in names(xts_data)) {
-    xts_var <- period.apply(
+    xts_var <- xts::period.apply(
       xts_data[, var],
       endpoints(xts_data, on = "minutes", k = new_sampling),
       FUN = sum
@@ -54,7 +52,7 @@ resample_dgm <- function(data, new_sampling) {
   }
 
   new_data <- data.frame(
-    datetime = index(sampled_xts),
+    datetime = zoo::index(sampled_xts),
     coredata(sampled_xts)
   )
 

### 2024-07-16 R/lomb_scargle_periodogram.R
Removed 'import from' from the source code 'lomb_scarlge_periodogram.R'

@@ -20,7 +20,6 @@
 #' the frequency is harmonic or not.
 #'
 #' @import ggplot2
-#' @importFrom pracma fibsearch
 #'
 #' @export
 #'
@@ -131,7 +130,7 @@ lomb_scargle_periodogram <- function(data, alpha = 0.01, sampling = 15, plot = T
   })
 
   p.values <- unlist(p.values)
-  level <- fibsearch(levopt, 0, 1, alpha, fmax = fmax, tm = t)$xmin
+  level <- pracma:fibsearch(levopt, 0, 1, alpha, fmax = fmax, tm = t)$xmin
 
   # Returns a list that contains:
   # 1. an LSP dataframe with the following cols: Freq, Power, Harmonc Status,

### 2024-07-16 R/is_dgm_friendly.R
Remove the 'import from' from the source code 'is_dgm_friendly'

@@ -9,7 +9,7 @@
 #'
 #' @return Boolean. If True, the dataframe is digirhythm friendly. If False, the
 #' dataframe is not digirhythm friendly.
-#' @importFrom crayon green blue red
+#'
 #' @export
 #'
 #' @examples
@@ -20,16 +20,16 @@ is_dgm_friendly <- function(data, verbose = FALSE) {
   is_dgm <- TRUE
 
   if (is.null(data)) {
-    message <- red("Data cannot be NULL")
+    message <- crayon::red("Data cannot be NULL")
     print_v(message, verbose)
     is_dgm <- FALSE
   } else {
     # Checking if the first column meets the requirements
     if ("POSIXct" %in% class(data[, 1]) | "POSIXt" %in% class(data[, 1])) {
-      message <- paste0(green("v Correct time format: "), "First column has a POSIXct Format")
+      message <- paste0(crayon::green("v Correct time format: "), "First column has a POSIXct Format")
       print_v(message, verbose)
     } else {
-      message <- paste0(red("x Inorrect time format: "), "First column does not a POSIXct Format")
+      message <- paste0(crayon::red("x Inorrect time format: "), "First column does not a POSIXct Format")
       print_v(message, verbose)
       is_dgm <- FALSE
     }
@@ -37,27 +37,27 @@ is_dgm_friendly <- function(data, verbose = FALSE) {
 
     # Checking if data contains at least 7 days of data
     if (length(unique(as.Date(data[, 1]))) >= 2) {
-      message <- paste0(green("v Number of days: "), "Bigger or equal to 2")
+      message <- paste0(crayon::green("v Number of days: "), "Bigger or equal to 2")
       print_v(message, verbose)
     } else {
-      message <- paste0(red("x Number of days: "), "Less than 2 (minimum of 2 days is recommended)")
+      message <- paste0(crayon::red("x Number of days: "), "Less than 2 (minimum of 2 days is recommended)")
       print_v(message, verbose)
       is_dgm <- FALSE
     }
 
     # Checking if we have more than one column
     if (ncol(data) == 1) {
-      message <- paste0(red("x Illogical number of columns: "), "The dataset has only one column. Minimum number of columns is 2")
+      message <- paste0(crayon::red("x Illogical number of columns: "), "The dataset has only one column. Minimum number of columns is 2")
       print_v(message, verbose)
       is_dgm <- FALSE
     } else {
       # Checking if data are numeric
[…]

### 2024-07-16 R/import_raw_activity_data.R
Removed the 'import from' from the source code 'import_raw_activity_data'

@@ -43,15 +43,7 @@
 #' @return A dataframe with datetime column and other activity columns, ready to
 #' be used with other functions in digirhythm
 #'
-#' @importFrom tidyr unite
-#' @importFrom magrittr %>%
-#' @importFrom readr read_delim
-#' @importFrom xts endpoints period.apply xts
-#' @importFrom zoo coredata index
-#' @importFrom dplyr filter select last tally
-#' @importFrom utils read.table
-#' @importFrom stringr str_trim
-#' @importFrom lubridate date round_date
+#' @import magrittr
 #'
 #' @examples
 #'
@@ -96,7 +88,7 @@ import_raw_activity_data <- function(filename,
   # Loading data from the CSV (with specific columns and skipping lines)
 
 
-  data <- read_delim(filename,
+  data <- readr::read_delim(filename,
     skip = skipLines,
     delim = sep,
     show_col_types = FALSE
@@ -105,9 +97,9 @@ import_raw_activity_data <- function(filename,
   data <- na.omit(data)
 
   data <- data %>%
-    mutate(across(where(is.character), str_trim))
+    mutate(across(where(is.character), stringr::str_trim))
 
-  data <- data %>% unite(datetime, c(act.cols.names[1], act.cols.names[2]), sep = "-")
+  data <- data %>% tidyr::unite(datetime, c(act.cols.names[1], act.cols.names[2]), sep = "-")
 
   data$datetime <- as.POSIXct(data$datetime, format = paste0(date_format, " -", time_format), tz = original_tz)
 
@@ -137,7 +129,7 @@ import_raw_activity_data <- function(filename,
 
 
   # Transforming data to an XTS for easy management of sampling and date removal
-  data_xts <- xts(
+  data_xts <- xts::xts(
     data[, 2:ncol(data)],
     order.by = data$datetime
   )
@@ -147,7 +139,7 @@ import_raw_activity_data <- function(filename,
   for (var in names(data_xts)) {
     var_xts <- xts::period.apply(
       data_xts[, var],
-      endpoints(data_xts, "minutes", k = sampling),
+      xts::endpoints(data_xts, "minutes", k = sampling),
       FUN = sum
     )
     data_xts_sampled <- cbind(data_xts_sampled, var_xts)
@@ -156,8 +148,8 @@ import_raw_activity_data <- function(filename,
 
   # Creating a dataframe from the sampled XTS (what we will return)
   df <- data.frame(
[…]

### 2024-07-16 R/resample_dgm.R

@@ -45,7 +45,7 @@ resample_dgm <- function(data, new_sampling) {
   for (var in names(xts_data)) {
     xts_var <- xts::period.apply(
       xts_data[, var],
-      endpoints(xts_data, on = "minutes", k = new_sampling),
+      xts::endpoints(xts_data, on = "minutes", k = new_sampling),
       FUN = sum
     )
     sampled_xts <- cbind(sampled_xts, xts_var)

### 2024-07-16 R/sliding_DI.R

@@ -11,6 +11,8 @@
 #'
 #' @return A ggplot2 object that contains the Sliding diurnality plot in addition to a dataframe with 2 col: date and sliding diurnality index
 #'
+#' @import ggplot2
+#'
 #' @examples
 #' data("df516b_2", package = "digiRhythm")
 #' data <- df516b_2
@@ -63,7 +65,7 @@ sliding_DI <- function(data,
 
   # Computing Cd
   X_day <- X[day_range]
-  Cd <- xts::period.apply(X_day, endpoints(X_day, "days"), sum)
+  Cd <- xts::period.apply(X_day, xts::endpoints(X_day, "days"), sum)
 
   # Computing day value
   day_val <- Cd / Td
@@ -73,7 +75,7 @@ sliding_DI <- function(data,
   offset <- 3600 * hour(lubridate::hms("12:00:00"))
   zoo::index(X_night) <- zoo::index(X_night) - offset
   Cn <- xts::period.apply(
-    X_night, endpoints(X_night, "days"),
+    X_night, xts::endpoints(X_night, "days"),
     sum
   )
 

### 2024-07-16 R/diurnality.R
Removed the 'import from' from the source code 'diurnality.R'

@@ -12,8 +12,7 @@
 #'
 #' @return A ggplot2 object that contains the diurnality plot in addition to a dataframe with 2 col: date and diurnality index
 #'
-#' @importFrom lubridate date hms hour minute
-#' @importFrom xts xts period.apply merge.xts
+#' @import ggplot2
 #'
 #' @examples
 #' data("df516b_2", package = "digiRhythm")
@@ -32,7 +31,7 @@ diurnality <- function(data,
   # di = (cd/td - cn/tn)/(cd/td + cn/tn)
 
   dates <- unique(lubridate::date(data[, 1]))
-  X <- xts(
+  X <- xts::xts(
     x = data[[activity]],
     order.by = data[, 1]
   )
@@ -87,7 +86,7 @@ diurnality <- function(data,
   hms_night_start <- hms(night_time[1])
   hms_midnight <- hms("00:00:00")
   hms_24t <- hms("24:00:00")
-  hms_night_end <- hms(night_time[2])
+  hms_night_end <- lubridate::hms(night_time[2])
   Tn <- (hms_24t - hms_night_start + hms_night_end - hms_midnight) / sample_size
 
   # Check conditions about the day and night time (overlapping, misplacement)
@@ -95,16 +94,16 @@ diurnality <- function(data,
     stop("The end of the day_time period should proceed the beginning of the night_time period")
   }
 
-  if (hour(hms_night_end) < 1) {
+  if (lubridate::hour(hms_night_end) < 1) {
     stop("The of the nightly period should be after midnight")
   }
 
-  if (hour(hms_night_end) > 11) {
+  if (lubridate::hour(hms_night_end) > 11) {
     stop("The end of the nightly period cannot be after mid-day! Come on!")
   }
 
   X_day <- X[day_range]
-  Cd <- period.apply(X_day, endpoints(X_day, "days"), sum)
+  Cd <- xts::period.apply(X_day, endpoints(X_day, "days"), sum)
 
   # Computing Td for the motion index
   # Td <- 40 # 40 samples * 15 minutes = 10 hours
@@ -120,7 +119,7 @@ diurnality <- function(data,
   shift <- hour(hms_24t - hms_night_start + hms_night_end - hms_midnight)
   shifted_night_range <- paste0("T00:00:00/T", shift, ":00:00")
   X_night <- X_night[shifted_night_range]
-  Cn <- period.apply(X_night, endpoints(X_night, "days"), sum)
[…]

### 2024-07-16 R/dfc.R
Removed 'import from' from the source code dfc.R

@@ -44,9 +44,7 @@
 #' Missing days are not permitted. If you have data with half day, it should be
 #' removed.
 #'
-#' @importFrom lubridate date
-#' @importFrom dplyr filter
-#' @importFrom stats ts
+#' @import ggplot2
 #'
 #' @export
 #' @examples
@@ -76,7 +74,7 @@ dfc <- function(
   }
 
 
-  data$date <- date(data$datetime)
+  data$date <- lubridate::date(data$datetime)
 
   days <- seq(
     data$date[1],
@@ -130,8 +128,8 @@ dfc <- function(
     # Filtering the next seven days by date (not by index - in case of missing data, filtering by index would make errors)
 
     data_week <- data %>%
-      filter(date >= days[index_start_day]) %>%
-      filter(date <= days[index_end_day])
+      dplyr::filter(date >= days[index_start_day]) %>%
+      dplyr::filter(date <= days[index_end_day])
 
 
     # Selecting the first column (datetime) and the activity column

### 2024-07-16 R/daily_average_activity.R
Removed 'import from' from the source code daily_average_activity.R

@@ -14,10 +14,8 @@
 #'
 #' @return None
 #'
-#' @importFrom magrittr %>%
-#' @importFrom stats time
+#' @import magrittr
 #' @import ggplot2
-#' @importFrom lubridate date tz
 #' @import dplyr
 #'
 #' @export
@@ -68,7 +66,7 @@ daily_average_activity <- function(
   s <- sum_of_activity_over_all_days_per_sample
 
   s$datetime <- paste(data_to_plot$date[1], s$time)
-  s$datetime <- as.POSIXct(s$datetime, format("%Y-%m-%d %H:%M"), tz = tz(df$datetime))
+  s$datetime <- as.POSIXct(s$datetime, format("%Y-%m-%d %H:%M"), tz = lubridate::tz(df$datetime))
 
   s <- s %>% select(datetime, average)
 

### 2024-07-16 R/daily_acticity_wrap_plot.R
Remove the 'import from' from the source code daily_activity_wrap_plot.R

@@ -16,11 +16,9 @@
 #'
 #' @return A ggplot2 object that contains the daily average activity plot
 #'
-#' @importFrom magrittr %>%
-#' @importFrom stats time
-#' @import ggplot2
-#' @importFrom lubridate date
 #' @import dplyr
+#' @import ggplot2
+#' @import magrittr
 #'
 #' @export
 #'

### 2024-07-16 R/diurnality.R
Fixed few build errors that appeared after massive change to the code base.

@@ -42,17 +42,17 @@ diurnality <- function(data,
   # Formatting the time range of the day
   start_day <- paste0(
     "T",
-    sprintf("%02d", attr(hms(day_time[1]), "hour")),
+    sprintf("%02d", attr(lubridate::hms(day_time[1]), "hour")),
     ":",
-    sprintf("%02d", attr(hms(day_time[1]), "minute")),
+    sprintf("%02d", attr(lubridate::hms(day_time[1]), "minute")),
     ":00"
   )
 
   end_day <- paste0(
     "T",
-    sprintf("%02d", attr(hms(day_time[2]), "hour")),
+    sprintf("%02d", attr(lubridate::hms(day_time[2]), "hour")),
     ":",
-    sprintf("%02d", attr(hms(day_time[2]), "minute")),
+    sprintf("%02d", attr(lubridate::hms(day_time[2]), "minute")),
     ":00"
   )
 
@@ -61,31 +61,31 @@ diurnality <- function(data,
   # Formatting the time range of the night
   start_night <- paste0(
     "T",
-    sprintf("%02d", attr(hms(night_time[1]), "hour")),
+    sprintf("%02d", attr(lubridate::hms(night_time[1]), "hour")),
     ":",
-    sprintf("%02d", attr(hms(night_time[1]), "minute")),
+    sprintf("%02d", attr(lubridate::hms(night_time[1]), "minute")),
     ":00"
   )
 
   end_night <- paste0(
     "T",
-    sprintf("%02d", attr(hms(night_time[2]), "hour")),
+    sprintf("%02d", attr(lubridate::hms(night_time[2]), "hour")),
     ":",
-    sprintf("%02d", attr(hms(night_time[2]), "minute")),
+    sprintf("%02d", attr(lubridate::hms(night_time[2]), "minute")),
     ":00"
   )
 
   night_range <- paste0(start_night, "/", end_night)
 
   # Compute the range of samples for the day and the night
-  hms_day_start <- hms(day_time[1])
-  hms_day_end <- hms(day_time[2])
-  sample_size <- hms(paste0("00:", sampling, ":00"))
+  hms_day_start <- lubridate::hms(day_time[1])
+  hms_day_end <- lubridate::hms(day_time[2])
+  sample_size <- lubridate::hms(paste0("00:", sampling, ":00"))
   Td <- abs((hms_day_end - hms_day_start) / sample_size)
 
[…]

### 2024-07-16 R/import_raw_activity_data.R

@@ -169,7 +169,7 @@ import_raw_activity_data <- function(filename,
 
   if (trim_first_day) {
     n_samples_day1 <- df %>%
-      filter(lubridate::date == unique(df$date)[1]) %>%
+      filter(date == unique(df$date)[1]) %>%
       dplyr::tally()
     if (n_samples_day1 < smallest_mandatory_daily_samples) {
       df <- df %>% filter(date != unique(df$date)[1])

### 2024-07-16 R/lomb_scargle_periodogram.R

@@ -130,7 +130,7 @@ lomb_scargle_periodogram <- function(data, alpha = 0.01, sampling = 15, plot = T
   })
 
   p.values <- unlist(p.values)
-  level <- pracma:fibsearch(levopt, 0, 1, alpha, fmax = fmax, tm = t)$xmin
+  level <- pracma::fibsearch(levopt, 0, 1, alpha, fmax = fmax, tm = t)$xmin
 
   # Returns a list that contains:
   # 1. an LSP dataframe with the following cols: Freq, Power, Harmonc Status,

### 2024-07-16 R/resample_dgm.R

@@ -53,7 +53,7 @@ resample_dgm <- function(data, new_sampling) {
 
   new_data <- data.frame(
     datetime = zoo::index(sampled_xts),
-    coredata(sampled_xts)
+    zoo::coredata(sampled_xts)
   )
 
   return(new_data)

### 2024-07-16 R/sliding_DI.R

@@ -48,7 +48,7 @@ sliding_DI <- function(data,
 
   hms_night_start <- lubridate::hms(substr(timedata$night_start[1:nrow(timedata) - 1], 12, 19))
   hms_midnight <- lubridate::hms("00:00:00")
-  hms_24t <- hms("24:00:00")
+  hms_24t <- lubridate::hms("24:00:00")
   hms_night_end <- lubridate::hms(substr(timedata$night_end[2:nrow(timedata)], 12, 19))
   Tn <- (hms_24t - hms_night_start + hms_night_end - hms_midnight) / sample_size
 
@@ -56,10 +56,10 @@ sliding_DI <- function(data,
   # if (hms_day_end > hms_night_start) {
   # stop("The end of the day_time period should proceed the beginning of the night_time period")
   # }
-  # if (hour(hms_night_end) < 1) {
+  # if (lubridate::hour(hms_night_end) < 1) {
   # stop("The of the nightly period should be after midnight")
   # }
-  # if (hour(hms_night_end) > 11) {
+  # if (lubridate::hour(hms_night_end) > 11) {
   # stop("The end of the nightly period cannot be after mid-day! Come on!")
   # }
 
@@ -72,7 +72,7 @@ sliding_DI <- function(data,
 
   # Computing Cn
   X_night <- X[night_range]
-  offset <- 3600 * hour(lubridate::hms("12:00:00"))
+  offset <- 3600 * lubridate::hour(lubridate::hms("12:00:00"))
   zoo::index(X_night) <- zoo::index(X_night) - offset
   Cn <- xts::period.apply(
     X_night, xts::endpoints(X_night, "days"),
@@ -99,7 +99,7 @@ sliding_DI <- function(data,
   # Computing the diurnality index
   df <- data.frame(
     date = zoo::index(d),
-    diurnality = (coredata(d[, "day_val"]) - coredata(d[, "night_val"])) / (coredata(d[, "day_val"]) + coredata(d[, "night_val"]))
+    diurnality = (zoo::coredata(d[, "day_val"]) - zoo::coredata(d[, "night_val"])) / (zoo::coredata(d[, "day_val"]) + zoo::coredata(d[, "night_val"]))
   )
 
   names(df) <- c("date", "diurnality")

### 2024-07-16 R/utils.R

@@ -87,8 +87,8 @@ dgm_periodicity <- function(data) {
 }
 
 
-#' Returns p-value of a frequency peack according to pbaluev (2008) given Z,
-#' fmax and tm. Copied from the LOMB library.
+#' Returns p-value of a frequency peak according to pbaluev (2008) given Z,
+#' fmax and tm. Reused from the LOMB library (https://rdrr.io/cran/lomb/)
 #'
 #' @param Z the power of the frequency
 #' @param fmax the maximum frequency in the spectrum

### 2024-07-16 build_cran_package.sh

@@ -4,6 +4,8 @@
 
 #Moving the files
 rm -rf ../dgm/*
+rm -rf ../dgm.Rcheck/*
+rm -rf ../digiRhythm*.tar.gz
 
 #Data files
 cp -R data ../dgm

### 2024-07-25 R/actogram.R
Fixed a glitch in the dynamic LSP and styled the package with stylr

@@ -30,7 +30,8 @@
 #' end <- "2020-08-13" # year-month-day
 #' activity_alias <- "Motion Index"
 #' my_actogram <- actogram(df, activity, activity_alias, start, end,
-#'   save = NULL)
+#'   save = NULL
+#' )
 #' print(my_actogram)
 actogram <- function(
     df,
@@ -48,10 +49,15 @@ actogram <- function(
   data_to_plot <- df %>%
     filter(lubridate::date(datetime) >= start) %>%
     filter(lubridate::date(datetime) <= end)
-  data_to_plot$time <- format(data_to_plot$datetime, format = "%H:%M", tz = "CET")
+  data_to_plot$time <- format(data_to_plot$datetime,
+    format = "%H:%M",
+    tz = "CET"
+  )
   data_to_plot <- data_to_plot %>% select(-datetime)
 
-  data_to_plot$date_numeric <- xtfrm(as.Date(data_to_plot$date, format = "%Y-%m-%d"))
+  data_to_plot$date_numeric <- xtfrm(as.Date(data_to_plot$date,
+    format = "%Y-%m-%d"
+  ))
 
 
   equally_spaced_select <- function(x) {
@@ -60,7 +66,9 @@ actogram <- function(
   }
 
   if (length(unique(data_to_plot$date_numeric)) > 15) {
-    breaks_for_y_axis <- equally_spaced_select(unique(data_to_plot$date_numeric))
+    breaks_for_y_axis <- equally_spaced_select(unique(
+      data_to_plot$date_numeric
+    ))
     labels_for_y_axis <- equally_spaced_select(unique(data_to_plot$date))
   } else {
     breaks_for_y_axis <- unique(data_to_plot$date_numeric)

### 2024-07-25 R/diurnality.R

@@ -2,15 +2,16 @@
 #'
 #' @param data a digiRhythm-friendly dataset
 #' @param activity The number of non-useful lines to skip (lines to header)
-#' @param day_time an array containing the start and end of the day period. Default:
-#' c("06:30:00", "16:30:00").
-#' @param night_time an array containing the start and end of the night period. Default:
-#' c("18:00:00", "T05:00:00").
+#' @param day_time an array containing the start and end of the day period.
+#' Default: c("06:30:00", "16:30:00").
+#' @param night_time an array containing the start and end of the night period.
+#' Default: c("18:00:00", "T05:00:00").
 #' @param save if NULL, the image is not saved. Otherwise, this parameter will
 #' be the name of the saved image. it should contain the path and name without
 #' the extension.
 #'
-#' @return A ggplot2 object that contains the diurnality plot in addition to a dataframe with 2 col: date and diurnality index
+#' @return A ggplot2 object that contains the diurnality plot in addition to a
+#' dataframe with 2 col: date and diurnality index
 #'
 #' @import ggplot2
 #'
@@ -91,7 +92,8 @@ diurnality <- function(data,
 
   # Check conditions about the day and night time (overlapping, misplacement)
   if (hms_day_end > hms_night_start) {
-    stop("The end of the day_time period should proceed the beginning of the night_time period")
+    stop("The end of the day_time period should proceed the beginning of the
+         night_time period")
   }
 
   if (lubridate::hour(hms_night_end) < 1) {
@@ -116,7 +118,8 @@ diurnality <- function(data,
 
   offset <- 3600 * lubridate::hour(hms_night_start)
   zoo::index(X_night) <- zoo::index(X_night) - offset
-  shift <- lubridate::hour(hms_24t - hms_night_start + hms_night_end - hms_midnight)
+  shift <- lubridate::hour(hms_24t - hms_night_start +
+    hms_night_end - hms_midnight)
   shifted_night_range <- paste0("T00:00:00/T", shift, ":00:00")
   X_night <- X_night[shifted_night_range]
   Cn <- xts::period.apply(X_night, xts::endpoints(X_night, "days"), sum)
@@ -143,7 +146,10 @@ diurnality <- function(data,
   # computing the dirunality index
   df <- data.frame(
     date = zoo::index(d),
[…]

### 2024-07-25 R/is_dgm_friendly.R

@@ -1,8 +1,8 @@
 #' Informs if a dataset is digiRhythm Friendly
 #'
-#' Takes an activity dataset as input and gives information about 1) If a dataset
-#' is digiRhythm friendly, i.e., the functions used can work with this dataset
-#' and 2) Tells what's wrong, if any.
+#' Takes an activity dataset as input and gives information about 1) If a
+#' dataset is digiRhythm friendly, i.e., the functions used can work with this
+#' dataset and 2) Tells what's wrong, if any.
 #'
 #' @param data The dataframe containing the activity data
 #' @param verbose if TRUE, prints info about the dataset
@@ -26,10 +26,12 @@ is_dgm_friendly <- function(data, verbose = FALSE) {
   } else {
     # Checking if the first column meets the requirements
     if ("POSIXct" %in% class(data[, 1]) | "POSIXt" %in% class(data[, 1])) {
-      message <- paste0(crayon::green("v Correct time format: "), "First column has a POSIXct Format")
+      message <- paste0(crayon::green("v Correct time format: "), "First column
+                        has a POSIXct Format")
       print_v(message, verbose)
     } else {
-      message <- paste0(crayon::red("x Inorrect time format: "), "First column does not a POSIXct Format")
+      message <- paste0(crayon::red("x Inorrect time format: "), "First column
+                        does not a POSIXct Format")
       print_v(message, verbose)
       is_dgm <- FALSE
     }
@@ -37,27 +39,36 @@ is_dgm_friendly <- function(data, verbose = FALSE) {
 
     # Checking if data contains at least 7 days of data
     if (length(unique(as.Date(data[, 1]))) >= 2) {
-      message <- paste0(crayon::green("v Number of days: "), "Bigger or equal to 2")
+      message <- paste0(crayon::green("v Number of days: "), "Bigger or equal
+                        than 2")
       print_v(message, verbose)
     } else {
-      message <- paste0(crayon::red("x Number of days: "), "Less than 2 (minimum of 2 days is recommended)")
+      message <- paste0(crayon::red("x Number of days: "), "Less than 2 (minimum
+                        of 2 days is recommended)")
       print_v(message, verbose)
       is_dgm <- FALSE
     }
 
     # Checking if we have more than one column
     if (ncol(data) == 1) {
-      message <- paste0(crayon::red("x Illogical number of columns: "), "The dataset has only one column. Minimum number of columns is 2")
+      message <- paste0(crayon::red("x Illogical number of columns: "), "The
[…]

### 2024-07-25 R/lomb_scargle_periodogram.R

@@ -1,23 +1,26 @@
 #' Computes the Lomb Scargle Periodogram and returns the information needed for
-#' computing the DFC and HP. A plot visualizing the Harmonic Frequencies presence
-#' in the spectrum is possible. The function is inspired from the Lomb library in
-#' a great part, with modifications to fit the requirements of harmonic powers and
-#' computation of the DFC. This function is inspired by the lsp function from the
-#' lomb package and adapted to add different colors for harmonic and non harmonic
-#' frequencies in the signal. For more information about lomb::lsp, please refer
+#' computing the DFC and HP. A plot visualizing the Harmonic Frequencies
+#' presence in the spectrum is possible. The function is inspired from the Lomb
+#' library in a great part, with modifications to fit the requirements of
+#' harmonic powers and computation of the DFC. This function is inspired by the
+#' lsp function from the lomb package and adapted to add different colors for
+#' harmonic and non harmonic frequencies in the signal. For more information
+#' about lomb::lsp, please refer
 #' to: https://cran.r-project.org/web/packages/lomb/
 #'
 #' @param data a digiRhythm friendly dataframe of only two columns
 #' @param sampling the sampling period in minutes. default = 15 min.
 #' @param alpha the statistical significance for the false alarm
+#' @param harm_cutoff the order of the highest harmonic needed to be considered.
+#' An integer equal to 1, 2, 3, ... Default is 12.
 #' @param plot if TRUE, the LSP will be plotted
 #' @param extra_info_plot if True, extra information will be shown on the plot
 #'
 #' @return a list that contains a dataframe (detailed below), the significance
-#' level and alpha (for the record). The dataframe contains the power the frequency,
-#' the frequency in HZ, the p values according to Baluev 2008, the period that corresponds
-#' to the frequency in seconds and in hours and finally, a boolean to tell whether
-#' the frequency is harmonic or not.
+#' level and significance (for the record). The dataframe contains the power the
[…]

### 2024-07-25 R/resample_dgm.R

@@ -1,10 +1,12 @@
 #' Change the sampling of a digiRhythm friendly dataset
 #'
-#' This function upsamples the data but does not downsample them. The new sampling
-#' should be a multiple of the current sampling period, and should be given in minutes.
+#' This function upsamples the data but does not downsample them. The new
+#' sampling should be a multiple of the current sampling period, and should be
+#' given in minutes.
 #'
 #' @param data The dataframe containing the activity data
-#' @param new_sampling The new sampling (multiple of current sampling) in minutes
+#' @param new_sampling The new sampling (multiple of current sampling) in
+#' minutes
 #'
 #' @return A digiRhythm friendly dataset with the new sampling
 #'
@@ -26,7 +28,8 @@ resample_dgm <- function(data, new_sampling) {
   original_sampling <- xts::periodicity(xts_data)$frequency
 
   if ((new_sampling %% original_sampling) != 0) {
-    stop("The new sampling should be a multiple of the current sampling in minutes")
+    stop("The new sampling should be a multiple of the current sampling in
+         minutes")
   }
 
   if (new_sampling < original_sampling) {

### 2024-07-25 R/utils.R

@@ -97,9 +97,8 @@ dgm_periodicity <- function(data) {
 #' pbaluev (2008).
 #'
 
-# Copied from the LOMB library
 pbaluev <- function(Z, fmax, tm) {
-  # code copied from astropy timeseries (https://docs.astropy.org/en/stable/timeseries/index.html)
+  # Adapted from astropy timeseries (https://docs.astropy.org/en/stable/timeseries/index.html)
   N <- length(tm)
   Dt <- mean(tm^2) - mean(tm)^2
   NH <- N - 1
@@ -124,8 +123,27 @@ pbaluev <- function(Z, fmax, tm) {
 #' @return Returns the level given the p-value computed with pbaluev (2008).
 #'
 
-
 levopt <- function(Z, alpha, fmax, tm) {
   prob <- pbaluev(Z, fmax, tm)
   (log(prob) - log(alpha))^2
+}
+
+#' Function to calculate the smallest possible harmonic to consider given
+#' a sampling frequency. The minimum possible harmonic = 2 x the period of the
+#' maximum frequency according to the Shanon theorem. Example: if the sampling
+#' period is 15 min, the minimum possible treatable period is 30 minutes and that
+#' corresponds to the 48th harmonic (24 hours * 60 minutes / 48 = 30 minutes)
+#'
+#' @param sampling_period_in_minutes The sampling period of the acquired
+#' data in minutes
+#'
+#' @return Returns the smallest possible harmonic (of 24 hours) to consider given
+#' a sampling frequency.
+
+lowest_possible_harmonic_period <- function(sampling_period_in_minutes) {
+  harmonics <- seq(1, 1000)
+  harmonic_periods <- 24 * 60 / harmonics
+  all_reachable_harmonic_period <- harmonics[harmonic_periods >= 2 * sampling_period_in_minutes]
+  lowest_reachable_harmonic_period <- max(all_reachable_harmonic_period)
+  return(lowest_reachable_harmonic_period)
 }

### 2024-07-25 examples/code_snipets_manuscript.R

@@ -20,11 +20,11 @@ library(patchwork)
 #devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
 
 # To install the stable version from CRAN
-install.packages('digiRhythm')
+#install.packages('digiRhythm')
 library(digiRhythm)
 
 ###############################################################################
-################################## FIGURE: 2 ##################################
+################################## FIGURE: 1 ##################################
 ###############################################################################
 
 # LSP on purely synthetic data ----------------------------
@@ -272,7 +272,7 @@ ggsave(
 ################################## FIGURE: 4 ##################################
 ###############################################################################
 
-lomb_scargle_periodogram(df, alpha = 0.01, sampling = 15, plot = TRUE, extra_info_plot = TRUE)
+my_lsp <- lomb_scargle_periodogram(df, alpha = 0.01, sampling = 15, plot = TRUE, extra_info_plot = TRUE)
 
 name <- paste0('figures/Figure 4.png')
 ggsave(
@@ -352,7 +352,10 @@ td <- as.data.frame(timedata)
 my_sliding_di <- sliding_DI(df,
                             activity,
                             td)
-name <- paste0('figures/Figure 8.png')
+
+print(str(td))
+print(head(td))
+name <- paste0('./figures/Figure 8.png')
 ggsave(
   name,
   plot = lsp <- last_plot(),
@@ -391,7 +394,7 @@ ggsave(
 
 my_dfc +
   theme(
-    text=element_text(family="Times", size=8),
+    text=element_text(family="Times", size=20),
     axis.text.y=element_text(size=15, colour="red"))
 
 name <- paste0('figures/Figure 10.png')

### 2024-07-25 original_codes/original_lomb_scargle_periodogram.R

@@ -1,9 +1,30 @@
 library(pracma)
 library(digiRhythm)
 data("df516b_2", package = "digiRhythm")
+source("~/projects/digiRhythm/R/utils.R")
+library(ggplot2)
+library(tidyverse)
+
+# Inputs
+sampling = 15
+harm_cutoff <- 12
+theoretical_cutoff <- lowest_possible_harmonic_period(sampling)
+
+# Check if the needed cutoff harmonic is bigger than the theoretical cutoff
+if (harm_cutoff > theoretical_cutoff) {
+  warning("The sought harmonic cutoff is bigger than what is possible given the
+       sampling period. The cutoff harmonic should correspond to a period that
+       is at least 2 times the sampling period. For example, with a sampling
+       period of 15 min, the lowest possible period that can be treated is 30
+       min, which corresponds to the 48th harmonic period.")
+  print(paste0('changing the harmoinc cutoff to ', theoretical_cutoff))
+  used_harmonic_cutoff <- theoretical_cutoff
+} else {
+  used_harmonic_cutoff <- harm_cutoff
+}
 
 for(days in c(1:15)){
-  sampling = 15
+
   useful_len = (60/sampling)*24*days
   data <- df516b_2[1:useful_len, c(1,2)]
 
@@ -14,11 +35,6 @@ for(days in c(1:15)){
   if (!is_dgm_friendly(data, verbose = TRUE)) {
     stop('The data is not digiRhythm friendly. type ?is_dgm_friendly in your console for more information')
   }
-
-  # if (length(unique(as.Date(data[,1]))) != 7 ) {
-  #   warning('This LSP function is customized to use data with 7 days span only. But, the data contains less or more than 7 days.
-  #        A dynamic number of days will be introduced in a later version.')
-  # }
 
   x <- data
   start <- as.Date(x[1,1])
@@ -60,9 +76,8 @@ for(days in c(1:15)){
 
   n <- length(y)
   tspan <- t[n] - t[1]
-  fr.d <- 1/tspan
   step <- 1/(tspan * ofac)
-
+  step
 
   # if (is.null(to)) {
   f.max <- floor(0.5 * n * ofac) * step
@@ -71,6 +86,20 @@ for(days in c(1:15)){
   # }
 
   freq <- seq(fr.d, f.max, by = step)
+
[…]

### 2024-07-25 README.Rmd
Updated the vignettes and the readme page

--- 2021-10-29 README.Rmd
@@
 digiRhythm is an R library developed at Agroscope and provides a set of tools to 
-analyze and visualize the rhythmic behavior of animals.
-
-**TO DO CORE FUNCTIONALITIES:**
-
-- [X] Create a function that read Icetag data stored in a CSV file.
-- [X] Add the import_raw_icetage_data as an example to the README file.
-- [X] Create a function that computes the degree of functional coupling.
-- [X] Add the the dfc as an example to the README file.
-- [ ] Create a function that computes the diurnality index.
-- [ ] Add the the dfc as an example to the README file.
-
-**TO DO - VISUALIZATION FUNCTIONS:**
-
-- [X] Create a function that visualize the actogram.
-- [ ] Create a function that visualize the actogram + average activity.
-- [ ] Create a function that visualize the actogram + average activity + DFC.
-
-**TO DO - UTILITIES AND DOCUMENTATIONS:**
-
-- [X] Define data inside the library
-- [X] Add and example about the improt_raw_icetag_data to the README.
-- [ ] Create a function to test if a dataset is digiRhythm friendly.
-- [ ] Create a proper documentation for the data set.
-- [ ] configure data sets with lazy loading
-- [ ] Create test functions.
+analyze and rhythmic behavior of animals using the Degree of Functional Coupling. 
+It also provides a set of vizualization tools such as Actogram, Diurnality Index,
+Average Daily Activity, and Daily Activity Wrap Plot.
 
@@
 ```{r}
-#Uncomment the below two lines if you use the library for the first time of if
-#you want to update the library
-#install.packages("devtools")
-#devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
+# Uncomment the below two lines if you use the library for the first time of if
+# you want to update the library
+# install.packages("devtools")
+# devtools::install_github("nasserdr/digiRhythm", dependencies = TRUE)
 ```
+
 ## Example
 
-This is a basic example which shows you how to solve a common problem:
+You can access few examples from the vignettes provided along with the package.
 
-```{r example, message=FALSE, warning=FALSE, paged.print=FALSE}
-library(digiRhythm)
-
-#The file name with the path
-url <- 'https://github.com/nasserdr/digiRhythm_sample_datasets/raw/main/516b_2.csv'
-download.file(url, destfile = '516b_2.csv')
-filename <- file.path(getwd(), '516b_2.csv')
-
-#The columns that we are interested in
-colstoread <- c("Date", "Time", "Motion Index", 'Steps')
-
-#Reading the activity data from the csv file
-data <- import_raw_icetag_data(filename = filename, skipLines = , act.cols.names = colstoread, sampling = 15)
-
-print(head(data))
-
-```
-
-This is an example on how to visualize the actogram
-
-```{r visualize, message=FALSE, warning=FALSE, paged.print=FALSE}
-data("df516b_2", package = "digiRhythm")
-df <- remove_activity_outliers(df516b_2)
-df_act_info(df)
-activity = names(df)[2]
-start <- "2020-30-04"
-end <- "2020-06-05"
-save <- TRUE
-outputdir <- 'testresults'
-outplotname <- 'myplot'
-width <- 10
-device <- 'tiff'
-height <-  5
-actogram(data, activity, start, end, save = FALSE,
-     outputdir = 'testresults', outplotname = 'actoplot', width = 10,
-     height =  5, device = 'tiff')
-```
-
-This is an example on how to compute the degree of functional coupling.
-
-```{r dfc_example, results = FALSE, warning=FALSE, paged.print=FALSE}
-data("df516b_2", package = "digiRhythm")
-df <- remove_activity_outliers(df516b_2)
-df_act_info(df)
-activity = names(df)[2]
-my_dfc <- dfc(df, activity , sampling = 15, show_lsp_plot = FALSE)
-
-#You may want to explore the two list inside my_dfc.
-#DFC and SPECTRUM are saved inside my_dfc, each as a list
-```
-
-This is an example on how to compute the diurnality index:
-
-```{r dindex_example, results = FALSE, warning=FALSE, paged.print=FALSE}
-data("df516b_2", package = "digiRhythm")
-df <- remove_activity_outliers(df516b_2)
-df_act_info(df)
-activity = names(df)[2]
-d_index <- diurnality(data, activity, plot = TRUE)
-
-```
-
-This is an example on how you can resample your data:
-
-```{r resample, results = FALSE, warning=FALSE, paged.print=FALSE}
-data("df516b_2", package = "digiRhythm")
-df <- df516b_2
-df <- remove_activity_outliers(df)
-new_sampling <- 30
-new_dgm <- resample_dgm(df, new_sampling)
-
-```
+[Vignettes](https://cran.r-project.org/web/packages/digiRhythm/index.html)

### 2024-07-25 R/actogram.R

@@ -14,10 +14,7 @@
 #'
 #' @return A ggplot2 object that contains the actogram plot
 #'
-#' @importFrom magrittr %>%
-#' @importFrom stats time
 #' @import ggplot2
-#' @importFrom lubridate date
 #' @import dplyr
 #'
 #' @export

### 2024-07-25 R/dfc.R
Modified some styling

@@ -2,19 +2,20 @@
 #' Lomb-Scargle Spectrum (LSP Spec) for one variable in an activity dataset.
 #' The dataset should be digiRhythm friendly.
 #'
-#' The computation of DFC/HP/LSP parameters is done using a rolling window of 7 days (i.e.,
-#' first, we compute the parameters of Days 1-7 then, of days 2-8 and so on).
-#' For each window of the 7 days, the function will compute the LSP spectrum to
-#' determine the power of each frequency. Using Baluev (2008), we will compute the
-#' significance of the amplitude of each frequency component and determine whether
-#' it is significant or not. Then, we will have all the significant frequencies,
-#' whose amplitudes' summation will be denominated as SUMSIG. Among all the available
-#' frequencies, some are harmonic (those that correspond to waves of period
-#' 24h, 12h, 24h/3, 24h/4, ...). As a result, we will have frequency components
-#' that are significant and harmonic, whose powers' summation is called SSH (sum
-#' significant and harmonic). The summation of all frequency components up to a
-#' frequency reflecting a 24h period is called SUMALL. Therefore, DFC and HP are
-#' computed as follows:
+#' The computation of DFC/HP/LSP parameters is done using a rolling window. If
+#' the rolling window is 7 days, first, we compute the parameters of Days 1-7
+#' then, of days 2-8 and so on). For each window of the 7 days, the function
+#' will compute the LSP spectrum to determine the power of each frequency. Using
+#' Baluev (2008), we will compute the significance of the amplitude of each
+#' frequency component and determine whether it is significant or not. Then, we
+#' will have all the significant frequencies, whose amplitudes' summation will
+#' be denominated as SUMSIG. Among all the available frequencies, some are
+#' harmonic (those that correspond to waves of period 24h, 12h, 24h/3, 24h/4,
+#' ...). As a result, we will have frequency components that are significant and
+#' harmonic, whose powers' summation is called SSH (sum significant and
+#' harmonic). The summation of all frequency components up to a frequency
+#' reflecting a 24h period is called SUMALL. Therefore, DFC and HP are computed
+#' as follows:
 #'
 #' DFC <- SSH / SUMSIG
 #' HP <- SSH / SUMALL
@@ -28,7 +29,8 @@
[…]

### 2024-07-25 R/is_dgm_friendly.R

@@ -52,7 +52,8 @@ is_dgm_friendly <- function(data, verbose = FALSE) {
     # Checking if we have more than one column
     if (ncol(data) == 1) {
       message <- paste0(crayon::red("x Illogical number of columns: "), "The
-                        dataset has only one column. Minimum number of columns is 2")
+                        dataset has only one column. Minimum number of columns
+                        is 2")
       print_v(message, verbose)
       is_dgm <- FALSE
     } else {

### 2024-07-25 R/utils.R

@@ -93,12 +93,14 @@ dgm_periodicity <- function(data) {
 #' @param Z the power of the frequency
 #' @param fmax the maximum frequency in the spectrum
 #' @param tm the time grid of the original time series
-#' @return an intermediate calculation step needed to compute the p-value according to
+#' @return an intermediate calculation step needed to compute the p-value
+#' according to
 #' pbaluev (2008).
 #'
 
 pbaluev <- function(Z, fmax, tm) {
-  # Adapted from astropy timeseries (https://docs.astropy.org/en/stable/timeseries/index.html)
+  # Adapted from astropy timeseries
+  # (https://docs.astropy.org/en/stable/timeseries/index.html)
   N <- length(tm)
   Dt <- mean(tm^2) - mean(tm)^2
   NH <- N - 1
@@ -131,19 +133,21 @@ levopt <- function(Z, alpha, fmax, tm) {
 #' Function to calculate the smallest possible harmonic to consider given
 #' a sampling frequency. The minimum possible harmonic = 2 x the period of the
 #' maximum frequency according to the Shanon theorem. Example: if the sampling
-#' period is 15 min, the minimum possible treatable period is 30 minutes and that
-#' corresponds to the 48th harmonic (24 hours * 60 minutes / 48 = 30 minutes)
+#' period is 15 min, the minimum possible treatable period is 30 minutes and
+#' that corresponds to the 48th harmonic (24 hours * 60 minutes / 48 =
+#' 30 minutes)
 #'
 #' @param sampling_period_in_minutes The sampling period of the acquired
 #' data in minutes
 #'
-#' @return Returns the smallest possible harmonic (of 24 hours) to consider given
-#' a sampling frequency.
+#' @return Returns the smallest possible harmonic (of 24 hours) to consider
+#' given a sampling frequency.
 
 lowest_possible_harmonic_period <- function(sampling_period_in_minutes) {
   harmonics <- seq(1, 1000)
   harmonic_periods <- 24 * 60 / harmonics
-  all_reachable_harmonic_period <- harmonics[harmonic_periods >= 2 * sampling_period_in_minutes]
+  all_reachable_harmonic_period <- harmonics[harmonic_periods >=
+    2 * sampling_period_in_minutes]
   lowest_reachable_harmonic_period <- max(all_reachable_harmonic_period)
   return(lowest_reachable_harmonic_period)
 }
